### 2022-02-16 manuscript/include-figures.tex

\section{Tajima's D - Fay \& Hu's H}

\subsection{Site-specific mutation-selection codon models.} 
\begin{center}
\includegraphics[width=0.75\linewidth, page=1]{/home/thibault/Documents/SelCoeff/results/MutSel.D_tajima.heatmap.pdf} \\
\includegraphics[width=0.75\linewidth, page=1]{/home/thibault/Documents/SelCoeff/results/MutSel.D_tajima.scatter.pdf} \\
\includegraphics[width=0.75\linewidth, page=1]{/home/thibault/Documents/SelCoeff/results/MutSel.H_fay_wu.heatmap.pdf} \\
\includegraphics[width=0.75\linewidth, page=1]{/home/thibault/Documents/SelCoeff/results/MutSel.H_fay_wu.scatter.pdf} \\
\includegraphics[width=0.75\linewidth, page=1]{/home/thibault/Documents/SelCoeff/results/MutSel.fay_wu.heatmap.pdf} \\
\includegraphics[width=0.75\linewidth, page=1]{/home/thibault/Documents/SelCoeff/results/MutSel.fay_wu.scatter.pdf} \\
\includegraphics[width=0.75\linewidth, page=1]{/home/thibault/Documents/SelCoeff/results/MutSel.tajima.heatmap.pdf} \\
\includegraphics[width=0.75\linewidth, page=1]{/home/thibault/Documents/SelCoeff/results/MutSel.tajima.scatter.pdf} \\
\includegraphics[width=0.75\linewidth, page=1]{/home/thibault/Documents/SelCoeff/results/MutSel.watterson.heatmap.pdf} \\
\includegraphics[width=0.75\linewidth, page=1]{/home/thibault/Documents/SelCoeff/results/MutSel.watterson.scatter.pdf} \\
\end{center}
\section{Bos taurus}

\subsection{UGBT}

\subsubsection*{SFS for site-specific Mutation-Selection codon models.} 
\begin{minipage}{0.49\linewidth} 
\includegraphics[width=\linewidth, page=1]{/home/thibault/Documents/SelCoeff/analysis/Bos_taurus.UGBT.MutSel.pdf} 
\end{minipage}
\begin{minipage}{0.49\linewidth}
\includegraphics[width=\linewidth, page=1]{/home/thibault/Documents/SelCoeff/analysis/Bos_taurus.UGBT.MutSel.normalize.pdf} 
\end{minipage}
\\ 
\section{Chlorocebus sabaeus}

\subsection{Zambia}

\subsubsection*{SFS for site-specific Mutation-Selection codon models.} 
\begin{minipage}{0.49\linewidth} 
\includegraphics[width=\linewidth, page=1]{/home/thibault/Documents/SelCoeff/analysis/Chlorocebus_sabaeus.Zambia.MutSel.pdf} 
\end{minipage}
\begin{minipage}{0.49\linewidth}
\includegraphics[width=\linewidth, page=1]{/home/thibault/Documents/SelCoeff/analysis/Chlorocebus_sabaeus.Zambia.MutSel.normalize.pdf} 
\end{minipage}
\\ 
\section{Homo sapiens}

\subsection{AMR}

\subsubsection*{SFS for site-specific Mutation-Selection codon models.} 
\begin{minipage}{0.49\linewidth} 
\includegraphics[width=\linewidth, page=1]{/home/thibault/Documents/SelCoeff/analysis/Homo_sapiens.AMR.MutSel.pdf} 
\end{minipage}
\begin{minipage}{0.49\linewidth}
\includegraphics[width=\linewidth, page=1]{/home/thibault/Documents/SelCoeff/analysis/Homo_sapiens.AMR.MutSel.normalize.pdf} 
\end{minipage}
\\