### 2022-03-03 manuscript/include-figures.tex
Use polyDFE model C, nearly-neutral mutations as |S|<1

\section{Tajima's D - Fay \& Hu's H}

\subsection{Site-specific mutation-selection codon models.} 
\begin{center}
\includegraphics[width=0.75\linewidth, page=1]{/home/thibault/Documents/SelCoeff/results/DFE.MutSel.P-Seq0.heatmap.pdf} \\
\includegraphics[width=0.75\linewidth, page=1]{/home/thibault/Documents/SelCoeff/results/DFE.MutSel.P-Sinf0.heatmap.pdf} \\
\includegraphics[width=0.75\linewidth, page=1]{/home/thibault/Documents/SelCoeff/results/DFE.MutSel.P-Ssup0.heatmap.pdf} \\
\includegraphics[width=0.75\linewidth, page=1]{/home/thibault/Documents/SelCoeff/results/DFE.MutSel.S.heatmap.pdf} \\
\includegraphics[width=0.75\linewidth, page=1]{/home/thibault/Documents/SelCoeff/results/Theta.MutSel.D_tajima.heatmap.pdf} \\
\includegraphics[width=0.75\linewidth, page=1]{/home/thibault/Documents/SelCoeff/results/Theta.MutSel.H_fay_wu.heatmap.pdf} \\
\includegraphics[width=0.75\linewidth, page=1]{/home/thibault/Documents/SelCoeff/results/Theta.MutSel.fay_wu.heatmap.pdf} \\
\includegraphics[width=0.75\linewidth, page=1]{/home/thibault/Documents/SelCoeff/results/Theta.MutSel.tajima.heatmap.pdf} \\
\includegraphics[width=0.75\linewidth, page=1]{/home/thibault/Documents/SelCoeff/results/Theta.MutSel.watterson.heatmap.pdf} \\
\end{center}
\subsection{Site-specific codon models.} 
\begin{center}
\includegraphics[width=0.75\linewidth, page=1]{/home/thibault/Documents/SelCoeff/results/DFE.Omega.P-Seq0.heatmap.pdf} \\
\includegraphics[width=0.75\linewidth, page=1]{/home/thibault/Documents/SelCoeff/results/DFE.Omega.P-Sinf0.heatmap.pdf} \\
\includegraphics[width=0.75\linewidth, page=1]{/home/thibault/Documents/SelCoeff/results/DFE.Omega.P-Ssup0.heatmap.pdf} \\
\includegraphics[width=0.75\linewidth, page=1]{/home/thibault/Documents/SelCoeff/results/DFE.Omega.S.heatmap.pdf} \\
\includegraphics[width=0.75\linewidth, page=1]{/home/thibault/Documents/SelCoeff/results/Theta.Omega.D_tajima.heatmap.pdf} \\
\includegraphics[width=0.75\linewidth, page=1]{/home/thibault/Documents/SelCoeff/results/Theta.Omega.H_fay_wu.heatmap.pdf} \\
\includegraphics[width=0.75\linewidth, page=1]{/home/thibault/Documents/SelCoeff/results/Theta.Omega.fay_wu.heatmap.pdf} \\
\includegraphics[width=0.75\linewidth, page=1]{/home/thibault/Documents/SelCoeff/results/Theta.Omega.tajima.heatmap.pdf} \\
\includegraphics[width=0.75\linewidth, page=1]{/home/thibault/Documents/SelCoeff/results/Theta.Omega.watterson.heatmap.pdf} \\
\end{center}
\section{Bos taurus}

\subsection{IRBT}

\subsubsection*{Site-specific Mutation-Selection codon models.} 
\begin{minipage}{0.49\linewidth} 
\includegraphics[width=\linewidth, page=1]{/home/thibault/Documents/SelCoeff/analysis/Bos_taurus.IRBT.MutSel.pdf} 
\end{minipage}
\begin{minipage}{0.49\linewidth} 
\includegraphics[width=\linewidth, page=1]{/home/thibault/Documents/SelCoeff/analysis/Bos_taurus.IRBT.MutSel.normalize.pdf} 
\end{minipage}
\begin{minipage}{0.49\linewidth} 
\includegraphics[width=\linewidth, page=1]{/home/thibault/Documents/SelCoeff/analysis/Bos_taurus.IRBT.MutSel.polyDFE_C.pdf} 
\end{minipage}
\begin{minipage}{0.49\linewidth} 
\includegraphics[width=\linewidth, page=1]{/home/thibault/Documents/SelCoeff/analysis/Bos_taurus.IRBT.MutSel.histogram.pdf} 
\end{minipage}
\\ 
\subsubsection*{Site-specific codon models.} 
\begin{minipage}{0.49\linewidth} 
\includegraphics[width=\linewidth, page=1]{/home/thibault/Documents/SelCoeff/analysis/Bos_taurus.IRBT.Omega.pdf} 
\end{minipage}
\begin{minipage}{0.49\linewidth} 
\includegraphics[width=\linewidth, page=1]{/home/thibault/Documents/SelCoeff/analysis/Bos_taurus.IRBT.Omega.normalize.pdf} 
\end{minipage}
\begin{minipage}{0.49\linewidth} 
\includegraphics[width=\linewidth, page=1]{/home/thibault/Documents/SelCoeff/analysis/Bos_taurus.IRBT.Omega.polyDFE_C.pdf} 
\end{minipage}
\begin{minipage}{0.49\linewidth} 
\includegraphics[width=\linewidth, page=1]{/home/thibault/Documents/SelCoeff/analysis/Bos_taurus.IRBT.Omega.histogram.pdf} 
\end{minipage}
\\ 
\subsection{UGBT}

\subsubsection*{Site-specific Mutation-Selection codon models.} 
\begin{minipage}{0.49\linewidth} 
\includegraphics[width=\linewidth, page=1]{/home/thibault/Documents/SelCoeff/analysis/Bos_taurus.UGBT.MutSel.pdf} 
\end{minipage}
\begin{minipage}{0.49\linewidth} 
\includegraphics[width=\linewidth, page=1]{/home/thibault/Documents/SelCoeff/analysis/Bos_taurus.UGBT.MutSel.normalize.pdf} 
\end{minipage}
\begin{minipage}{0.49\linewidth} 
\includegraphics[width=\linewidth, page=1]{/home/thibault/Documents/SelCoeff/analysis/Bos_taurus.UGBT.MutSel.polyDFE_C.pdf} 
\end{minipage}
\begin{minipage}{0.49\linewidth} 
\includegraphics[width=\linewidth, page=1]{/home/thibault/Documents/SelCoeff/analysis/Bos_taurus.UGBT.MutSel.histogram.pdf} 
\end{minipage}
\\ 
\subsubsection*{Site-specific codon models.} 
\begin{minipage}{0.49\linewidth} 
\includegraphics[width=\linewidth, page=1]{/home/thibault/Documents/SelCoeff/analysis/Bos_taurus.UGBT.Omega.pdf} 
\end{minipage}
\begin{minipage}{0.49\linewidth} 
\includegraphics[width=\linewidth, page=1]{/home/thibault/Documents/SelCoeff/analysis/Bos_taurus.UGBT.Omega.normalize.pdf} 
\end{minipage}
\begin{minipage}{0.49\linewidth} 
\includegraphics[width=\linewidth, page=1]{/home/thibault/Documents/SelCoeff/analysis/Bos_taurus.UGBT.Omega.polyDFE_C.pdf} 
\end{minipage}
\begin{minipage}{0.49\linewidth} 
\includegraphics[width=\linewidth, page=1]{/home/thibault/Documents/SelCoeff/analysis/Bos_taurus.UGBT.Omega.histogram.pdf} 
\end{minipage}
\\ 
\section{Canis familiaris}

\subsection{Canis}

\subsubsection*{Site-specific Mutation-Selection codon models.} 
\begin{minipage}{0.49\linewidth} 
\includegraphics[width=\linewidth, page=1]{/home/thibault/Documents/SelCoeff/analysis/Canis_familiaris.dogs.MutSel.pdf} 
\end{minipage}
\begin{minipage}{0.49\linewidth} 
\includegraphics[width=\linewidth, page=1]{/home/thibault/Documents/SelCoeff/analysis/Canis_familiaris.dogs.MutSel.normalize.pdf} 
\end{minipage}
\begin{minipage}{0.49\linewidth} 
\includegraphics[width=\linewidth, page=1]{/home/thibault/Documents/SelCoeff/analysis/Canis_familiaris.dogs.MutSel.polyDFE_C.pdf} 
\end{minipage}
\begin{minipage}{0.49\linewidth} 
\includegraphics[width=\linewidth, page=1]{/home/thibault/Documents/SelCoeff/analysis/Canis_familiaris.dogs.MutSel.histogram.pdf} 
\end{minipage}
\\ 
\subsubsection*{Site-specific codon models.} 
\begin{minipage}{0.49\linewidth} 
\includegraphics[width=\linewidth, page=1]{/home/thibault/Documents/SelCoeff/analysis/Canis_familiaris.dogs.Omega.pdf} 
\end{minipage}
\begin{minipage}{0.49\linewidth} 
\includegraphics[width=\linewidth, page=1]{/home/thibault/Documents/SelCoeff/analysis/Canis_familiaris.dogs.Omega.normalize.pdf} 
\end{minipage}
\begin{minipage}{0.49\linewidth} 
\includegraphics[width=\linewidth, page=1]{/home/thibault/Documents/SelCoeff/analysis/Canis_familiaris.dogs.Omega.polyDFE_C.pdf} 
\end{minipage}
\begin{minipage}{0.49\linewidth} 
\includegraphics[width=\linewidth, page=1]{/home/thibault/Documents/SelCoeff/analysis/Canis_familiaris.dogs.Omega.histogram.pdf} 
\end{minipage}
\\ 
\section{Capra hircus}

\subsection{AUCH}

\subsubsection*{Site-specific Mutation-Selection codon models.} 
\begin{minipage}{0.49\linewidth} 
\includegraphics[width=\linewidth, page=1]{/home/thibault/Documents/SelCoeff/analysis/Capra_hircus.AUCH.MutSel.pdf} 
\end{minipage}
\begin{minipage}{0.49\linewidth} 
\includegraphics[width=\linewidth, page=1]{/home/thibault/Documents/SelCoeff/analysis/Capra_hircus.AUCH.MutSel.normalize.pdf} 
\end{minipage}
\begin{minipage}{0.49\linewidth} 
\includegraphics[width=\linewidth, page=1]{/home/thibault/Documents/SelCoeff/analysis/Capra_hircus.AUCH.MutSel.polyDFE_C.pdf} 
\end{minipage}
\begin{minipage}{0.49\linewidth} 
\includegraphics[width=\linewidth, page=1]{/home/thibault/Documents/SelCoeff/analysis/Capra_hircus.AUCH.MutSel.histogram.pdf} 
\end{minipage}
\\ 
\subsubsection*{Site-specific codon models.} 
\begin{minipage}{0.49\linewidth} 
\includegraphics[width=\linewidth, page=1]{/home/thibault/Documents/SelCoeff/analysis/Capra_hircus.AUCH.Omega.pdf} 
\end{minipage}
\begin{minipage}{0.49\linewidth} 
\includegraphics[width=\linewidth, page=1]{/home/thibault/Documents/SelCoeff/analysis/Capra_hircus.AUCH.Omega.normalize.pdf} 
\end{minipage}
\begin{minipage}{0.49\linewidth} 
\includegraphics[width=\linewidth, page=1]{/home/thibault/Documents/SelCoeff/analysis/Capra_hircus.AUCH.Omega.polyDFE_C.pdf} 
\end{minipage}
\begin{minipage}{0.49\linewidth} 
\includegraphics[width=\linewidth, page=1]{/home/thibault/Documents/SelCoeff/analysis/Capra_hircus.AUCH.Omega.histogram.pdf} 
\end{minipage}
\\ 
\subsection{FRCH}

\subsubsection*{Site-specific Mutation-Selection codon models.} 
\begin{minipage}{0.49\linewidth} 
\includegraphics[width=\linewidth, page=1]{/home/thibault/Documents/SelCoeff/analysis/Capra_hircus.FRCH.MutSel.pdf} 
\end{minipage}
\begin{minipage}{0.49\linewidth} 
\includegraphics[width=\linewidth, page=1]{/home/thibault/Documents/SelCoeff/analysis/Capra_hircus.FRCH.MutSel.normalize.pdf} 
\end{minipage}
\begin{minipage}{0.49\linewidth} 
\includegraphics[width=\linewidth, page=1]{/home/thibault/Documents/SelCoeff/analysis/Capra_hircus.FRCH.MutSel.polyDFE_C.pdf} 
\end{minipage}
\begin{minipage}{0.49\linewidth} 
\includegraphics[width=\linewidth, page=1]{/home/thibault/Documents/SelCoeff/analysis/Capra_hircus.FRCH.MutSel.histogram.pdf} 
\end{minipage}
\\ 
\subsubsection*{Site-specific codon models.} 
\begin{minipage}{0.49\linewidth} 
\includegraphics[width=\linewidth, page=1]{/home/thibault/Documents/SelCoeff/analysis/Capra_hircus.FRCH.Omega.pdf} 
\end{minipage}
\begin{minipage}{0.49\linewidth} 
\includegraphics[width=\linewidth, page=1]{/home/thibault/Documents/SelCoeff/analysis/Capra_hircus.FRCH.Omega.normalize.pdf} 
\end{minipage}
\begin{minipage}{0.49\linewidth} 
\includegraphics[width=\linewidth, page=1]{/home/thibault/Documents/SelCoeff/analysis/Capra_hircus.FRCH.Omega.polyDFE_C.pdf} 
\end{minipage}
\begin{minipage}{0.49\linewidth} 
\includegraphics[width=\linewidth, page=1]{/home/thibault/Documents/SelCoeff/analysis/Capra_hircus.FRCH.Omega.histogram.pdf} 
\end{minipage}
\\ 
\subsection{IRCA}

\subsubsection*{Site-specific Mutation-Selection codon models.} 
\begin{minipage}{0.49\linewidth} 
\includegraphics[width=\linewidth, page=1]{/home/thibault/Documents/SelCoeff/analysis/Capra_hircus.IRCA.MutSel.pdf} 
\end{minipage}
\begin{minipage}{0.49\linewidth} 
\includegraphics[width=\linewidth, page=1]{/home/thibault/Documents/SelCoeff/analysis/Capra_hircus.IRCA.MutSel.normalize.pdf} 
\end{minipage}
\begin{minipage}{0.49\linewidth} 
\includegraphics[width=\linewidth, page=1]{/home/thibault/Documents/SelCoeff/analysis/Capra_hircus.IRCA.MutSel.polyDFE_C.pdf} 
\end{minipage}
\begin{minipage}{0.49\linewidth} 
\includegraphics[width=\linewidth, page=1]{/home/thibault/Documents/SelCoeff/analysis/Capra_hircus.IRCA.MutSel.histogram.pdf} 
\end{minipage}
\\ 
\subsubsection*{Site-specific codon models.} 
\begin{minipage}{0.49\linewidth} 
\includegraphics[width=\linewidth, page=1]{/home/thibault/Documents/SelCoeff/analysis/Capra_hircus.IRCA.Omega.pdf} 
\end{minipage}
\begin{minipage}{0.49\linewidth} 
\includegraphics[width=\linewidth, page=1]{/home/thibault/Documents/SelCoeff/analysis/Capra_hircus.IRCA.Omega.normalize.pdf} 
\end{minipage}
\begin{minipage}{0.49\linewidth} 
\includegraphics[width=\linewidth, page=1]{/home/thibault/Documents/SelCoeff/analysis/Capra_hircus.IRCA.Omega.polyDFE_C.pdf} 
\end{minipage}
\begin{minipage}{0.49\linewidth} 
\includegraphics[width=\linewidth, page=1]{/home/thibault/Documents/SelCoeff/analysis/Capra_hircus.IRCA.Omega.histogram.pdf} 
\end{minipage}
\\ 
\subsection{IRCH}

\subsubsection*{Site-specific Mutation-Selection codon models.} 
\begin{minipage}{0.49\linewidth} 
\includegraphics[width=\linewidth, page=1]{/home/thibault/Documents/SelCoeff/analysis/Capra_hircus.IRCH.MutSel.pdf} 
\end{minipage}
\begin{minipage}{0.49\linewidth} 
\includegraphics[width=\linewidth, page=1]{/home/thibault/Documents/SelCoeff/analysis/Capra_hircus.IRCH.MutSel.normalize.pdf} 
\end{minipage}
\begin{minipage}{0.49\linewidth} 
\includegraphics[width=\linewidth, page=1]{/home/thibault/Documents/SelCoeff/analysis/Capra_hircus.IRCH.MutSel.polyDFE_C.pdf} 
\end{minipage}
\begin{minipage}{0.49\linewidth} 
\includegraphics[width=\linewidth, page=1]{/home/thibault/Documents/SelCoeff/analysis/Capra_hircus.IRCH.MutSel.histogram.pdf} 
\end{minipage}
\\ 
\subsubsection*{Site-specific codon models.} 
\begin{minipage}{0.49\linewidth} 
\includegraphics[width=\linewidth, page=1]{/home/thibault/Documents/SelCoeff/analysis/Capra_hircus.IRCH.Omega.pdf} 
\end{minipage}
\begin{minipage}{0.49\linewidth} 
\includegraphics[width=\linewidth, page=1]{/home/thibault/Documents/SelCoeff/analysis/Capra_hircus.IRCH.Omega.normalize.pdf} 
\end{minipage}
\begin{minipage}{0.49\linewidth} 
\includegraphics[width=\linewidth, page=1]{/home/thibault/Documents/SelCoeff/analysis/Capra_hircus.IRCH.Omega.polyDFE_C.pdf} 
\end{minipage}
\begin{minipage}{0.49\linewidth} 
\includegraphics[width=\linewidth, page=1]{/home/thibault/Documents/SelCoeff/analysis/Capra_hircus.IRCH.Omega.histogram.pdf} 
\end{minipage}
\\ 
\subsection{ITCH}

\subsubsection*{Site-specific Mutation-Selection codon models.} 
\begin{minipage}{0.49\linewidth} 
\includegraphics[width=\linewidth, page=1]{/home/thibault/Documents/SelCoeff/analysis/Capra_hircus.ITCH.MutSel.pdf} 
\end{minipage}
\begin{minipage}{0.49\linewidth} 
\includegraphics[width=\linewidth, page=1]{/home/thibault/Documents/SelCoeff/analysis/Capra_hircus.ITCH.MutSel.normalize.pdf} 
\end{minipage}
\begin{minipage}{0.49\linewidth} 
\includegraphics[width=\linewidth, page=1]{/home/thibault/Documents/SelCoeff/analysis/Capra_hircus.ITCH.MutSel.polyDFE_C.pdf} 
\end{minipage}
\begin{minipage}{0.49\linewidth} 
\includegraphics[width=\linewidth, page=1]{/home/thibault/Documents/SelCoeff/analysis/Capra_hircus.ITCH.MutSel.histogram.pdf} 
\end{minipage}
\\ 
\subsubsection*{Site-specific codon models.} 
\begin{minipage}{0.49\linewidth} 
\includegraphics[width=\linewidth, page=1]{/home/thibault/Documents/SelCoeff/analysis/Capra_hircus.ITCH.Omega.pdf} 
\end{minipage}
\begin{minipage}{0.49\linewidth} 
\includegraphics[width=\linewidth, page=1]{/home/thibault/Documents/SelCoeff/analysis/Capra_hircus.ITCH.Omega.normalize.pdf} 
\end{minipage}
\begin{minipage}{0.49\linewidth} 
\includegraphics[width=\linewidth, page=1]{/home/thibault/Documents/SelCoeff/analysis/Capra_hircus.ITCH.Omega.polyDFE_C.pdf} 
\end{minipage}
\begin{minipage}{0.49\linewidth} 
\includegraphics[width=\linewidth, page=1]{/home/thibault/Documents/SelCoeff/analysis/Capra_hircus.ITCH.Omega.histogram.pdf} 
\end{minipage}
\\ 
\subsection{MOCH}

\subsubsection*{Site-specific Mutation-Selection codon models.} 
\begin{minipage}{0.49\linewidth} 
\includegraphics[width=\linewidth, page=1]{/home/thibault/Documents/SelCoeff/analysis/Capra_hircus.MOCH.MutSel.pdf} 
\end{minipage}
\begin{minipage}{0.49\linewidth} 
\includegraphics[width=\linewidth, page=1]{/home/thibault/Documents/SelCoeff/analysis/Capra_hircus.MOCH.MutSel.normalize.pdf} 
\end{minipage}
\begin{minipage}{0.49\linewidth} 
\includegraphics[width=\linewidth, page=1]{/home/thibault/Documents/SelCoeff/analysis/Capra_hircus.MOCH.MutSel.polyDFE_C.pdf} 
\end{minipage}
\begin{minipage}{0.49\linewidth} 
\includegraphics[width=\linewidth, page=1]{/home/thibault/Documents/SelCoeff/analysis/Capra_hircus.MOCH.MutSel.histogram.pdf} 
\end{minipage}
\\ 
\subsubsection*{Site-specific codon models.} 
\begin{minipage}{0.49\linewidth} 
\includegraphics[width=\linewidth, page=1]{/home/thibault/Documents/SelCoeff/analysis/Capra_hircus.MOCH.Omega.pdf} 
\end{minipage}
\begin{minipage}{0.49\linewidth} 
\includegraphics[width=\linewidth, page=1]{/home/thibault/Documents/SelCoeff/analysis/Capra_hircus.MOCH.Omega.normalize.pdf} 
\end{minipage}
\begin{minipage}{0.49\linewidth} 
\includegraphics[width=\linewidth, page=1]{/home/thibault/Documents/SelCoeff/analysis/Capra_hircus.MOCH.Omega.polyDFE_C.pdf} 
\end{minipage}
\begin{minipage}{0.49\linewidth} 
\includegraphics[width=\linewidth, page=1]{/home/thibault/Documents/SelCoeff/analysis/Capra_hircus.MOCH.Omega.histogram.pdf} 
\end{minipage}
\\ 
\section{Chlorocebus sabaeus}

\subsection{Barbados}

\subsubsection*{Site-specific Mutation-Selection codon models.} 
\begin{minipage}{0.49\linewidth} 
\includegraphics[width=\linewidth, page=1]{/home/thibault/Documents/SelCoeff/analysis/Chlorocebus_sabaeus.Barbados.MutSel.pdf} 
\end{minipage}
\begin{minipage}{0.49\linewidth} 
\includegraphics[width=\linewidth, page=1]{/home/thibault/Documents/SelCoeff/analysis/Chlorocebus_sabaeus.Barbados.MutSel.normalize.pdf} 
\end{minipage}
\begin{minipage}{0.49\linewidth} 
\includegraphics[width=\linewidth, page=1]{/home/thibault/Documents/SelCoeff/analysis/Chlorocebus_sabaeus.Barbados.MutSel.polyDFE_C.pdf} 
\end{minipage}
\begin{minipage}{0.49\linewidth} 
\includegraphics[width=\linewidth, page=1]{/home/thibault/Documents/SelCoeff/analysis/Chlorocebus_sabaeus.Barbados.MutSel.histogram.pdf} 
\end{minipage}
\\ 
\subsubsection*{Site-specific codon models.} 
\begin{minipage}{0.49\linewidth} 
\includegraphics[width=\linewidth, page=1]{/home/thibault/Documents/SelCoeff/analysis/Chlorocebus_sabaeus.Barbados.Omega.pdf} 
\end{minipage}
\begin{minipage}{0.49\linewidth} 
\includegraphics[width=\linewidth, page=1]{/home/thibault/Documents/SelCoeff/analysis/Chlorocebus_sabaeus.Barbados.Omega.normalize.pdf} 
\end{minipage}
\begin{minipage}{0.49\linewidth} 
\includegraphics[width=\linewidth, page=1]{/home/thibault/Documents/SelCoeff/analysis/Chlorocebus_sabaeus.Barbados.Omega.polyDFE_C.pdf} 
\end{minipage}
\begin{minipage}{0.49\linewidth} 
\includegraphics[width=\linewidth, page=1]{/home/thibault/Documents/SelCoeff/analysis/Chlorocebus_sabaeus.Barbados.Omega.histogram.pdf} 
\end{minipage}
\\ 
\subsection{Central African Republic (CAR)}

\subsubsection*{Site-specific Mutation-Selection codon models.} 
\begin{minipage}{0.49\linewidth} 
\includegraphics[width=\linewidth, page=1]{/home/thibault/Documents/SelCoeff/analysis/Chlorocebus_sabaeus.Central_African_Republic.MutSel.pdf} 
\end{minipage}
\begin{minipage}{0.49\linewidth} 
\includegraphics[width=\linewidth, page=1]{/home/thibault/Documents/SelCoeff/analysis/Chlorocebus_sabaeus.Central_African_Republic.MutSel.normalize.pdf} 
\end{minipage}
\begin{minipage}{0.49\linewidth} 
\includegraphics[width=\linewidth, page=1]{/home/thibault/Documents/SelCoeff/analysis/Chlorocebus_sabaeus.Central_African_Republic.MutSel.polyDFE_C.pdf} 
\end{minipage}
\begin{minipage}{0.49\linewidth} 
\includegraphics[width=\linewidth, page=1]{/home/thibault/Documents/SelCoeff/analysis/Chlorocebus_sabaeus.Central_African_Republic.MutSel.histogram.pdf} 
\end{minipage}
\\ 
\subsubsection*{Site-specific codon models.} 
\begin{minipage}{0.49\linewidth} 
\includegraphics[width=\linewidth, page=1]{/home/thibault/Documents/SelCoeff/analysis/Chlorocebus_sabaeus.Central_African_Republic.Omega.pdf} 
\end{minipage}
\begin{minipage}{0.49\linewidth} 
\includegraphics[width=\linewidth, page=1]{/home/thibault/Documents/SelCoeff/analysis/Chlorocebus_sabaeus.Central_African_Republic.Omega.normalize.pdf} 
\end{minipage}
\begin{minipage}{0.49\linewidth} 
\includegraphics[width=\linewidth, page=1]{/home/thibault/Documents/SelCoeff/analysis/Chlorocebus_sabaeus.Central_African_Republic.Omega.polyDFE_C.pdf} 
\end{minipage}
\begin{minipage}{0.49\linewidth} 
\includegraphics[width=\linewidth, page=1]{/home/thibault/Documents/SelCoeff/analysis/Chlorocebus_sabaeus.Central_African_Republic.Omega.histogram.pdf} 
\end{minipage}
\\ 
\subsection{Ethiopia}

\subsubsection*{Site-specific Mutation-Selection codon models.} 
\begin{minipage}{0.49\linewidth} 
\includegraphics[width=\linewidth, page=1]{/home/thibault/Documents/SelCoeff/analysis/Chlorocebus_sabaeus.Ethiopia.MutSel.pdf} 
\end{minipage}
\begin{minipage}{0.49\linewidth} 
\includegraphics[width=\linewidth, page=1]{/home/thibault/Documents/SelCoeff/analysis/Chlorocebus_sabaeus.Ethiopia.MutSel.normalize.pdf} 
\end{minipage}
\begin{minipage}{0.49\linewidth} 
\includegraphics[width=\linewidth, page=1]{/home/thibault/Documents/SelCoeff/analysis/Chlorocebus_sabaeus.Ethiopia.MutSel.polyDFE_C.pdf} 
\end{minipage}
\begin{minipage}{0.49\linewidth} 
\includegraphics[width=\linewidth, page=1]{/home/thibault/Documents/SelCoeff/analysis/Chlorocebus_sabaeus.Ethiopia.MutSel.histogram.pdf} 
\end{minipage}
\\ 
\subsubsection*{Site-specific codon models.} 
\begin{minipage}{0.49\linewidth} 
\includegraphics[width=\linewidth, page=1]{/home/thibault/Documents/SelCoeff/analysis/Chlorocebus_sabaeus.Ethiopia.Omega.pdf} 
\end{minipage}
\begin{minipage}{0.49\linewidth} 
\includegraphics[width=\linewidth, page=1]{/home/thibault/Documents/SelCoeff/analysis/Chlorocebus_sabaeus.Ethiopia.Omega.normalize.pdf} 
\end{minipage}
\begin{minipage}{0.49\linewidth} 
\includegraphics[width=\linewidth, page=1]{/home/thibault/Documents/SelCoeff/analysis/Chlorocebus_sabaeus.Ethiopia.Omega.polyDFE_C.pdf} 
\end{minipage}
\begin{minipage}{0.49\linewidth} 
\includegraphics[width=\linewidth, page=1]{/home/thibault/Documents/SelCoeff/analysis/Chlorocebus_sabaeus.Ethiopia.Omega.histogram.pdf} 
\end{minipage}
\\ 
\subsection{Gambia}

\subsubsection*{Site-specific Mutation-Selection codon models.} 
\begin{minipage}{0.49\linewidth} 
\includegraphics[width=\linewidth, page=1]{/home/thibault/Documents/SelCoeff/analysis/Chlorocebus_sabaeus.Gambia.MutSel.pdf} 
\end{minipage}
\begin{minipage}{0.49\linewidth} 
\includegraphics[width=\linewidth, page=1]{/home/thibault/Documents/SelCoeff/analysis/Chlorocebus_sabaeus.Gambia.MutSel.normalize.pdf} 
\end{minipage}
\begin{minipage}{0.49\linewidth} 
\includegraphics[width=\linewidth, page=1]{/home/thibault/Documents/SelCoeff/analysis/Chlorocebus_sabaeus.Gambia.MutSel.polyDFE_C.pdf} 
\end{minipage}
\begin{minipage}{0.49\linewidth} 
\includegraphics[width=\linewidth, page=1]{/home/thibault/Documents/SelCoeff/analysis/Chlorocebus_sabaeus.Gambia.MutSel.histogram.pdf} 
\end{minipage}
\\ 
\subsubsection*{Site-specific codon models.} 
\begin{minipage}{0.49\linewidth} 
\includegraphics[width=\linewidth, page=1]{/home/thibault/Documents/SelCoeff/analysis/Chlorocebus_sabaeus.Gambia.Omega.pdf} 
\end{minipage}
\begin{minipage}{0.49\linewidth} 
\includegraphics[width=\linewidth, page=1]{/home/thibault/Documents/SelCoeff/analysis/Chlorocebus_sabaeus.Gambia.Omega.normalize.pdf} 
\end{minipage}
\begin{minipage}{0.49\linewidth} 
\includegraphics[width=\linewidth, page=1]{/home/thibault/Documents/SelCoeff/analysis/Chlorocebus_sabaeus.Gambia.Omega.polyDFE_C.pdf} 
\end{minipage}
\begin{minipage}{0.49\linewidth} 
\includegraphics[width=\linewidth, page=1]{/home/thibault/Documents/SelCoeff/analysis/Chlorocebus_sabaeus.Gambia.Omega.histogram.pdf} 
\end{minipage}
\\ 
\subsection{Kenya}

\subsubsection*{Site-specific Mutation-Selection codon models.} 
\begin{minipage}{0.49\linewidth} 
\includegraphics[width=\linewidth, page=1]{/home/thibault/Documents/SelCoeff/analysis/Chlorocebus_sabaeus.Kenya.MutSel.pdf} 
\end{minipage}
\begin{minipage}{0.49\linewidth} 
\includegraphics[width=\linewidth, page=1]{/home/thibault/Documents/SelCoeff/analysis/Chlorocebus_sabaeus.Kenya.MutSel.normalize.pdf} 
\end{minipage}
\begin{minipage}{0.49\linewidth} 
\includegraphics[width=\linewidth, page=1]{/home/thibault/Documents/SelCoeff/analysis/Chlorocebus_sabaeus.Kenya.MutSel.polyDFE_C.pdf} 
\end{minipage}
\begin{minipage}{0.49\linewidth} 
\includegraphics[width=\linewidth, page=1]{/home/thibault/Documents/SelCoeff/analysis/Chlorocebus_sabaeus.Kenya.MutSel.histogram.pdf} 
\end{minipage}
\\ 
\subsubsection*{Site-specific codon models.} 
\begin{minipage}{0.49\linewidth} 
\includegraphics[width=\linewidth, page=1]{/home/thibault/Documents/SelCoeff/analysis/Chlorocebus_sabaeus.Kenya.Omega.pdf} 
\end{minipage}
\begin{minipage}{0.49\linewidth} 
\includegraphics[width=\linewidth, page=1]{/home/thibault/Documents/SelCoeff/analysis/Chlorocebus_sabaeus.Kenya.Omega.normalize.pdf} 
\end{minipage}
\begin{minipage}{0.49\linewidth} 
\includegraphics[width=\linewidth, page=1]{/home/thibault/Documents/SelCoeff/analysis/Chlorocebus_sabaeus.Kenya.Omega.polyDFE_C.pdf} 
\end{minipage}
\begin{minipage}{0.49\linewidth} 
\includegraphics[width=\linewidth, page=1]{/home/thibault/Documents/SelCoeff/analysis/Chlorocebus_sabaeus.Kenya.Omega.histogram.pdf} 
\end{minipage}
\\ 
\subsection{Nevis}

\subsubsection*{Site-specific Mutation-Selection codon models.} 
\begin{minipage}{0.49\linewidth} 
\includegraphics[width=\linewidth, page=1]{/home/thibault/Documents/SelCoeff/analysis/Chlorocebus_sabaeus.Nevis.MutSel.pdf} 
\end{minipage}
\begin{minipage}{0.49\linewidth} 
\includegraphics[width=\linewidth, page=1]{/home/thibault/Documents/SelCoeff/analysis/Chlorocebus_sabaeus.Nevis.MutSel.normalize.pdf} 
\end{minipage}
\begin{minipage}{0.49\linewidth} 
\includegraphics[width=\linewidth, page=1]{/home/thibault/Documents/SelCoeff/analysis/Chlorocebus_sabaeus.Nevis.MutSel.polyDFE_C.pdf} 
\end{minipage}
\begin{minipage}{0.49\linewidth} 
\includegraphics[width=\linewidth, page=1]{/home/thibault/Documents/SelCoeff/analysis/Chlorocebus_sabaeus.Nevis.MutSel.histogram.pdf} 
\end{minipage}
\\ 
\subsubsection*{Site-specific codon models.} 
\begin{minipage}{0.49\linewidth} 
\includegraphics[width=\linewidth, page=1]{/home/thibault/Documents/SelCoeff/analysis/Chlorocebus_sabaeus.Nevis.Omega.pdf} 
\end{minipage}
\begin{minipage}{0.49\linewidth} 
\includegraphics[width=\linewidth, page=1]{/home/thibault/Documents/SelCoeff/analysis/Chlorocebus_sabaeus.Nevis.Omega.normalize.pdf} 
\end{minipage}
\begin{minipage}{0.49\linewidth} 
\includegraphics[width=\linewidth, page=1]{/home/thibault/Documents/SelCoeff/analysis/Chlorocebus_sabaeus.Nevis.Omega.polyDFE_C.pdf} 
\end{minipage}
\begin{minipage}{0.49\linewidth} 
\includegraphics[width=\linewidth, page=1]{/home/thibault/Documents/SelCoeff/analysis/Chlorocebus_sabaeus.Nevis.Omega.histogram.pdf} 
\end{minipage}
\\ 
\subsection{Saint Kitts (SK)}

\subsubsection*{Site-specific Mutation-Selection codon models.} 
\begin{minipage}{0.49\linewidth} 
\includegraphics[width=\linewidth, page=1]{/home/thibault/Documents/SelCoeff/analysis/Chlorocebus_sabaeus.Saint_Kitts.MutSel.pdf} 
\end{minipage}
\begin{minipage}{0.49\linewidth} 
\includegraphics[width=\linewidth, page=1]{/home/thibault/Documents/SelCoeff/analysis/Chlorocebus_sabaeus.Saint_Kitts.MutSel.normalize.pdf} 
\end{minipage}
\begin{minipage}{0.49\linewidth} 
\includegraphics[width=\linewidth, page=1]{/home/thibault/Documents/SelCoeff/analysis/Chlorocebus_sabaeus.Saint_Kitts.MutSel.polyDFE_C.pdf} 
\end{minipage}
\begin{minipage}{0.49\linewidth} 
\includegraphics[width=\linewidth, page=1]{/home/thibault/Documents/SelCoeff/analysis/Chlorocebus_sabaeus.Saint_Kitts.MutSel.histogram.pdf} 
\end{minipage}
\\ 
\subsubsection*{Site-specific codon models.} 
\begin{minipage}{0.49\linewidth} 
\includegraphics[width=\linewidth, page=1]{/home/thibault/Documents/SelCoeff/analysis/Chlorocebus_sabaeus.Saint_Kitts.Omega.pdf} 
\end{minipage}
\begin{minipage}{0.49\linewidth} 
\includegraphics[width=\linewidth, page=1]{/home/thibault/Documents/SelCoeff/analysis/Chlorocebus_sabaeus.Saint_Kitts.Omega.normalize.pdf} 
\end{minipage}
\begin{minipage}{0.49\linewidth} 
\includegraphics[width=\linewidth, page=1]{/home/thibault/Documents/SelCoeff/analysis/Chlorocebus_sabaeus.Saint_Kitts.Omega.polyDFE_C.pdf} 
\end{minipage}
\begin{minipage}{0.49\linewidth} 
\includegraphics[width=\linewidth, page=1]{/home/thibault/Documents/SelCoeff/analysis/Chlorocebus_sabaeus.Saint_Kitts.Omega.histogram.pdf} 
\end{minipage}
\\ 
\subsection{South Africa (SA)}

\subsubsection*{Site-specific Mutation-Selection codon models.} 
\begin{minipage}{0.49\linewidth} 
\includegraphics[width=\linewidth, page=1]{/home/thibault/Documents/SelCoeff/analysis/Chlorocebus_sabaeus.South_Africa.MutSel.pdf} 
\end{minipage}
\begin{minipage}{0.49\linewidth} 
\includegraphics[width=\linewidth, page=1]{/home/thibault/Documents/SelCoeff/analysis/Chlorocebus_sabaeus.South_Africa.MutSel.normalize.pdf} 
\end{minipage}
\begin{minipage}{0.49\linewidth} 
\includegraphics[width=\linewidth, page=1]{/home/thibault/Documents/SelCoeff/analysis/Chlorocebus_sabaeus.South_Africa.MutSel.polyDFE_C.pdf} 
\end{minipage}
\begin{minipage}{0.49\linewidth} 
\includegraphics[width=\linewidth, page=1]{/home/thibault/Documents/SelCoeff/analysis/Chlorocebus_sabaeus.South_Africa.MutSel.histogram.pdf} 
\end{minipage}
\\ 
\subsubsection*{Site-specific codon models.} 
\begin{minipage}{0.49\linewidth} 
\includegraphics[width=\linewidth, page=1]{/home/thibault/Documents/SelCoeff/analysis/Chlorocebus_sabaeus.South_Africa.Omega.pdf} 
\end{minipage}
\begin{minipage}{0.49\linewidth} 
\includegraphics[width=\linewidth, page=1]{/home/thibault/Documents/SelCoeff/analysis/Chlorocebus_sabaeus.South_Africa.Omega.normalize.pdf} 
\end{minipage}
\begin{minipage}{0.49\linewidth} 
\includegraphics[width=\linewidth, page=1]{/home/thibault/Documents/SelCoeff/analysis/Chlorocebus_sabaeus.South_Africa.Omega.polyDFE_C.pdf} 
\end{minipage}
\begin{minipage}{0.49\linewidth} 
\includegraphics[width=\linewidth, page=1]{/home/thibault/Documents/SelCoeff/analysis/Chlorocebus_sabaeus.South_Africa.Omega.histogram.pdf} 
\end{minipage}
\\ 
\subsection{Zambia}

\subsubsection*{Site-specific Mutation-Selection codon models.} 
\begin{minipage}{0.49\linewidth} 
\includegraphics[width=\linewidth, page=1]{/home/thibault/Documents/SelCoeff/analysis/Chlorocebus_sabaeus.Zambia.MutSel.pdf} 
\end{minipage}
\begin{minipage}{0.49\linewidth} 
\includegraphics[width=\linewidth, page=1]{/home/thibault/Documents/SelCoeff/analysis/Chlorocebus_sabaeus.Zambia.MutSel.normalize.pdf} 
\end{minipage}
\begin{minipage}{0.49\linewidth} 
\includegraphics[width=\linewidth, page=1]{/home/thibault/Documents/SelCoeff/analysis/Chlorocebus_sabaeus.Zambia.MutSel.polyDFE_C.pdf} 
\end{minipage}
\begin{minipage}{0.49\linewidth} 
\includegraphics[width=\linewidth, page=1]{/home/thibault/Documents/SelCoeff/analysis/Chlorocebus_sabaeus.Zambia.MutSel.histogram.pdf} 
\end{minipage}
\\ 
\subsubsection*{Site-specific codon models.} 
\begin{minipage}{0.49\linewidth} 
\includegraphics[width=\linewidth, page=1]{/home/thibault/Documents/SelCoeff/analysis/Chlorocebus_sabaeus.Zambia.Omega.pdf} 
\end{minipage}
\begin{minipage}{0.49\linewidth} 
\includegraphics[width=\linewidth, page=1]{/home/thibault/Documents/SelCoeff/analysis/Chlorocebus_sabaeus.Zambia.Omega.normalize.pdf} 
\end{minipage}
\begin{minipage}{0.49\linewidth} 
\includegraphics[width=\linewidth, page=1]{/home/thibault/Documents/SelCoeff/analysis/Chlorocebus_sabaeus.Zambia.Omega.polyDFE_C.pdf} 
\end{minipage}
\begin{minipage}{0.49\linewidth} 
\includegraphics[width=\linewidth, page=1]{/home/thibault/Documents/SelCoeff/analysis/Chlorocebus_sabaeus.Zambia.Omega.histogram.pdf} 
\end{minipage}
\\ 
\section{Equus caballus}

\subsection{Equus}

\subsubsection*{Site-specific Mutation-Selection codon models.} 
\begin{minipage}{0.49\linewidth} 
\includegraphics[width=\linewidth, page=1]{/home/thibault/Documents/SelCoeff/analysis/Equus_caballus.up.MutSel.pdf} 
\end{minipage}
\begin{minipage}{0.49\linewidth} 
\includegraphics[width=\linewidth, page=1]{/home/thibault/Documents/SelCoeff/analysis/Equus_caballus.up.MutSel.normalize.pdf} 
\end{minipage}
\begin{minipage}{0.49\linewidth} 
\includegraphics[width=\linewidth, page=1]{/home/thibault/Documents/SelCoeff/analysis/Equus_caballus.up.MutSel.polyDFE_C.pdf} 
\end{minipage}
\begin{minipage}{0.49\linewidth} 
\includegraphics[width=\linewidth, page=1]{/home/thibault/Documents/SelCoeff/analysis/Equus_caballus.up.MutSel.histogram.pdf} 
\end{minipage}
\\ 
\subsubsection*{Site-specific codon models.} 
\begin{minipage}{0.49\linewidth} 
\includegraphics[width=\linewidth, page=1]{/home/thibault/Documents/SelCoeff/analysis/Equus_caballus.up.Omega.pdf} 
\end{minipage}
\begin{minipage}{0.49\linewidth} 
\includegraphics[width=\linewidth, page=1]{/home/thibault/Documents/SelCoeff/analysis/Equus_caballus.up.Omega.normalize.pdf} 
\end{minipage}
\begin{minipage}{0.49\linewidth} 
\includegraphics[width=\linewidth, page=1]{/home/thibault/Documents/SelCoeff/analysis/Equus_caballus.up.Omega.polyDFE_C.pdf} 
\end{minipage}
\begin{minipage}{0.49\linewidth} 
\includegraphics[width=\linewidth, page=1]{/home/thibault/Documents/SelCoeff/analysis/Equus_caballus.up.Omega.histogram.pdf} 
\end{minipage}
\\ 
\section{Homo sapiens}

\subsection{AFR}

\subsubsection*{Site-specific Mutation-Selection codon models.} 
\begin{minipage}{0.49\linewidth} 
\includegraphics[width=\linewidth, page=1]{/home/thibault/Documents/SelCoeff/analysis/Homo_sapiens.AFR.MutSel.pdf} 
\end{minipage}
\begin{minipage}{0.49\linewidth} 
\includegraphics[width=\linewidth, page=1]{/home/thibault/Documents/SelCoeff/analysis/Homo_sapiens.AFR.MutSel.normalize.pdf} 
\end{minipage}
\begin{minipage}{0.49\linewidth} 
\includegraphics[width=\linewidth, page=1]{/home/thibault/Documents/SelCoeff/analysis/Homo_sapiens.AFR.MutSel.polyDFE_C.pdf} 
\end{minipage}
\begin{minipage}{0.49\linewidth} 
\includegraphics[width=\linewidth, page=1]{/home/thibault/Documents/SelCoeff/analysis/Homo_sapiens.AFR.MutSel.histogram.pdf} 
\end{minipage}
\\ 
\subsubsection*{Site-specific codon models.} 
\begin{minipage}{0.49\linewidth} 
\includegraphics[width=\linewidth, page=1]{/home/thibault/Documents/SelCoeff/analysis/Homo_sapiens.AFR.Omega.pdf} 
\end{minipage}
\begin{minipage}{0.49\linewidth} 
\includegraphics[width=\linewidth, page=1]{/home/thibault/Documents/SelCoeff/analysis/Homo_sapiens.AFR.Omega.normalize.pdf} 
\end{minipage}
\begin{minipage}{0.49\linewidth} 
\includegraphics[width=\linewidth, page=1]{/home/thibault/Documents/SelCoeff/analysis/Homo_sapiens.AFR.Omega.polyDFE_C.pdf} 
\end{minipage}
\begin{minipage}{0.49\linewidth} 
\includegraphics[width=\linewidth, page=1]{/home/thibault/Documents/SelCoeff/analysis/Homo_sapiens.AFR.Omega.histogram.pdf} 
\end{minipage}
\\ 
\subsection{AMR}

\subsubsection*{Site-specific Mutation-Selection codon models.} 
\begin{minipage}{0.49\linewidth} 
\includegraphics[width=\linewidth, page=1]{/home/thibault/Documents/SelCoeff/analysis/Homo_sapiens.AMR.MutSel.pdf} 
\end{minipage}
\begin{minipage}{0.49\linewidth} 
\includegraphics[width=\linewidth, page=1]{/home/thibault/Documents/SelCoeff/analysis/Homo_sapiens.AMR.MutSel.normalize.pdf} 
\end{minipage}
\begin{minipage}{0.49\linewidth} 
\includegraphics[width=\linewidth, page=1]{/home/thibault/Documents/SelCoeff/analysis/Homo_sapiens.AMR.MutSel.polyDFE_C.pdf} 
\end{minipage}
\begin{minipage}{0.49\linewidth} 
\includegraphics[width=\linewidth, page=1]{/home/thibault/Documents/SelCoeff/analysis/Homo_sapiens.AMR.MutSel.histogram.pdf} 
\end{minipage}
\\ 
\subsubsection*{Site-specific codon models.} 
\begin{minipage}{0.49\linewidth} 
\includegraphics[width=\linewidth, page=1]{/home/thibault/Documents/SelCoeff/analysis/Homo_sapiens.AMR.Omega.pdf} 
\end{minipage}
\begin{minipage}{0.49\linewidth} 
\includegraphics[width=\linewidth, page=1]{/home/thibault/Documents/SelCoeff/analysis/Homo_sapiens.AMR.Omega.normalize.pdf} 
\end{minipage}
\begin{minipage}{0.49\linewidth} 
\includegraphics[width=\linewidth, page=1]{/home/thibault/Documents/SelCoeff/analysis/Homo_sapiens.AMR.Omega.polyDFE_C.pdf} 
\end{minipage}
\begin{minipage}{0.49\linewidth} 
\includegraphics[width=\linewidth, page=1]{/home/thibault/Documents/SelCoeff/analysis/Homo_sapiens.AMR.Omega.histogram.pdf} 
\end{minipage}
\\ 
\subsection{EAS}

\subsubsection*{Site-specific Mutation-Selection codon models.} 
\begin{minipage}{0.49\linewidth} 
\includegraphics[width=\linewidth, page=1]{/home/thibault/Documents/SelCoeff/analysis/Homo_sapiens.EAS.MutSel.pdf} 
\end{minipage}
\begin{minipage}{0.49\linewidth} 
\includegraphics[width=\linewidth, page=1]{/home/thibault/Documents/SelCoeff/analysis/Homo_sapiens.EAS.MutSel.normalize.pdf} 
\end{minipage}
\begin{minipage}{0.49\linewidth} 
\includegraphics[width=\linewidth, page=1]{/home/thibault/Documents/SelCoeff/analysis/Homo_sapiens.EAS.MutSel.polyDFE_C.pdf} 
\end{minipage}
\begin{minipage}{0.49\linewidth} 
\includegraphics[width=\linewidth, page=1]{/home/thibault/Documents/SelCoeff/analysis/Homo_sapiens.EAS.MutSel.histogram.pdf} 
\end{minipage}
\\ 
\subsubsection*{Site-specific codon models.} 
\begin{minipage}{0.49\linewidth} 
\includegraphics[width=\linewidth, page=1]{/home/thibault/Documents/SelCoeff/analysis/Homo_sapiens.EAS.Omega.pdf} 
\end{minipage}
\begin{minipage}{0.49\linewidth} 
\includegraphics[width=\linewidth, page=1]{/home/thibault/Documents/SelCoeff/analysis/Homo_sapiens.EAS.Omega.normalize.pdf} 
\end{minipage}
\begin{minipage}{0.49\linewidth} 
\includegraphics[width=\linewidth, page=1]{/home/thibault/Documents/SelCoeff/analysis/Homo_sapiens.EAS.Omega.polyDFE_C.pdf} 
\end{minipage}
\begin{minipage}{0.49\linewidth} 
\includegraphics[width=\linewidth, page=1]{/home/thibault/Documents/SelCoeff/analysis/Homo_sapiens.EAS.Omega.histogram.pdf} 
\end{minipage}
\\ 
\subsection{EUR}

\subsubsection*{Site-specific Mutation-Selection codon models.} 
\begin{minipage}{0.49\linewidth} 
\includegraphics[width=\linewidth, page=1]{/home/thibault/Documents/SelCoeff/analysis/Homo_sapiens.EUR.MutSel.pdf} 
\end{minipage}
\begin{minipage}{0.49\linewidth} 
\includegraphics[width=\linewidth, page=1]{/home/thibault/Documents/SelCoeff/analysis/Homo_sapiens.EUR.MutSel.normalize.pdf} 
\end{minipage}
\begin{minipage}{0.49\linewidth} 
\includegraphics[width=\linewidth, page=1]{/home/thibault/Documents/SelCoeff/analysis/Homo_sapiens.EUR.MutSel.polyDFE_C.pdf} 
\end{minipage}
\begin{minipage}{0.49\linewidth} 
\includegraphics[width=\linewidth, page=1]{/home/thibault/Documents/SelCoeff/analysis/Homo_sapiens.EUR.MutSel.histogram.pdf} 
\end{minipage}
\\ 
\subsubsection*{Site-specific codon models.} 
\begin{minipage}{0.49\linewidth} 
\includegraphics[width=\linewidth, page=1]{/home/thibault/Documents/SelCoeff/analysis/Homo_sapiens.EUR.Omega.pdf} 
\end{minipage}
\begin{minipage}{0.49\linewidth} 
\includegraphics[width=\linewidth, page=1]{/home/thibault/Documents/SelCoeff/analysis/Homo_sapiens.EUR.Omega.normalize.pdf} 
\end{minipage}
\begin{minipage}{0.49\linewidth} 
\includegraphics[width=\linewidth, page=1]{/home/thibault/Documents/SelCoeff/analysis/Homo_sapiens.EUR.Omega.polyDFE_C.pdf} 
\end{minipage}
\begin{minipage}{0.49\linewidth} 
\includegraphics[width=\linewidth, page=1]{/home/thibault/Documents/SelCoeff/analysis/Homo_sapiens.EUR.Omega.histogram.pdf} 
\end{minipage}
\\ 
\subsection{SAS}

\subsubsection*{Site-specific Mutation-Selection codon models.} 
\begin{minipage}{0.49\linewidth} 
\includegraphics[width=\linewidth, page=1]{/home/thibault/Documents/SelCoeff/analysis/Homo_sapiens.SAS.MutSel.pdf} 
\end{minipage}
\begin{minipage}{0.49\linewidth} 
\includegraphics[width=\linewidth, page=1]{/home/thibault/Documents/SelCoeff/analysis/Homo_sapiens.SAS.MutSel.normalize.pdf} 
\end{minipage}
\begin{minipage}{0.49\linewidth} 
\includegraphics[width=\linewidth, page=1]{/home/thibault/Documents/SelCoeff/analysis/Homo_sapiens.SAS.MutSel.polyDFE_C.pdf} 
\end{minipage}
\begin{minipage}{0.49\linewidth} 
\includegraphics[width=\linewidth, page=1]{/home/thibault/Documents/SelCoeff/analysis/Homo_sapiens.SAS.MutSel.histogram.pdf} 
\end{minipage}
\\ 
\subsubsection*{Site-specific codon models.} 
\begin{minipage}{0.49\linewidth} 
\includegraphics[width=\linewidth, page=1]{/home/thibault/Documents/SelCoeff/analysis/Homo_sapiens.SAS.Omega.pdf} 
\end{minipage}
\begin{minipage}{0.49\linewidth} 
\includegraphics[width=\linewidth, page=1]{/home/thibault/Documents/SelCoeff/analysis/Homo_sapiens.SAS.Omega.normalize.pdf} 
\end{minipage}
\begin{minipage}{0.49\linewidth} 
\includegraphics[width=\linewidth, page=1]{/home/thibault/Documents/SelCoeff/analysis/Homo_sapiens.SAS.Omega.polyDFE_C.pdf} 
\end{minipage}
\begin{minipage}{0.49\linewidth} 
\includegraphics[width=\linewidth, page=1]{/home/thibault/Documents/SelCoeff/analysis/Homo_sapiens.SAS.Omega.histogram.pdf} 
\end{minipage}
\\ 
\section{Ovis aries}

\subsection{IROA}

\subsubsection*{Site-specific Mutation-Selection codon models.} 
\begin{minipage}{0.49\linewidth} 
\includegraphics[width=\linewidth, page=1]{/home/thibault/Documents/SelCoeff/analysis/Ovis_aries.IROA.MutSel.pdf} 
\end{minipage}
\begin{minipage}{0.49\linewidth} 
\includegraphics[width=\linewidth, page=1]{/home/thibault/Documents/SelCoeff/analysis/Ovis_aries.IROA.MutSel.normalize.pdf} 
\end{minipage}
\begin{minipage}{0.49\linewidth} 
\includegraphics[width=\linewidth, page=1]{/home/thibault/Documents/SelCoeff/analysis/Ovis_aries.IROA.MutSel.polyDFE_C.pdf} 
\end{minipage}
\begin{minipage}{0.49\linewidth} 
\includegraphics[width=\linewidth, page=1]{/home/thibault/Documents/SelCoeff/analysis/Ovis_aries.IROA.MutSel.histogram.pdf} 
\end{minipage}
\\ 
\subsubsection*{Site-specific codon models.} 
\begin{minipage}{0.49\linewidth} 
\includegraphics[width=\linewidth, page=1]{/home/thibault/Documents/SelCoeff/analysis/Ovis_aries.IROA.Omega.pdf} 
\end{minipage}
\begin{minipage}{0.49\linewidth} 
\includegraphics[width=\linewidth, page=1]{/home/thibault/Documents/SelCoeff/analysis/Ovis_aries.IROA.Omega.normalize.pdf} 
\end{minipage}
\begin{minipage}{0.49\linewidth} 
\includegraphics[width=\linewidth, page=1]{/home/thibault/Documents/SelCoeff/analysis/Ovis_aries.IROA.Omega.polyDFE_C.pdf} 
\end{minipage}
\begin{minipage}{0.49\linewidth} 
\includegraphics[width=\linewidth, page=1]{/home/thibault/Documents/SelCoeff/analysis/Ovis_aries.IROA.Omega.histogram.pdf} 
\end{minipage}
\\ 
\subsection{IROO}

\subsubsection*{Site-specific Mutation-Selection codon models.} 
\begin{minipage}{0.49\linewidth} 
\includegraphics[width=\linewidth, page=1]{/home/thibault/Documents/SelCoeff/analysis/Ovis_aries.IROO.MutSel.pdf} 
\end{minipage}
\begin{minipage}{0.49\linewidth} 
\includegraphics[width=\linewidth, page=1]{/home/thibault/Documents/SelCoeff/analysis/Ovis_aries.IROO.MutSel.normalize.pdf} 
\end{minipage}
\begin{minipage}{0.49\linewidth} 
\includegraphics[width=\linewidth, page=1]{/home/thibault/Documents/SelCoeff/analysis/Ovis_aries.IROO.MutSel.polyDFE_C.pdf} 
\end{minipage}
\begin{minipage}{0.49\linewidth} 
\includegraphics[width=\linewidth, page=1]{/home/thibault/Documents/SelCoeff/analysis/Ovis_aries.IROO.MutSel.histogram.pdf} 
\end{minipage}
\\ 
\subsubsection*{Site-specific codon models.} 
\begin{minipage}{0.49\linewidth} 
\includegraphics[width=\linewidth, page=1]{/home/thibault/Documents/SelCoeff/analysis/Ovis_aries.IROO.Omega.pdf} 
\end{minipage}
\begin{minipage}{0.49\linewidth} 
\includegraphics[width=\linewidth, page=1]{/home/thibault/Documents/SelCoeff/analysis/Ovis_aries.IROO.Omega.normalize.pdf} 
\end{minipage}
\begin{minipage}{0.49\linewidth} 
\includegraphics[width=\linewidth, page=1]{/home/thibault/Documents/SelCoeff/analysis/Ovis_aries.IROO.Omega.polyDFE_C.pdf} 
\end{minipage}
\begin{minipage}{0.49\linewidth} 
\includegraphics[width=\linewidth, page=1]{/home/thibault/Documents/SelCoeff/analysis/Ovis_aries.IROO.Omega.histogram.pdf} 
\end{minipage}
\\ 
\subsection{IROV}

\subsubsection*{Site-specific Mutation-Selection codon models.} 
\begin{minipage}{0.49\linewidth} 
\includegraphics[width=\linewidth, page=1]{/home/thibault/Documents/SelCoeff/analysis/Ovis_aries.IROV.MutSel.pdf} 
\end{minipage}
\begin{minipage}{0.49\linewidth} 
\includegraphics[width=\linewidth, page=1]{/home/thibault/Documents/SelCoeff/analysis/Ovis_aries.IROV.MutSel.normalize.pdf} 
\end{minipage}
\begin{minipage}{0.49\linewidth} 
\includegraphics[width=\linewidth, page=1]{/home/thibault/Documents/SelCoeff/analysis/Ovis_aries.IROV.MutSel.polyDFE_C.pdf} 
\end{minipage}
\begin{minipage}{0.49\linewidth} 
\includegraphics[width=\linewidth, page=1]{/home/thibault/Documents/SelCoeff/analysis/Ovis_aries.IROV.MutSel.histogram.pdf} 
\end{minipage}
\\ 
\subsubsection*{Site-specific codon models.} 
\begin{minipage}{0.49\linewidth} 
\includegraphics[width=\linewidth, page=1]{/home/thibault/Documents/SelCoeff/analysis/Ovis_aries.IROV.Omega.pdf} 
\end{minipage}
\begin{minipage}{0.49\linewidth} 
\includegraphics[width=\linewidth, page=1]{/home/thibault/Documents/SelCoeff/analysis/Ovis_aries.IROV.Omega.normalize.pdf} 
\end{minipage}
\begin{minipage}{0.49\linewidth} 
\includegraphics[width=\linewidth, page=1]{/home/thibault/Documents/SelCoeff/analysis/Ovis_aries.IROV.Omega.polyDFE_C.pdf} 
\end{minipage}
\begin{minipage}{0.49\linewidth} 
\includegraphics[width=\linewidth, page=1]{/home/thibault/Documents/SelCoeff/analysis/Ovis_aries.IROV.Omega.histogram.pdf} 
\end{minipage}
\\ 
\subsection{ISGC}

\subsubsection*{Site-specific Mutation-Selection codon models.} 
\begin{minipage}{0.49\linewidth} 
\includegraphics[width=\linewidth, page=1]{/home/thibault/Documents/SelCoeff/analysis/Ovis_aries.ISGC.MutSel.pdf} 
\end{minipage}
\begin{minipage}{0.49\linewidth} 
\includegraphics[width=\linewidth, page=1]{/home/thibault/Documents/SelCoeff/analysis/Ovis_aries.ISGC.MutSel.normalize.pdf} 
\end{minipage}
\begin{minipage}{0.49\linewidth} 
\includegraphics[width=\linewidth, page=1]{/home/thibault/Documents/SelCoeff/analysis/Ovis_aries.ISGC.MutSel.polyDFE_C.pdf} 
\end{minipage}
\begin{minipage}{0.49\linewidth} 
\includegraphics[width=\linewidth, page=1]{/home/thibault/Documents/SelCoeff/analysis/Ovis_aries.ISGC.MutSel.histogram.pdf} 
\end{minipage}
\\ 
\subsubsection*{Site-specific codon models.} 
\begin{minipage}{0.49\linewidth} 
\includegraphics[width=\linewidth, page=1]{/home/thibault/Documents/SelCoeff/analysis/Ovis_aries.ISGC.Omega.pdf} 
\end{minipage}
\begin{minipage}{0.49\linewidth} 
\includegraphics[width=\linewidth, page=1]{/home/thibault/Documents/SelCoeff/analysis/Ovis_aries.ISGC.Omega.normalize.pdf} 
\end{minipage}
\begin{minipage}{0.49\linewidth} 
\includegraphics[width=\linewidth, page=1]{/home/thibault/Documents/SelCoeff/analysis/Ovis_aries.ISGC.Omega.polyDFE_C.pdf} 
\end{minipage}
\begin{minipage}{0.49\linewidth} 
\includegraphics[width=\linewidth, page=1]{/home/thibault/Documents/SelCoeff/analysis/Ovis_aries.ISGC.Omega.histogram.pdf} 
\end{minipage}
\\ 
\subsection{MOOA}

\subsubsection*{Site-specific Mutation-Selection codon models.} 
\begin{minipage}{0.49\linewidth} 
\includegraphics[width=\linewidth, page=1]{/home/thibault/Documents/SelCoeff/analysis/Ovis_aries.MOOA.MutSel.pdf} 
\end{minipage}
\begin{minipage}{0.49\linewidth} 
\includegraphics[width=\linewidth, page=1]{/home/thibault/Documents/SelCoeff/analysis/Ovis_aries.MOOA.MutSel.normalize.pdf} 
\end{minipage}
\begin{minipage}{0.49\linewidth} 
\includegraphics[width=\linewidth, page=1]{/home/thibault/Documents/SelCoeff/analysis/Ovis_aries.MOOA.MutSel.polyDFE_C.pdf} 
\end{minipage}
\begin{minipage}{0.49\linewidth} 
\includegraphics[width=\linewidth, page=1]{/home/thibault/Documents/SelCoeff/analysis/Ovis_aries.MOOA.MutSel.histogram.pdf} 
\end{minipage}
\\ 
\subsubsection*{Site-specific codon models.} 
\begin{minipage}{0.49\linewidth} 
\includegraphics[width=\linewidth, page=1]{/home/thibault/Documents/SelCoeff/analysis/Ovis_aries.MOOA.Omega.pdf} 
\end{minipage}
\begin{minipage}{0.49\linewidth} 
\includegraphics[width=\linewidth, page=1]{/home/thibault/Documents/SelCoeff/analysis/Ovis_aries.MOOA.Omega.normalize.pdf} 
\end{minipage}
\begin{minipage}{0.49\linewidth} 
\includegraphics[width=\linewidth, page=1]{/home/thibault/Documents/SelCoeff/analysis/Ovis_aries.MOOA.Omega.polyDFE_C.pdf} 
\end{minipage}
\begin{minipage}{0.49\linewidth} 
\includegraphics[width=\linewidth, page=1]{/home/thibault/Documents/SelCoeff/analysis/Ovis_aries.MOOA.Omega.histogram.pdf} 
\end{minipage}
\\