### 2023-04-03 manuscript/include-MutSel-non-adaptive.tex
Updated manuscript before NG corrections

\subsection{Bos taurus}

\subsubsection{Iran}

\begin{minipage}{0.49\linewidth}
    \includegraphics[width=\linewidth, page=1]{../data_processed/opportunities_bins3Mask0.9/DFE.Bos_taurus.IRBT.MutSel.pdf}
\end{minipage}
\begin{minipage}{0.49\linewidth}
    \includegraphics[width=\linewidth, page=1]{../data_processed/ancestral_bins3Mask0.9/Bos_taurus.IRBT.pdf}
\end{minipage}
\\
\begin{minipage}{0.49\linewidth}
    \includegraphics[width=\linewidth, page=1]{../experiments/3bins-mC/analysis/Bos_taurus.IRBT.MutSel-sfs.normalize.pdf}
\end{minipage}
\begin{minipage}{0.4\linewidth}
    \includegraphics[width=\linewidth, page=1]{../experiments/3bins-mC/analysis/Bos_taurus.IRBT.MutSel.polyDFE_C.pdf}
\end{minipage}
\begin{minipage}{0.09\linewidth}
    \includegraphics[width=\linewidth, page=1]{artworks/legend.polycat.top}
\end{minipage}
\\

\subsubsection{Uganda}

\begin{minipage}{0.49\linewidth}
    \includegraphics[width=\linewidth, page=1]{../data_processed/opportunities_bins3Mask0.9/DFE.Bos_taurus.UGBT.MutSel.pdf}
\end{minipage}
\begin{minipage}{0.49\linewidth}
    \includegraphics[width=\linewidth, page=1]{../data_processed/ancestral_bins3Mask0.9/Bos_taurus.UGBT.pdf}
\end{minipage}
\\
\begin{minipage}{0.49\linewidth}
    \includegraphics[width=\linewidth, page=1]{../experiments/3bins-mC/analysis/Bos_taurus.UGBT.MutSel-sfs.normalize.pdf}
\end{minipage}
\begin{minipage}{0.4\linewidth}
    \includegraphics[width=\linewidth, page=1]{../experiments/3bins-mC/analysis/Bos_taurus.UGBT.MutSel.polyDFE_C.pdf}
\end{minipage}
\begin{minipage}{0.09\linewidth}
    \includegraphics[width=\linewidth, page=1]{artworks/legend.polycat.top}
\end{minipage}
\\

\subsection{Capra}

\subsubsection{Australia}

\begin{minipage}{0.49\linewidth}
    \includegraphics[width=\linewidth, page=1]{../data_processed/opportunities_bins3Mask0.9/DFE.Capra_hircus.AUCH.MutSel.pdf}
\end{minipage}
\begin{minipage}{0.49\linewidth}
    \includegraphics[width=\linewidth, page=1]{../data_processed/ancestral_bins3Mask0.9/Capra_hircus.AUCH.pdf}
\end{minipage}
\\
\begin{minipage}{0.49\linewidth}
    \includegraphics[width=\linewidth, page=1]{../experiments/3bins-mC/analysis/Capra_hircus.AUCH.MutSel-sfs.normalize.pdf}
\end{minipage}
\begin{minipage}{0.4\linewidth}
    \includegraphics[width=\linewidth, page=1]{../experiments/3bins-mC/analysis/Capra_hircus.AUCH.MutSel.polyDFE_C.pdf}
\end{minipage}
\begin{minipage}{0.09\linewidth}
    \includegraphics[width=\linewidth, page=1]{artworks/legend.polycat.top}
\end{minipage}
\\

\subsubsection{France}

\begin{minipage}{0.49\linewidth}
    \includegraphics[width=\linewidth, page=1]{../data_processed/opportunities_bins3Mask0.9/DFE.Capra_hircus.FRCH.MutSel.pdf}
\end{minipage}
\begin{minipage}{0.49\linewidth}
    \includegraphics[width=\linewidth, page=1]{../data_processed/ancestral_bins3Mask0.9/Capra_hircus.FRCH.pdf}
\end{minipage}
\\
\begin{minipage}{0.49\linewidth}
    \includegraphics[width=\linewidth, page=1]{../experiments/3bins-mC/analysis/Capra_hircus.FRCH.MutSel-sfs.normalize.pdf}
\end{minipage}
\begin{minipage}{0.4\linewidth}
    \includegraphics[width=\linewidth, page=1]{../experiments/3bins-mC/analysis/Capra_hircus.FRCH.MutSel.polyDFE_C.pdf}
\end{minipage}
\begin{minipage}{0.09\linewidth}
    \includegraphics[width=\linewidth, page=1]{artworks/legend.polycat.top}
\end{minipage}
\\

\subsubsection{Iran (C. aegagrus)}

\begin{minipage}{0.49\linewidth}
    \includegraphics[width=\linewidth, page=1]{../data_processed/opportunities_bins3Mask0.9/DFE.Capra_hircus.IRCA.MutSel.pdf}
\end{minipage}
\begin{minipage}{0.49\linewidth}
    \includegraphics[width=\linewidth, page=1]{../data_processed/ancestral_bins3Mask0.9/Capra_hircus.IRCA.pdf}
\end{minipage}
\\
\begin{minipage}{0.49\linewidth}
    \includegraphics[width=\linewidth, page=1]{../experiments/3bins-mC/analysis/Capra_hircus.IRCA.MutSel-sfs.normalize.pdf}
\end{minipage}
\begin{minipage}{0.4\linewidth}
    \includegraphics[width=\linewidth, page=1]{../experiments/3bins-mC/analysis/Capra_hircus.IRCA.MutSel.polyDFE_C.pdf}
\end{minipage}
\begin{minipage}{0.09\linewidth}
    \includegraphics[width=\linewidth, page=1]{artworks/legend.polycat.top}
\end{minipage}
\\

\subsubsection{Iran}

\begin{minipage}{0.49\linewidth}
    \includegraphics[width=\linewidth, page=1]{../data_processed/opportunities_bins3Mask0.9/DFE.Capra_hircus.IRCH.MutSel.pdf}
\end{minipage}
\begin{minipage}{0.49\linewidth}
    \includegraphics[width=\linewidth, page=1]{../data_processed/ancestral_bins3Mask0.9/Capra_hircus.IRCH.pdf}
\end{minipage}
\\
\begin{minipage}{0.49\linewidth}
    \includegraphics[width=\linewidth, page=1]{../experiments/3bins-mC/analysis/Capra_hircus.IRCH.MutSel-sfs.normalize.pdf}
\end{minipage}
\begin{minipage}{0.4\linewidth}
    \includegraphics[width=\linewidth, page=1]{../experiments/3bins-mC/analysis/Capra_hircus.IRCH.MutSel.polyDFE_C.pdf}
\end{minipage}
\begin{minipage}{0.09\linewidth}
    \includegraphics[width=\linewidth, page=1]{artworks/legend.polycat.top}
\end{minipage}
\\

\subsubsection{Italy}

\begin{minipage}{0.49\linewidth}
    \includegraphics[width=\linewidth, page=1]{../data_processed/opportunities_bins3Mask0.9/DFE.Capra_hircus.ITCH.MutSel.pdf}
\end{minipage}
\begin{minipage}{0.49\linewidth}
    \includegraphics[width=\linewidth, page=1]{../data_processed/ancestral_bins3Mask0.9/Capra_hircus.ITCH.pdf}
\end{minipage}
\\
\begin{minipage}{0.49\linewidth}
    \includegraphics[width=\linewidth, page=1]{../experiments/3bins-mC/analysis/Capra_hircus.ITCH.MutSel-sfs.normalize.pdf}
\end{minipage}
\begin{minipage}{0.4\linewidth}
    \includegraphics[width=\linewidth, page=1]{../experiments/3bins-mC/analysis/Capra_hircus.ITCH.MutSel.polyDFE_C.pdf}
\end{minipage}
\begin{minipage}{0.09\linewidth}
    \includegraphics[width=\linewidth, page=1]{artworks/legend.polycat.top}
\end{minipage}
\\

\subsubsection{Morocco}

\begin{minipage}{0.49\linewidth}
    \includegraphics[width=\linewidth, page=1]{../data_processed/opportunities_bins3Mask0.9/DFE.Capra_hircus.MOCH.MutSel.pdf}
\end{minipage}
\begin{minipage}{0.49\linewidth}
    \includegraphics[width=\linewidth, page=1]{../data_processed/ancestral_bins3Mask0.9/Capra_hircus.MOCH.pdf}
\end{minipage}
\\
\begin{minipage}{0.49\linewidth}
    \includegraphics[width=\linewidth, page=1]{../experiments/3bins-mC/analysis/Capra_hircus.MOCH.MutSel-sfs.normalize.pdf}
\end{minipage}
\begin{minipage}{0.4\linewidth}
    \includegraphics[width=\linewidth, page=1]{../experiments/3bins-mC/analysis/Capra_hircus.MOCH.MutSel.polyDFE_C.pdf}
\end{minipage}
\begin{minipage}{0.09\linewidth}
    \includegraphics[width=\linewidth, page=1]{artworks/legend.polycat.top}
\end{minipage}
\\

\subsection{Chlorocebus sabaeus}

\subsubsection{Barbados}

\begin{minipage}{0.49\linewidth}
    \includegraphics[width=\linewidth, page=1]{../data_processed/opportunities_bins3Mask0.9/DFE.Chlorocebus_sabaeus.Barbados.MutSel.pdf}
\end{minipage}
\begin{minipage}{0.49\linewidth}
    \includegraphics[width=\linewidth, page=1]{../data_processed/ancestral_bins3Mask0.9/Chlorocebus_sabaeus.Barbados.pdf}
\end{minipage}
\\
\begin{minipage}{0.49\linewidth}
    \includegraphics[width=\linewidth, page=1]{../experiments/3bins-mC/analysis/Chlorocebus_sabaeus.Barbados.MutSel-sfs.normalize.pdf}
\end{minipage}
\begin{minipage}{0.4\linewidth}
    \includegraphics[width=\linewidth, page=1]{../experiments/3bins-mC/analysis/Chlorocebus_sabaeus.Barbados.MutSel.polyDFE_C.pdf}
\end{minipage}
\begin{minipage}{0.09\linewidth}
    \includegraphics[width=\linewidth, page=1]{artworks/legend.polycat.top}
\end{minipage}
\\

\subsubsection{Central Afr. Rep.}

\begin{minipage}{0.49\linewidth}
    \includegraphics[width=\linewidth, page=1]{../data_processed/opportunities_bins3Mask0.9/DFE.Chlorocebus_sabaeus.Central_African_Republic.MutSel.pdf}
\end{minipage}
\begin{minipage}{0.49\linewidth}
    \includegraphics[width=\linewidth, page=1]{../data_processed/ancestral_bins3Mask0.9/Chlorocebus_sabaeus.Central_African_Republic.pdf}
\end{minipage}
\\
\begin{minipage}{0.49\linewidth}
    \includegraphics[width=\linewidth, page=1]{../experiments/3bins-mC/analysis/Chlorocebus_sabaeus.Central_African_Republic.MutSel-sfs.normalize.pdf}
\end{minipage}
\begin{minipage}{0.4\linewidth}
    \includegraphics[width=\linewidth, page=1]{../experiments/3bins-mC/analysis/Chlorocebus_sabaeus.Central_African_Republic.MutSel.polyDFE_C.pdf}
\end{minipage}
\begin{minipage}{0.09\linewidth}
    \includegraphics[width=\linewidth, page=1]{artworks/legend.polycat.top}
\end{minipage}
\\

\subsubsection{Ethiopia}

\begin{minipage}{0.49\linewidth}
    \includegraphics[width=\linewidth, page=1]{../data_processed/opportunities_bins3Mask0.9/DFE.Chlorocebus_sabaeus.Ethiopia.MutSel.pdf}
\end{minipage}
\begin{minipage}{0.49\linewidth}
    \includegraphics[width=\linewidth, page=1]{../data_processed/ancestral_bins3Mask0.9/Chlorocebus_sabaeus.Ethiopia.pdf}
\end{minipage}
\\
\begin{minipage}{0.49\linewidth}
    \includegraphics[width=\linewidth, page=1]{../experiments/3bins-mC/analysis/Chlorocebus_sabaeus.Ethiopia.MutSel-sfs.normalize.pdf}
\end{minipage}
\begin{minipage}{0.4\linewidth}
    \includegraphics[width=\linewidth, page=1]{../experiments/3bins-mC/analysis/Chlorocebus_sabaeus.Ethiopia.MutSel.polyDFE_C.pdf}
\end{minipage}
\begin{minipage}{0.09\linewidth}
    \includegraphics[width=\linewidth, page=1]{artworks/legend.polycat.top}
\end{minipage}
\\

\subsubsection{Gambia}

\begin{minipage}{0.49\linewidth}
    \includegraphics[width=\linewidth, page=1]{../data_processed/opportunities_bins3Mask0.9/DFE.Chlorocebus_sabaeus.Gambia.MutSel.pdf}
\end{minipage}
\begin{minipage}{0.49\linewidth}
    \includegraphics[width=\linewidth, page=1]{../data_processed/ancestral_bins3Mask0.9/Chlorocebus_sabaeus.Gambia.pdf}
\end{minipage}
\\
\begin{minipage}{0.49\linewidth}
    \includegraphics[width=\linewidth, page=1]{../experiments/3bins-mC/analysis/Chlorocebus_sabaeus.Gambia.MutSel-sfs.normalize.pdf}
\end{minipage}
\begin{minipage}{0.4\linewidth}
    \includegraphics[width=\linewidth, page=1]{../experiments/3bins-mC/analysis/Chlorocebus_sabaeus.Gambia.MutSel.polyDFE_C.pdf}
\end{minipage}
\begin{minipage}{0.09\linewidth}
    \includegraphics[width=\linewidth, page=1]{artworks/legend.polycat.top}
\end{minipage}
\\

\subsubsection{Kenya}

\begin{minipage}{0.49\linewidth}
    \includegraphics[width=\linewidth, page=1]{../data_processed/opportunities_bins3Mask0.9/DFE.Chlorocebus_sabaeus.Kenya.MutSel.pdf}
\end{minipage}
\begin{minipage}{0.49\linewidth}
    \includegraphics[width=\linewidth, page=1]{../data_processed/ancestral_bins3Mask0.9/Chlorocebus_sabaeus.Kenya.pdf}
\end{minipage}
\\
\begin{minipage}{0.49\linewidth}
    \includegraphics[width=\linewidth, page=1]{../experiments/3bins-mC/analysis/Chlorocebus_sabaeus.Kenya.MutSel-sfs.normalize.pdf}
\end{minipage}
\begin{minipage}{0.4\linewidth}
    \includegraphics[width=\linewidth, page=1]{../experiments/3bins-mC/analysis/Chlorocebus_sabaeus.Kenya.MutSel.polyDFE_C.pdf}
\end{minipage}
\begin{minipage}{0.09\linewidth}
    \includegraphics[width=\linewidth, page=1]{artworks/legend.polycat.top}
\end{minipage}
\\

\subsubsection{Nevis}

\begin{minipage}{0.49\linewidth}
    \includegraphics[width=\linewidth, page=1]{../data_processed/opportunities_bins3Mask0.9/DFE.Chlorocebus_sabaeus.Nevis.MutSel.pdf}
\end{minipage}
\begin{minipage}{0.49\linewidth}
    \includegraphics[width=\linewidth, page=1]{../data_processed/ancestral_bins3Mask0.9/Chlorocebus_sabaeus.Nevis.pdf}
\end{minipage}
\\
\begin{minipage}{0.49\linewidth}
    \includegraphics[width=\linewidth, page=1]{../experiments/3bins-mC/analysis/Chlorocebus_sabaeus.Nevis.MutSel-sfs.normalize.pdf}
\end{minipage}
\begin{minipage}{0.4\linewidth}
    \includegraphics[width=\linewidth, page=1]{../experiments/3bins-mC/analysis/Chlorocebus_sabaeus.Nevis.MutSel.polyDFE_C.pdf}
\end{minipage}
\begin{minipage}{0.09\linewidth}
    \includegraphics[width=\linewidth, page=1]{artworks/legend.polycat.top}
\end{minipage}
\\

\subsubsection{Saint Kitts}

\begin{minipage}{0.49\linewidth}
    \includegraphics[width=\linewidth, page=1]{../data_processed/opportunities_bins3Mask0.9/DFE.Chlorocebus_sabaeus.Saint_Kitts.MutSel.pdf}
\end{minipage}
\begin{minipage}{0.49\linewidth}
    \includegraphics[width=\linewidth, page=1]{../data_processed/ancestral_bins3Mask0.9/Chlorocebus_sabaeus.Saint_Kitts.pdf}
\end{minipage}
\\
\begin{minipage}{0.49\linewidth}
    \includegraphics[width=\linewidth, page=1]{../experiments/3bins-mC/analysis/Chlorocebus_sabaeus.Saint_Kitts.MutSel-sfs.normalize.pdf}
\end{minipage}
\begin{minipage}{0.4\linewidth}
    \includegraphics[width=\linewidth, page=1]{../experiments/3bins-mC/analysis/Chlorocebus_sabaeus.Saint_Kitts.MutSel.polyDFE_C.pdf}
\end{minipage}
\begin{minipage}{0.09\linewidth}
    \includegraphics[width=\linewidth, page=1]{artworks/legend.polycat.top}
\end{minipage}
\\

\subsubsection{South Africa}

\begin{minipage}{0.49\linewidth}
    \includegraphics[width=\linewidth, page=1]{../data_processed/opportunities_bins3Mask0.9/DFE.Chlorocebus_sabaeus.South_Africa.MutSel.pdf}
\end{minipage}
\begin{minipage}{0.49\linewidth}
    \includegraphics[width=\linewidth, page=1]{../data_processed/ancestral_bins3Mask0.9/Chlorocebus_sabaeus.South_Africa.pdf}
\end{minipage}
\\
\begin{minipage}{0.49\linewidth}
    \includegraphics[width=\linewidth, page=1]{../experiments/3bins-mC/analysis/Chlorocebus_sabaeus.South_Africa.MutSel-sfs.normalize.pdf}
\end{minipage}
\begin{minipage}{0.4\linewidth}
    \includegraphics[width=\linewidth, page=1]{../experiments/3bins-mC/analysis/Chlorocebus_sabaeus.South_Africa.MutSel.polyDFE_C.pdf}
\end{minipage}
\begin{minipage}{0.09\linewidth}
    \includegraphics[width=\linewidth, page=1]{artworks/legend.polycat.top}
\end{minipage}
\\

\subsubsection{Zambia}

\begin{minipage}{0.49\linewidth}
    \includegraphics[width=\linewidth, page=1]{../data_processed/opportunities_bins3Mask0.9/DFE.Chlorocebus_sabaeus.Zambia.MutSel.pdf}
\end{minipage}
\begin{minipage}{0.49\linewidth}
    \includegraphics[width=\linewidth, page=1]{../data_processed/ancestral_bins3Mask0.9/Chlorocebus_sabaeus.Zambia.pdf}
\end{minipage}
\\
\begin{minipage}{0.49\linewidth}
    \includegraphics[width=\linewidth, page=1]{../experiments/3bins-mC/analysis/Chlorocebus_sabaeus.Zambia.MutSel-sfs.normalize.pdf}
\end{minipage}
\begin{minipage}{0.4\linewidth}
    \includegraphics[width=\linewidth, page=1]{../experiments/3bins-mC/analysis/Chlorocebus_sabaeus.Zambia.MutSel.polyDFE_C.pdf}
\end{minipage}
\begin{minipage}{0.09\linewidth}
    \includegraphics[width=\linewidth, page=1]{artworks/legend.polycat.top}
\end{minipage}
\\

\subsection{Equus caballus}

\begin{minipage}{0.49\linewidth}
    \includegraphics[width=\linewidth, page=1]{../data_processed/opportunities_bins3Mask0.9/DFE.Equus_caballus.up.MutSel.pdf}
\end{minipage}
\begin{minipage}{0.49\linewidth}
    \includegraphics[width=\linewidth, page=1]{../data_processed/ancestral_bins3Mask0.9/Equus_caballus.up.pdf}
\end{minipage}
\\
\begin{minipage}{0.49\linewidth}
    \includegraphics[width=\linewidth, page=1]{../experiments/3bins-mC/analysis/Equus_caballus.up.MutSel-sfs.normalize.pdf}
\end{minipage}
\begin{minipage}{0.4\linewidth}
    \includegraphics[width=\linewidth, page=1]{../experiments/3bins-mC/analysis/Equus_caballus.up.MutSel.polyDFE_C.pdf}
\end{minipage}
\begin{minipage}{0.09\linewidth}
    \includegraphics[width=\linewidth, page=1]{artworks/legend.polycat.top}
\end{minipage}
\\

\subsection{Homo sapiens}

\subsubsection{African}

\begin{minipage}{0.49\linewidth}
    \includegraphics[width=\linewidth, page=1]{../data_processed/opportunities_bins3Mask0.9/DFE.Homo_sapiens.AFR.MutSel.pdf}
\end{minipage}
\begin{minipage}{0.49\linewidth}
    \includegraphics[width=\linewidth, page=1]{../data_processed/ancestral_bins3Mask0.9/Homo_sapiens.AFR.pdf}
\end{minipage}
\\
\begin{minipage}{0.49\linewidth}
    \includegraphics[width=\linewidth, page=1]{../experiments/3bins-mC/analysis/Homo_sapiens.AFR.MutSel-sfs.normalize.pdf}
\end{minipage}
\begin{minipage}{0.4\linewidth}
    \includegraphics[width=\linewidth, page=1]{../experiments/3bins-mC/analysis/Homo_sapiens.AFR.MutSel.polyDFE_C.pdf}
\end{minipage}
\begin{minipage}{0.09\linewidth}
    \includegraphics[width=\linewidth, page=1]{artworks/legend.polycat.top}
\end{minipage}
\\

\subsubsection{Admixed American}

\begin{minipage}{0.49\linewidth}
    \includegraphics[width=\linewidth, page=1]{../data_processed/opportunities_bins3Mask0.9/DFE.Homo_sapiens.AMR.MutSel.pdf}
\end{minipage}
\begin{minipage}{0.49\linewidth}
    \includegraphics[width=\linewidth, page=1]{../data_processed/ancestral_bins3Mask0.9/Homo_sapiens.AMR.pdf}
\end{minipage}
\\
\begin{minipage}{0.49\linewidth}
    \includegraphics[width=\linewidth, page=1]{../experiments/3bins-mC/analysis/Homo_sapiens.AMR.MutSel-sfs.normalize.pdf}
\end{minipage}
\begin{minipage}{0.4\linewidth}
    \includegraphics[width=\linewidth, page=1]{../experiments/3bins-mC/analysis/Homo_sapiens.AMR.MutSel.polyDFE_C.pdf}
\end{minipage}
\begin{minipage}{0.09\linewidth}
    \includegraphics[width=\linewidth, page=1]{artworks/legend.polycat.top}
\end{minipage}
\\

\subsubsection{East Asian}

\begin{minipage}{0.49\linewidth}
    \includegraphics[width=\linewidth, page=1]{../data_processed/opportunities_bins3Mask0.9/DFE.Homo_sapiens.EAS.MutSel.pdf}
\end{minipage}
\begin{minipage}{0.49\linewidth}
    \includegraphics[width=\linewidth, page=1]{../data_processed/ancestral_bins3Mask0.9/Homo_sapiens.EAS.pdf}
\end{minipage}
\\
\begin{minipage}{0.49\linewidth}
    \includegraphics[width=\linewidth, page=1]{../experiments/3bins-mC/analysis/Homo_sapiens.EAS.MutSel-sfs.normalize.pdf}
\end{minipage}
\begin{minipage}{0.4\linewidth}
    \includegraphics[width=\linewidth, page=1]{../experiments/3bins-mC/analysis/Homo_sapiens.EAS.MutSel.polyDFE_C.pdf}
\end{minipage}
\begin{minipage}{0.09\linewidth}
    \includegraphics[width=\linewidth, page=1]{artworks/legend.polycat.top}
\end{minipage}
\\

\subsubsection{European}

\begin{minipage}{0.49\linewidth}
    \includegraphics[width=\linewidth, page=1]{../data_processed/opportunities_bins3Mask0.9/DFE.Homo_sapiens.EUR.MutSel.pdf}
\end{minipage}
\begin{minipage}{0.49\linewidth}
    \includegraphics[width=\linewidth, page=1]{../data_processed/ancestral_bins3Mask0.9/Homo_sapiens.EUR.pdf}
\end{minipage}
\\
\begin{minipage}{0.49\linewidth}
    \includegraphics[width=\linewidth, page=1]{../experiments/3bins-mC/analysis/Homo_sapiens.EUR.MutSel-sfs.normalize.pdf}
\end{minipage}
\begin{minipage}{0.4\linewidth}
    \includegraphics[width=\linewidth, page=1]{../experiments/3bins-mC/analysis/Homo_sapiens.EUR.MutSel.polyDFE_C.pdf}
\end{minipage}
\begin{minipage}{0.09\linewidth}
    \includegraphics[width=\linewidth, page=1]{artworks/legend.polycat.top}
\end{minipage}
\\

\subsubsection{South Asian}

\begin{minipage}{0.49\linewidth}
    \includegraphics[width=\linewidth, page=1]{../data_processed/opportunities_bins3Mask0.9/DFE.Homo_sapiens.SAS.MutSel.pdf}
\end{minipage}
\begin{minipage}{0.49\linewidth}
    \includegraphics[width=\linewidth, page=1]{../data_processed/ancestral_bins3Mask0.9/Homo_sapiens.SAS.pdf}
\end{minipage}
\\
\begin{minipage}{0.49\linewidth}
    \includegraphics[width=\linewidth, page=1]{../experiments/3bins-mC/analysis/Homo_sapiens.SAS.MutSel-sfs.normalize.pdf}
\end{minipage}
\begin{minipage}{0.4\linewidth}
    \includegraphics[width=\linewidth, page=1]{../experiments/3bins-mC/analysis/Homo_sapiens.SAS.MutSel.polyDFE_C.pdf}
\end{minipage}
\begin{minipage}{0.09\linewidth}
    \includegraphics[width=\linewidth, page=1]{artworks/legend.polycat.top}
\end{minipage}
\\

\subsection{Ovis}

\subsubsection{Iran}

\begin{minipage}{0.49\linewidth}
    \includegraphics[width=\linewidth, page=1]{../data_processed/opportunities_bins3Mask0.9/DFE.Ovis_aries.IROA.MutSel.pdf}
\end{minipage}
\begin{minipage}{0.49\linewidth}
    \includegraphics[width=\linewidth, page=1]{../data_processed/ancestral_bins3Mask0.9/Ovis_aries.IROA.pdf}
\end{minipage}
\\
\begin{minipage}{0.49\linewidth}
    \includegraphics[width=\linewidth, page=1]{../experiments/3bins-mC/analysis/Ovis_aries.IROA.MutSel-sfs.normalize.pdf}
\end{minipage}
\begin{minipage}{0.4\linewidth}
    \includegraphics[width=\linewidth, page=1]{../experiments/3bins-mC/analysis/Ovis_aries.IROA.MutSel.polyDFE_C.pdf}
\end{minipage}
\begin{minipage}{0.09\linewidth}
    \includegraphics[width=\linewidth, page=1]{artworks/legend.polycat.top}
\end{minipage}
\\

\subsubsection{Iran (O. orientalis)}

\begin{minipage}{0.49\linewidth}
    \includegraphics[width=\linewidth, page=1]{../data_processed/opportunities_bins3Mask0.9/DFE.Ovis_aries.IROO.MutSel.pdf}
\end{minipage}
\begin{minipage}{0.49\linewidth}
    \includegraphics[width=\linewidth, page=1]{../data_processed/ancestral_bins3Mask0.9/Ovis_aries.IROO.pdf}
\end{minipage}
\\
\begin{minipage}{0.49\linewidth}
    \includegraphics[width=\linewidth, page=1]{../experiments/3bins-mC/analysis/Ovis_aries.IROO.MutSel-sfs.normalize.pdf}
\end{minipage}
\begin{minipage}{0.4\linewidth}
    \includegraphics[width=\linewidth, page=1]{../experiments/3bins-mC/analysis/Ovis_aries.IROO.MutSel.polyDFE_C.pdf}
\end{minipage}
\begin{minipage}{0.09\linewidth}
    \includegraphics[width=\linewidth, page=1]{artworks/legend.polycat.top}
\end{minipage}
\\

\subsubsection{Iran (O. vignei)}

\begin{minipage}{0.49\linewidth}
    \includegraphics[width=\linewidth, page=1]{../data_processed/opportunities_bins3Mask0.9/DFE.Ovis_aries.IROV.MutSel.pdf}
\end{minipage}
\begin{minipage}{0.49\linewidth}
    \includegraphics[width=\linewidth, page=1]{../data_processed/ancestral_bins3Mask0.9/Ovis_aries.IROV.pdf}
\end{minipage}
\\
\begin{minipage}{0.49\linewidth}
    \includegraphics[width=\linewidth, page=1]{../experiments/3bins-mC/analysis/Ovis_aries.IROV.MutSel-sfs.normalize.pdf}
\end{minipage}
\begin{minipage}{0.4\linewidth}
    \includegraphics[width=\linewidth, page=1]{../experiments/3bins-mC/analysis/Ovis_aries.IROV.MutSel.polyDFE_C.pdf}
\end{minipage}
\begin{minipage}{0.09\linewidth}
    \includegraphics[width=\linewidth, page=1]{artworks/legend.polycat.top}
\end{minipage}
\\

\subsubsection{Various}

\begin{minipage}{0.49\linewidth}
    \includegraphics[width=\linewidth, page=1]{../data_processed/opportunities_bins3Mask0.9/DFE.Ovis_aries.ISGC.MutSel.pdf}
\end{minipage}
\begin{minipage}{0.49\linewidth}
    \includegraphics[width=\linewidth, page=1]{../data_processed/ancestral_bins3Mask0.9/Ovis_aries.ISGC.pdf}
\end{minipage}
\\
\begin{minipage}{0.49\linewidth}
    \includegraphics[width=\linewidth, page=1]{../experiments/3bins-mC/analysis/Ovis_aries.ISGC.MutSel-sfs.normalize.pdf}
\end{minipage}
\begin{minipage}{0.4\linewidth}
    \includegraphics[width=\linewidth, page=1]{../experiments/3bins-mC/analysis/Ovis_aries.ISGC.MutSel.polyDFE_C.pdf}
\end{minipage}
\begin{minipage}{0.09\linewidth}
    \includegraphics[width=\linewidth, page=1]{artworks/legend.polycat.top}
\end{minipage}
\\

\subsubsection{Morocco}

\begin{minipage}{0.49\linewidth}
    \includegraphics[width=\linewidth, page=1]{../data_processed/opportunities_bins3Mask0.9/DFE.Ovis_aries.MOOA.MutSel.pdf}
\end{minipage}
\begin{minipage}{0.49\linewidth}
    \includegraphics[width=\linewidth, page=1]{../data_processed/ancestral_bins3Mask0.9/Ovis_aries.MOOA.pdf}
\end{minipage}
\\
\begin{minipage}{0.49\linewidth}
    \includegraphics[width=\linewidth, page=1]{../experiments/3bins-mC/analysis/Ovis_aries.MOOA.MutSel-sfs.normalize.pdf}
\end{minipage}
\begin{minipage}{0.49\linewidth}
    \includegraphics[width=\linewidth, page=1]{../experiments/3bins-mC/analysis/Ovis_aries.MOOA.MutSel.polyDFE_C.pdf}
\end{minipage}
\\