### 2023-05-03 manuscript/supp-mat.tex

%TC:ignore
\documentclass{article}
\usepackage{caption}
\usepackage{xcolor, colortbl}
\definecolor{LIGHTGREY}{gray}{0.9}
\definecolor{BLUELINK}{HTML}{0645AD}
\definecolor{DARKBLUELINK}{HTML}{0B0080}
\PassOptionsToPackage{hyphens}{url}
\usepackage[colorlinks=false]{hyperref}
% for linking between references, figures, TOC, etc in the pdf document
\hypersetup{colorlinks,
    linkcolor=DARKBLUELINK,
    anchorcolor=DARKBLUELINK,
    citecolor=DARKBLUELINK,
    filecolor=DARKBLUELINK,
    menucolor=DARKBLUELINK,
    urlcolor=BLUELINK
} % Color citation links in purple
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{naturalnames}{hyperref}

\usepackage[backend=biber,eprint=false,isbn=false,url=false,intitle=true,style=nature,date=year]{biblatex}
\addbibresource{codon_models.bib}

\usepackage[margin=50pt]{geometry}
\usepackage{amssymb,amsfonts,amsmath,amsthm,mathtools}
\usepackage{lmodern}
\usepackage{bm,bbold}
\usepackage{verbatim}
\usepackage{float}
\usepackage{listings, enumerate, enumitem}
\usepackage[export]{adjustbox}
\usepackage{tabu}
\usepackage{longtable}
\tabulinesep=0.6mm
\newcommand\cellwidth{\TX@col@width}
\usepackage{hhline}
\setlength{\arrayrulewidth}{1.2pt}
\usepackage{multicol,multirow,array}
\usepackage{etoolbox}
\AtBeginEnvironment{tabu}{\footnotesize}
\usepackage{booktabs}
\usepackage{makecell}
\usepackage{orcidlink}
\usepackage{graphicx}
\pdfinclusioncopyfonts=1

\renewcommand{\thetable}{S\arabic{table}}
\renewcommand{\thefigure}{S\arabic{figure}}

\newcommand{\UniDimArray}[1]{\bm{#1}}
\newcommand{\der}{\text{d}}
\newcommand{\e}{\text{e}}
\newcommand{\Ne}{N_{\text{e}}}
\newcommand{\proba}{\mathbb{P}}
\newcommand{\pfix}{\proba_{\text{fix}}}
\newcommand{\dn}{d_N}
\newcommand{\ds}{d_S}
\newcommand{\dnds}{\dn / \ds}
\newcommand{\Sphy}{S_{0}}
\newcommand{\SphyMean}{\overline{\Sphy}}
\newcommand{\SphyDel}{\mathcal{D}_0}
\newcommand{\SphyNeu}{\mathcal{N}_0}
\newcommand{\SphyBen}{\mathcal{B}_0}
\newcommand{\Sphyclass}{x}
\newcommand{\SphyclassAlt}{y}
\newcommand{\given}{\mid}
\newcommand{\Spop}{S}
\newcommand{\SpopDel}{\mathcal{D}}
\newcommand{\SpopNeu}{\mathcal{N}}
\newcommand{\SpopBen}{\mathcal{B}}
\newcommand{\ProbaPopDel}{\proba [ \SpopDel]}
\newcommand{\ProbaPopNeu}{\proba [ \SpopNeu ]}
\newcommand{\ProbaPopBen}{\proba [ \SpopBen ]}
\newcommand{\AdvMean}{\beta_b}
\newcommand{\DelMean}{\beta_d}
\newcommand{\thetaSyn}{\theta_{\text{S}}}

\title{\textbf{Mammalian protein-coding genes exhibit widespread beneficial mutations that are not adaptive}}

\author{
    \large
    \textbf{T. {Latrille}$^{1\dag}$\orcidlink{0000-0002-9643-4668}, J. {Joseph}$^{2\dag}$\orcidlink{0009-0002-1312-9930}, D.~A. {Hartasánchez}$^{1}$\orcidlink{0000-0003-2596-6883}, N. {Salamin}$^{1}$\orcidlink{0000-0002-3963-4954}}\\
    \scriptsize $^{1}$Department of Computational Biology, Université de Lausanne, Lausanne, Switzerland\\
    \scriptsize $^{2}$Laboratoire de Biométrie et Biologie Evolutive, UMR5558, Université Lyon 1, Villeurbanne, France \\
    \scriptsize $^{\dag}$These authors contributed equally to this work\\
    \normalsize \texttt{\href{mailto:thibault.latrille@ens-lyon.org}{thibault.latrille@ens-lyon.org}} \\
}

\date{}

\begin{document}
    \maketitle
    \part*{Supplementary materials}
    \tableofcontents
    \clearpage

    \section{Beneficial mutations in the terminal lineages and populations}\label{sec:beneficial-mutations}

    \subsection{Mutation-selection codon models}
    \label{sec:site-specific-mutation-selection-codon-models}
    In \textit{BayesCode} (\url{https://github.com/ThibaultLatrille/bayescode}), we ran the mutation-selection codon models \textit{mutselomega} for 2000 points of MCMC with the options:
    \begin{scriptsize}
        \begin{verbatim}
 mutselomega ---omegashift 0.0 --ncat 30 -a my_alignment.phy -t my_tree.newick -u 2000 my_genename
        \end{verbatim}
    \end{scriptsize}
    The collection of site-specific fitness profiles ($\UniDimArray{F^{(i)}}, \forall i$) are then obtained by running \textit{readmutselomega}, reading 1000 points of MCMC (first 1000 are considered as burn-in) with the options:
    \begin{scriptsize}
        \begin{verbatim}
 readmutselomega --every 1 --until 2000 --burnin 1000 --ss my_genename
        \end{verbatim}
    \end{scriptsize}
    The gene-specific $4 \times 4$ nucleotide mutation rate matrix ($\UniDimArray{\mu}$) is also obtained by running \textit{readmutselomega}, reading 1000 points of MCMC (first 1000 are considered as burn-in) with the options:
    \begin{scriptsize}
        \begin{verbatim}
 readmutselomega --every 1 --until 2000 --burnin 1000 --nuc my_genename
        \end{verbatim}
    \end{scriptsize}

    \subsection{Summary tables}\label{subsec:summary-table-mutsel}

    \subsubsection{$\dnds$ for deleterious ($\SphyDel$), nearly-neutral ($\SphyNeu$) and beneficial back-mutations ($\SphyBen$)}
    \begin{center}
        \captionof{table}{$\dnds$ for deleterious ($\SphyDel$), nearly-neutral ($\SphyNeu$) and beneficial back-mutations ($\SphyBen$).}
        \scriptsize
        \begin{longtable*}{|l|l|r|r|r|r|r|r|}
            \toprule
            Population &             Species & Watterson's $\thetaSyn$  & $\proba_{div}[ \SphyBen ]$ & $\dnds $ & $\dn ( \SphyDel ) / \ds$ & $\dn ( \SphyNeu ) / \ds$ & $\dn ( \SphyBen ) / \ds$ \\
            \midrule
            \endhead
            \midrule
            \multicolumn{8}{r}{{Continued on next page}} \\
            \midrule
            \endfoot

            \bottomrule
            \endlastfoot
            \rowcolor{LIGHTGREY} Equus c. &      Equus caballus &               $ 0.002$ &                    $ 0.118$ &  $ 0.147$ &            $ 0.075$ &                $ 0.823$ &           $ 1.203$ \\
            Iran &          Bos taurus &               $ 0.002$ &                    $ 0.123$ &  $ 0.131$ &            $ 0.074$ &                $ 0.627$ &           $ 1.206$ \\
            Uganda &          Bos taurus &               $ 0.005$ &                    $ 0.125$ &  $ 0.134$ &            $ 0.075$ &                $ 0.637$ &           $ 1.252$ \\
            \rowcolor{LIGHTGREY} Australia &        Capra hircus &               $ 0.003$ &                    $ 0.121$ &  $ 0.125$ &            $ 0.068$ &                $ 0.629$ &           $ 1.099$ \\
            \rowcolor{LIGHTGREY} France &        Capra hircus &               $ 0.004$ &                    $ 0.120$ &  $ 0.126$ &            $ 0.068$ &                $ 0.633$ &           $ 1.102$ \\
            \rowcolor{LIGHTGREY} Iran (C. aegagrus) &        Capra hircus &               $ 0.004$ &                    $ 0.120$ &  $ 0.125$ &            $ 0.068$ &                $ 0.630$ &           $ 1.100$ \\
            \rowcolor{LIGHTGREY} Iran &        Capra hircus &               $ 0.004$ &                    $ 0.121$ &  $ 0.125$ &            $ 0.067$ &                $ 0.630$ &           $ 1.108$ \\
            \rowcolor{LIGHTGREY} Italy &        Capra hircus &               $ 0.004$ &                    $ 0.120$ &  $ 0.126$ &            $ 0.068$ &                $ 0.634$ &           $ 1.102$ \\
            \rowcolor{LIGHTGREY} Morocco &        Capra hircus &               $ 0.004$ &                    $ 0.122$ &  $ 0.125$ &            $ 0.067$ &                $ 0.631$ &           $ 1.112$ \\
            Iran &          Ovis aries &               $ 0.007$ &                    $ 0.109$ &  $ 0.146$ &            $ 0.091$ &                $ 0.616$ &           $ 1.148$ \\
            Iran (O. orientalis) &          Ovis aries &               $ 0.009$ &                    $ 0.108$ &  $ 0.148$ &            $ 0.093$ &                $ 0.617$ &           $ 1.145$ \\
            Iran (O. vignei) &          Ovis aries &               $ 0.007$ &                    $ 0.109$ &  $ 0.146$ &            $ 0.091$ &                $ 0.617$ &           $ 1.134$ \\
            Various &          Ovis aries &               $ 0.008$ &                    $ 0.107$ &  $ 0.148$ &            $ 0.093$ &                $ 0.617$ &           $ 1.134$ \\
            Morocco &          Ovis aries &               $ 0.008$ &                    $ 0.108$ &  $ 0.148$ &            $ 0.093$ &                $ 0.619$ &           $ 1.149$ \\
            \rowcolor{LIGHTGREY} Barbados & Chlorocebus sabaeus &               $ 0.004$ &                    $ 0.122$ &  $ 0.133$ &            $ 0.071$ &                $ 0.696$ &           $ 1.397$ \\
            \rowcolor{LIGHTGREY} Central Afr. Rep. & Chlorocebus sabaeus &               $ 0.006$ &                    $ 0.124$ &  $ 0.134$ &            $ 0.071$ &                $ 0.701$ &           $ 1.431$ \\
            \rowcolor{LIGHTGREY} Ethiopia & Chlorocebus sabaeus &               $ 0.005$ &                    $ 0.124$ &  $ 0.134$ &            $ 0.071$ &                $ 0.699$ &           $ 1.422$ \\
            \rowcolor{LIGHTGREY} Gambia & Chlorocebus sabaeus &               $ 0.005$ &                    $ 0.123$ &  $ 0.134$ &            $ 0.071$ &                $ 0.703$ &           $ 1.411$ \\
            \rowcolor{LIGHTGREY} Kenya & Chlorocebus sabaeus &               $ 0.005$ &                    $ 0.123$ &  $ 0.134$ &            $ 0.071$ &                $ 0.699$ &           $ 1.419$ \\
            \rowcolor{LIGHTGREY} Nevis & Chlorocebus sabaeus &               $ 0.003$ &                    $ 0.123$ &  $ 0.133$ &            $ 0.071$ &                $ 0.699$ &           $ 1.407$ \\
            \rowcolor{LIGHTGREY} South Africa  & Chlorocebus sabaeus &               $ 0.006$ &                    $ 0.125$ &  $ 0.134$ &            $ 0.070$ &                $ 0.704$ &           $ 1.447$ \\
            \rowcolor{LIGHTGREY} Saint Kitts & Chlorocebus sabaeus &               $ 0.004$ &                    $ 0.123$ &  $ 0.133$ &            $ 0.070$ &                $ 0.699$ &           $ 1.413$ \\
            \rowcolor{LIGHTGREY} Zambia & Chlorocebus sabaeus &               $ 0.006$ &                    $ 0.123$ &  $ 0.134$ &            $ 0.071$ &                $ 0.699$ &           $ 1.410$ \\
            African &        Homo sapiens &               $ 0.002$ &                    $ 0.099$ &  $ 0.192$ &            $ 0.118$ &                $ 0.865$ &           $ 1.609$ \\
            Admixed American &        Homo sapiens &               $ 0.002$ &                    $ 0.099$ &  $ 0.192$ &            $ 0.118$ &                $ 0.865$ &           $ 1.607$ \\
            East Asian &        Homo sapiens &               $ 0.002$ &                    $ 0.098$ &  $ 0.192$ &            $ 0.118$ &                $ 0.869$ &           $ 1.607$ \\
            European &        Homo sapiens &               $ 0.002$ &                    $ 0.098$ &  $ 0.192$ &            $ 0.118$ &                $ 0.865$ &           $ 1.598$ \\
            South Asian &        Homo sapiens &               $ 0.002$ &                    $ 0.099$ &  $ 0.192$ &            $ 0.118$ &                $ 0.867$ &           $ 1.610$ \\
        \end{longtable*}
    \end{center}
    \begin{itemize}
        \item Watterson's $\thetaSyn$ is the observed genetic diversity calculated for synonymous changes.
        \item $\proba_{div}[\SphyBen]$ is the proportion of substitutions in the terminal branch that are $\SphyBen$.
        \item $\dn(\SphyDel) / \ds$ (eq.~6) is the ratio of non-synonymous over synonymous substitutions, when restricted to non-synonymous substitutions in the terminal branch that are $\SphyDel$.
        \item $\dn(\SphyNeu) / \ds$ (eq.~6) is the ratio of non-synonymous over synonymous substitutions, when restricted to non-synonymous substitutions in the terminal branch that are $\SphyNeu$.
        \item $\dn(\SphyBen) / \ds$ (eq.~6) is the ratio of non-synonymous over synonymous substitutions, when restricted to non-synonymous substitutions in the terminal branch that are $\SphyBen$.
    \end{itemize}

    \newpage

    \subsubsection{$\dnds$ over-estimation due to beneficial back-mutations}
    \begin{center}
        \captionof{table}{$\dnds$ over-estimation due to beneficial back-mutations.}
        \scriptsize
        \begin{longtable*}{|l|l|r|r|r|r|r|}
            \toprule
            Population & Species & Watterson's $\thetaSyn$ & $\proba_{div}[\Sphy > 1]$ & $\dnds $ & $\dn(\Sphy < 1) / \ds$ & $\delta(\dnds )$ \\
            \midrule
            \endhead
            \midrule
            \multicolumn{7}{r}{{Continued on next page}} \\
            \midrule
            \endfoot

            \bottomrule
            \endlastfoot
            \rowcolor{LIGHTGREY} Equus c. &      Equus caballus &               $ 0.002$ &                    $ 0.118$ &  $ 0.147$ &           $ 0.131$ &     $  10.5$ \\
            Iran &          Bos taurus &               $ 0.002$ &                    $ 0.123$ &  $ 0.131$ &           $ 0.117$ &     $  11.1$ \\
            Uganda &          Bos taurus &               $ 0.005$ &                    $ 0.125$ &  $ 0.134$ &           $ 0.118$ &     $  11.3$ \\
            \rowcolor{LIGHTGREY} Australia &        Capra hircus &               $ 0.003$ &                    $ 0.121$ &  $ 0.125$ &           $ 0.111$ &     $  10.8$ \\
            \rowcolor{LIGHTGREY} France &        Capra hircus &               $ 0.004$ &                    $ 0.120$ &  $ 0.126$ &           $ 0.112$ &     $  10.8$ \\
            \rowcolor{LIGHTGREY} Iran (C. aegagrus) &        Capra hircus &               $ 0.004$ &                    $ 0.120$ &  $ 0.125$ &           $ 0.111$ &     $  10.8$ \\
            \rowcolor{LIGHTGREY} Iran &        Capra hircus &               $ 0.004$ &                    $ 0.121$ &  $ 0.125$ &           $ 0.111$ &     $  10.9$ \\
            \rowcolor{LIGHTGREY} Italy &        Capra hircus &               $ 0.004$ &                    $ 0.120$ &  $ 0.126$ &           $ 0.112$ &     $  10.8$ \\
            \rowcolor{LIGHTGREY} Morocco &        Capra hircus &               $ 0.004$ &                    $ 0.122$ &  $ 0.125$ &           $ 0.111$ &     $  11.0$ \\
            Iran &          Ovis aries &               $ 0.007$ &                    $ 0.109$ &  $ 0.146$ &           $ 0.132$ &     $ 9.671$ \\
            Iran (O. orientalis) &          Ovis aries &               $ 0.009$ &                    $ 0.108$ &  $ 0.148$ &           $ 0.134$ &     $ 9.516$ \\
            Iran (O. vignei) &          Ovis aries &               $ 0.007$ &                    $ 0.109$ &  $ 0.146$ &           $ 0.132$ &     $ 9.598$ \\
            Various &          Ovis aries &               $ 0.008$ &                    $ 0.107$ &  $ 0.148$ &           $ 0.134$ &     $ 9.439$ \\
            Morocco &          Ovis aries &               $ 0.008$ &                    $ 0.108$ &  $ 0.148$ &           $ 0.134$ &     $ 9.582$ \\
            \rowcolor{LIGHTGREY} Barbados & Chlorocebus sabaeus &               $ 0.004$ &                    $ 0.122$ &  $ 0.133$ &           $ 0.118$ &     $  11.2$ \\
            \rowcolor{LIGHTGREY} Central Afr. Rep. & Chlorocebus sabaeus &               $ 0.006$ &                    $ 0.124$ &  $ 0.134$ &           $ 0.119$ &     $  11.3$ \\
            \rowcolor{LIGHTGREY} Ethiopia & Chlorocebus sabaeus &               $ 0.005$ &                    $ 0.124$ &  $ 0.134$ &           $ 0.118$ &     $  11.3$ \\
            \rowcolor{LIGHTGREY} Gambia & Chlorocebus sabaeus &               $ 0.005$ &                    $ 0.123$ &  $ 0.134$ &           $ 0.119$ &     $  11.2$ \\
            \rowcolor{LIGHTGREY} Kenya & Chlorocebus sabaeus &               $ 0.005$ &                    $ 0.123$ &  $ 0.134$ &           $ 0.119$ &     $  11.3$ \\
            \rowcolor{LIGHTGREY} Nevis & Chlorocebus sabaeus &               $ 0.003$ &                    $ 0.123$ &  $ 0.133$ &           $ 0.118$ &     $  11.3$ \\
            \rowcolor{LIGHTGREY} South Africa  & Chlorocebus sabaeus &               $ 0.006$ &                    $ 0.125$ &  $ 0.134$ &           $ 0.118$ &     $  11.5$ \\
            \rowcolor{LIGHTGREY} Saint Kitts & Chlorocebus sabaeus &               $ 0.004$ &                    $ 0.123$ &  $ 0.133$ &           $ 0.118$ &     $  11.3$ \\
            \rowcolor{LIGHTGREY} Zambia & Chlorocebus sabaeus &               $ 0.006$ &                    $ 0.123$ &  $ 0.134$ &           $ 0.119$ &     $  11.2$ \\
            African &        Homo sapiens &               $ 0.002$ &                    $ 0.099$ &  $ 0.192$ &           $ 0.175$ &     $ 8.799$ \\
            Admixed American &        Homo sapiens &               $ 0.002$ &                    $ 0.099$ &  $ 0.192$ &           $ 0.175$ &     $ 8.780$ \\
            East Asian &        Homo sapiens &               $ 0.002$ &                    $ 0.098$ &  $ 0.192$ &           $ 0.175$ &     $ 8.768$ \\
            European &        Homo sapiens &               $ 0.002$ &                    $ 0.098$ &  $ 0.192$ &           $ 0.175$ &     $ 8.728$ \\
            South Asian &        Homo sapiens &               $ 0.002$ &                    $ 0.099$ &  $ 0.192$ &           $ 0.175$ &     $ 8.788$ \\
        \end{longtable*}
    \end{center}
    \begin{itemize}
        \item Watterson's $\thetaSyn$ is the observed genetic diversity calculated for synonymous changes.
        \item $\proba_{div}[\SphyBen]$ is the proportion of substitutions in the terminal branch that are $\SphyBen$.
        \item $\dnds$ (eq.~6) is the ratio of non-synonymous over synonymous substitutions estimated for all the non-synonymous substitutions in the terminal branch.
        \item $\dn(\Sphy < 1) / \ds$ (eq.~6) is the ratio of non-synonymous over synonymous substitutions, when restricted to non-synonymous substitutions in the terminal branch that are not $\SphyBen$.
        This is the estimated divergence when we removed beneficial back-mutations.
        \item $\delta(\dnds)$ (eq.~7) is the fraction of the divergence ($\dnds$) that is over-estimated, $\dnds$ is compared to the estimated divergence when we removed beneficial back-mutations $\dn(\Sphy < 1) / \ds$.
    \end{itemize}
    We estimated that between 9 and 11\% of $\dnds$ is over-estimated, corresponding to beneficial back-mutation inflating the $\dnds$ statistic.

    \newpage

    \subsubsection{Probabilities of beneficial back-mutations}
    \begin{center}
        \captionof{table}{Probability of beneficial back-mutations among all beneficial ones.}
        \scriptsize
        \begin{longtable*}{|l|l|r|r|r|r|r|r|}
            \toprule
            Population & Species & Watterson's $\thetaSyn$ & $\proba[\SphyBen]$ & $\proba [ \SpopBen ]$ & $\frac{\proba[\SphyBen]}{\proba[ \SpopBen ]}$ & $\proba [ \SpopBen \given \SphyBen]$ & $\proba[\SphyBen\given \SpopBen ]$ \\
            \midrule
            \endhead
            \midrule
            \multicolumn{8}{r}{{Continued on next page}} \\
            \midrule
            \endfoot

            \bottomrule
            \endlastfoot
            \rowcolor{LIGHTGREY} Equus c. &      Equus caballus &               $ 0.002$ &              $ 0.014$ &              $ 0.010$ &                                          $ 1.395$ &                         $ 0.580$ &                      $ 0.809$ \\
            Iran &          Bos taurus &               $ 0.002$ &              $ 0.013$ &              $ 0.041$ &                                          $ 0.323$ &                         $ 0.857$ &                      $ 0.277$ \\
            Uganda &          Bos taurus &               $ 0.005$ &              $ 0.013$ &              $ 0.013$ &                                          $ 0.991$ &                         $ 0.488$ &                      $ 0.483$ \\
            \rowcolor{LIGHTGREY} Australia &        Capra hircus &               $ 0.003$ &              $ 0.014$ &              $ 0.026$ &                                          $ 0.532$ &                         $ 0.368$ &                      $ 0.196$ \\
            \rowcolor{LIGHTGREY} France &        Capra hircus &               $ 0.004$ &              $ 0.014$ &              $ 0.024$ &                                          $ 0.572$ &                         $ 0.368$ &                      $ 0.211$ \\
            \rowcolor{LIGHTGREY} Iran (C. aegagrus) &        Capra hircus &               $ 0.004$ &              $ 0.014$ &              $ 0.026$ &                                          $ 0.516$ &                         $ 0.221$ &                      $ 0.114$ \\
            \rowcolor{LIGHTGREY} Iran &        Capra hircus &               $ 0.004$ &              $ 0.014$ &              $ 0.024$ &                                          $ 0.573$ &                         $ 0.368$ &                      $ 0.211$ \\
            \rowcolor{LIGHTGREY} Italy &        Capra hircus &               $ 0.004$ &              $ 0.014$ &              $ 0.017$ &                                          $ 0.814$ &                         $ 0.270$ &                      $ 0.220$ \\
            \rowcolor{LIGHTGREY} Morocco &        Capra hircus &               $ 0.004$ &              $ 0.014$ &              $ 0.018$ &                                          $ 0.751$ &                         $ 0.368$ &                      $ 0.276$ \\
            Iran &          Ovis aries &               $ 0.007$ &              $ 0.014$ &              $ 0.006$ &                                          $ 2.321$ &                         $ 0.169$ &                      $ 0.392$ \\
            Iran (O. orientalis) &          Ovis aries &               $ 0.009$ &              $ 0.014$ &              $ 0.012$ &                                          $ 1.178$ &                         $ 0.153$ &                      $ 0.180$ \\
            Iran (O. vignei) &          Ovis aries &               $ 0.007$ &              $ 0.014$ &              $ 0.020$ &                                          $ 0.683$ &                         $ 0.195$ &                      $ 0.133$ \\
            Various &          Ovis aries &               $ 0.008$ &              $ 0.014$ &              $ 0.012$ &                                          $ 1.201$ &                         $ 0.190$ &                      $ 0.228$ \\
            Morocco &          Ovis aries &               $ 0.008$ &              $ 0.014$ &              $ 0.005$ &                                          $ 2.589$ &                         $ 0.136$ &                      $ 0.353$ \\
            \rowcolor{LIGHTGREY} Barbados & Chlorocebus sabaeus &               $ 0.004$ &              $ 0.012$ &              $ 0.022$ &                                          $ 0.541$ &                         $ 0.565$ &                      $ 0.306$ \\
            \rowcolor{LIGHTGREY} Central Afr. Rep. & Chlorocebus sabaeus &               $ 0.006$ &              $ 0.012$ &              $ 0.018$ &                                          $ 0.639$ &                         $ 0.429$ &                      $ 0.274$ \\
            \rowcolor{LIGHTGREY} Ethiopia & Chlorocebus sabaeus &               $ 0.005$ &              $ 0.012$ &              $ 0.022$ &                                          $ 0.516$ &                         $ 0.462$ &                      $ 0.239$ \\
            \rowcolor{LIGHTGREY} Gambia & Chlorocebus sabaeus &               $ 0.005$ &              $ 0.012$ &              $ 0.007$ &                                          $ 1.592$ &                         $ 0.486$ &                      $ 0.774$ \\
            \rowcolor{LIGHTGREY} Kenya & Chlorocebus sabaeus &               $ 0.005$ &              $ 0.012$ &              $ 0.023$ &                                          $ 0.508$ &                         $ 0.478$ &                      $ 0.243$ \\
            \rowcolor{LIGHTGREY} Nevis & Chlorocebus sabaeus &               $ 0.003$ &              $ 0.012$ &              $ 0.015$ &                                          $ 0.762$ &                         $ 0.516$ &                      $ 0.393$ \\
            \rowcolor{LIGHTGREY} South Africa  & Chlorocebus sabaeus &               $ 0.006$ &              $ 0.012$ &              $ 0.016$ &                                          $ 0.734$ &                         $ 0.490$ &                      $ 0.360$ \\
            \rowcolor{LIGHTGREY} Saint Kitts & Chlorocebus sabaeus &               $ 0.004$ &              $ 0.012$ &              $ 0.016$ &                                          $ 0.713$ &                         $ 0.515$ &                      $ 0.367$ \\
            \rowcolor{LIGHTGREY} Zambia & Chlorocebus sabaeus &               $ 0.006$ &              $ 0.012$ &              $ 0.023$ &                                          $ 0.500$ &                         $ 0.499$ &                      $ 0.249$ \\
            African &        Homo sapiens &               $ 0.002$ &              $ 0.012$ &              $ 0.025$ &                                          $ 0.470$ &                         $ 0.678$ &                      $ 0.319$ \\
            Admixed American &        Homo sapiens &               $ 0.002$ &              $ 0.012$ &              $ 0.021$ &                                          $ 0.569$ &                         $ 0.635$ &                      $ 0.361$ \\
            East Asian &        Homo sapiens &               $ 0.002$ &              $ 0.012$ &              $ 0.031$ &                                          $ 0.381$ &                         $ 0.611$ &                      $ 0.233$ \\
            European &        Homo sapiens &               $ 0.002$ &              $ 0.012$ &              $ 0.034$ &                                          $ 0.350$ &                         $ 0.619$ &                      $ 0.217$ \\
            South Asian &        Homo sapiens &               $ 0.002$ &              $ 0.012$ &              $ 0.033$ &                                          $ 0.354$ &                         $ 0.628$ &                      $ 0.222$ \\
        \end{longtable*}
    \end{center}
    \begin{itemize}
        \item Watterson's $\thetaSyn$ is the observed genetic diversity calculated for synonymous changes.
        \item $\proba [ \SphyBen ]$ (eq.~5) is the probability for a new mutation to be a beneficial back-mutation.
        These mutations have a selection coefficient predicted at the phylogenetic-scale larger than 1, thus toward a more fit amino-acid.
        \item $\proba [ \SpopBen ]$ (eq.~15) is the probability for a mutation to be beneficial.
        These mutations have a selection coefficient at the population-scale larger than 1.
        \item $\proba [ \SpopBen \given \SphyBen]$ (eq.~12) is the probability for a mutation to be beneficial at the population scale, given it is predicted to be a beneficial back-mutation at the phylogenetic scale (the \textit{precision}).
        \item $\proba [ \SphyBen \given \SpopBen]$ (eq.~14) is the probability for a mutation to a beneficial back-mutations, given it is beneficial at the population scale (the \textit{recall}).
        This probability is obtained using Bayes' formula.
    \end{itemize}
    \newpage

    \subsection{Replicability across mammals}
    The proportion of deleterious ($\ProbaPopDel$), nearly-neutral ($\ProbaPopNeu$) and of beneficial ($\ProbaPopBen$) mutations estimated at the population-genetic scale across the different populations is shown for each class of selection ($\Sphyclass \in \{\SphyDel, \SphyNeu, \SphyBen \}$).

    \captionof{figure}{Estimation of selection at the population scale for $\SphyDel$ mutations}
    \begin{minipage}{0.9\linewidth}
    \includegraphics[width=\linewidth, page=1]{artworks/Theta.neg.stacked.pdf}
    \end{minipage}
    \begin{minipage}{0.09\linewidth}
        \includegraphics[width=\linewidth, page=1]{artworks/legend.polycat}
    \end{minipage}

    \captionof{figure}{Estimation of selection at the population scale for $\SphyNeu$ mutations}
    \begin{minipage}{0.9\linewidth}
        \includegraphics[width=\linewidth, page=1]{artworks/Theta.weak.stacked.pdf}
    \end{minipage}
    \begin{minipage}{0.09\linewidth}
        \includegraphics[width=\linewidth, page=1]{artworks/legend.polycat}
    \end{minipage}

    \newpage
    \captionof{figure}{Estimation of selection at the population scale for $\SphyBen$ mutations}
    \begin{minipage}{0.9\linewidth}
        \includegraphics[width=\linewidth, page=1]{artworks/Theta.pos.stacked.pdf}
    \end{minipage}
    \begin{minipage}{0.09\linewidth}
        \includegraphics[width=\linewidth, page=1]{artworks/legend.polycat}
    \end{minipage}

    \subsection{Correlation with diversity}
    For each population, the proportion of deleterious ($\ProbaPopDel$), nearly-neutral ($\ProbaPopNeu$) and of beneficial ($\ProbaPopBen$) mutations estimated at the population-genetic scale is shown as function of Watterson's's diversity ($\thetaSyn$).

    \subsubsection{Proportion of deleterious mutations ($\SpopDel$)}\label{subsec:proportion-deleterious-mutations}
    \captionof{figure}{Proportion of deleterious mutations ($\SpopDel$)}
    \begin{minipage}{0.49\linewidth}
        \includegraphics[width=\linewidth, page=1]{artworks/results.watterson.all_P-Sneg.scatter.pdf}
    \end{minipage}
    \begin{minipage}{0.49\linewidth}
        \includegraphics[width=\linewidth, page=1]{artworks/results.watterson.neg_P-Sneg.scatter.pdf}
    \end{minipage}
    \begin{itemize}
        \item Watterson's $\thetaSyn$ is the observed genetic diversity calculated for synonymous changes.
        \item $\proba [ \SpopDel ]$ (eq.~15) is the probability for a mutation to be deleterious.
        These mutations have a selection coefficient at the population-scale lower than -1.
        \item $\proba [ \SpopDel \given \SphyDel ]$ (eq.~12) is the probability for a mutation to be deleterious at the population scale, given it is predicted to be a deleterious at the phylogenetic scale.
    \end{itemize}

    We can see that higher synonymous diversity ($\thetaSyn$) is typically accompanied by a higher proportion of effectively deleterious mutations at the population scale ($\proba [ \SpopDel ]$).
    This trend is also confirmed when we restricted the analysis to class of mutations that are supposedly deleterious at the phylogenetic scale ($\SphyDel$).
    This result is qualitatively in accordance with the nearly-neutral theory of evolution which argues that very slightly deleterious mutations are more efficiently purified in large populations.

    \subsubsection{Proportion of nearly-neutral mutations ($\SpopNeu$)}\label{subsec:proportion-nearly-neutral-mutations}
    \captionof{figure}{Proportion of nearly-neutral mutations ($\SpopNeu$)}
    \begin{minipage}{0.49\linewidth}
        \includegraphics[width=\linewidth, page=1]{artworks/results.watterson.all_P-Sweak.scatter.pdf}
    \end{minipage}
    \begin{minipage}{0.49\linewidth}
        \includegraphics[width=\linewidth, page=1]{artworks/results.watterson.weak_P-Sweak.scatter.pdf}
    \end{minipage}
    \begin{itemize}
        \item Watterson's $\thetaSyn$ is the observed genetic diversity calculated for synonymous changes.
        \item $\proba [ \SpopNeu ]$ (eq.~15) is the probability for a mutation to be nearly-neutral.
        These mutations have a selection coefficient at the population-scale between -1 and 1.
        \item $\proba [ \SpopNeu \given \SphyNeu ]$ (eq.~12) is the probability for a mutation to be nearly-neutral at the population scale, given it is predicted to be a nearly-neutral at the phylogenetic scale.
    \end{itemize}

    We confirmed that higher synonymous diversity ($\thetaSyn$) is typically accompanied by a smaller proportion of neutral mutations at the population scale ($\proba [ \SpopNeu ]$ in the range 0.06-0.18).
    This result is more pronounced ($\proba [ \SpopNeu \given \SphyNeu ]$ in the range 0.36-0.73) when we restricted the analysis to class of mutations that are supposedly nearly-neutral at the phylogenetic scale ($\SphyNeu$).
    This result suggests that populations with higher diversity (e.g.~\textit{Bos} or \textit{Ovis}) are more likely to discriminate whether mutations are beneficial or deleterious.
    Alternatively stated, mutations in populations with low diversity (e.g.~\textit{Homo}) are effectively nearly-neutral and behave as would a neutral mutation.
    This result is qualitatively in accordance with the nearly-neutral theory of evolution which argues that mutations are less efficiently selected for in small populations.

    \subsubsection{Proportion of beneficial mutations ($\SpopBen$)}\label{subsec:proportion-beneficial-mutations}
    \captionof{figure}{Proportion of beneficial mutations ($\SpopBen$)}
    \begin{minipage}{0.49\linewidth}
        \includegraphics[width=\linewidth, page=1]{artworks/results.watterson.all_P-Spos.scatter.pdf}
    \end{minipage}
    \begin{minipage}{0.49\linewidth}
        \includegraphics[width=\linewidth, page=1]{artworks/results.watterson.pos_P-Spos.scatter.pdf}
    \end{minipage}
    \begin{itemize}
        \item Watterson's $\thetaSyn$ is the observed genetic diversity calculated for synonymous changes.
        \item $\proba [ \SpopBen ]$ (eq.~15) is the probability for a mutation to be beneficial.
        These mutations have a selection coefficient at the population-scale larger than 1.
        \item $\proba [ \SpopBen \given \SphyBen]$ (eq.~12) is the probability for a mutation to be beneficial at the population scale, given it is predicted to be a beneficial back-mutation at the phylogenetic scale (the \textit{precision}).
    \end{itemize}
    Higher synonymous diversity ($\thetaSyn$) is accompanied by a smaller proportion of beneficial mutations at the population scale ($\proba [ \SpopBen ]$).
    This trend is also confirmed when we restricted the analysis to class of mutations that are supposedly beneficial back-mutations at the phylogenetic scale ($\SphyBen$).
    This result is much more difficult to interpret.
    The fraction of beneficial back mutations is supposed to depend on long term demographic history, which is not directly accessible.
    If the long term effective population size is relatively similar to the short term one, we expect little opportunity for beneficial back-mutations.
    And thus, counter-intuitively we expect a diminution of positively selected mutations.
    However, we can see that the proportion of beneficial back-mutations among advantageous one does not decrease with synonymous diversity.
    This means that somehow adaptive mutations also decrease with synonymous diversity for a reason we fail to explain.

    \newpage

    \section{Clinically related terms for mutations}
    \subsection{Terms associated with deleterious mutations $\SphyDel$}
    \begin{center}
        \captionof{table}{Terms associated with deleterious mutations $\SphyDel$}
        \begin{tabular}{|l|r|r|r|r|r|}
            \toprule
            SNP clinical ontology & $n_{\mathrm{Observed}}$ & $n_{\mathrm{Expected}}$ & Odds ratio & $p_{\mathrm{v}}$ & $p_{\mathrm{v-adjusted}}$ \\
            \midrule
            Benign                & 2969                    & $4043.0$                & $ 0.734$   & $ 1.000$             & $ 1.000~~$                    \\
            Likely benign         & 2994                    & $3399.8$                & $ 0.881$   & $ 0.999$             & $ 1.000~~$                    \\
            Risk factor           & 102                     & $ 118.2$                & $ 0.863$   & $ 0.798$             & $ 1.000~~$                    \\
            Likely pathogenic     & 221                     & $  68.5$                  & $ 3.226$   & $1.7\times 10^{-8}$  & $\bm{6.7\times 10^{-8}{^*}}$  \\
            Pathogenic            & 560                     & $ 193.6$                & $ 2.893$   & $4.2\times 10^{-17}$ & $\bm{2.1\times 10^{-16}{^*}}$ \\
            \bottomrule
        \end{tabular}
    \end{center}
    In humans, non-synonymous SNPs in the test group ($\SphyDel$) are contrasted to SNPs in the control group ($\SphyNeu$).
    For each clinical term, a 2x2 contingency tables is built by counting the number of SNPs based on their selection coefficient and their clinical terms (whether they have this specific term or not).
    Fisher's exact tests are then performed for these 2x2 contingency tables.
    $^*$ for $p_{\mathrm{v}}^{\mathrm{adj}}$ corrected for multiple comparison (Holm–Bonferroni correction) lower than the risk $\alpha=0.05$.
    SNPs predicted with $\SphyDel$ are statistically associated to clinical terms such as \textit{Likely Pathogenic} and \textit{Pathogenic}.

    \subsection{Terms associated with beneficial back-mutations $\SphyBen$}
    \begin{center}
        \captionof{table}{Terms associated with beneficial back-mutations $\SphyBen$}
        \begin{tabular}{|l|r|r|r|r|r|}
            \toprule
            SNP clinical ontology & $n_{\mathrm{Observed}}$ & $n_{\mathrm{Expected}}$ & Odds ratio & $p_{\mathrm{v}}$ & $p_{\mathrm{v-adjusted}}$ \\
            \midrule
            Benign                & 319                     & $ 261.7$                & $ 1.219$   & $ 0.002$         & $\bm{ 0.009{^*}}$         \\
            Likely benign         & 263                     & $ 222.7$                & $ 1.181$   & $ 0.012$         & $\bm{ 0.049{^*}}$         \\
            Risk factor           & 5                       & $ 7.847$                & $ 0.637$   & $ 0.879$         & $ 0.879~~$                \\
            Likely pathogenic     & 7                       & $ 4.552$                & $ 1.538$   & $ 0.227$         & $ 0.682~~$                \\
            Pathogenic            & 16                      & $  12.9$                  & $ 1.241$   & $ 0.268$         & $ 0.682~~$                \\
            \bottomrule
        \end{tabular}
    \end{center}
    In humans, non-synonymous SNPs in the test group ($\SphyBen$) are contrasted to SNPs in the control group ($\SphyNeu$).
    For each clinical term, a 2x2 contingency tables is built by counting the number of SNPs based on their selection coefficient and their clinical terms (whether they have this specific term or not).
    Fisher's exact tests are then performed for these 2x2 contingency tables.
    $^*$ for $p_{\mathrm{v}}^{\mathrm{adj}}$ corrected for multiple comparison (Holm–Bonferroni correction) lower than the risk $\alpha=0.05$.
    Beneficial back-mutations are associated with clinical terms such as \textit{Benign} and \textit{Likely Benign}.
    \newpage

    \section{Including divergence data for the estimation of $\Spop$}

    \subsection{Including divergence in polyDFE}
    Divergence data (number of substitutions per site) can also be included into polyDFE to estimate the DFE.
    For the class of a given class of selection coefficient ($\Sphyclass \in \{\SphyDel, \SphyNeu, \SphyBen \}$), the number of substitutions has already been computed and is given by $D\left( \Sphyclass \right)$ (see eq.~6), and the number of sites is given by $L \left( \Sphyclass \right)$ (see eq.~4).
    Otherwise, the procedure is the same as described section \textit{Scaled selection coefficients ($\Spop$) in a population-based method}.

    \subsection{Probabilities of beneficial back-mutations}
    \begin{center}
        \captionof{table}{Probability of beneficial back-mutations among all beneficial ones.}
        \scriptsize
        \begin{longtable*}{|l|l|r|r|r|r|r|r|}
            \toprule
            Population & Species & Watterson's $\thetaSyn$ & $\proba[\SphyBen]$ & $\proba [ \SpopBen ]$ & $\frac{\proba[\SphyBen]}{\proba[ \SpopBen ]}$ & $\proba [ \SpopBen \given \SphyBen]$ & $\proba[\SphyBen\given \SpopBen ]$ \\
            \midrule
            \endhead
            \midrule
            \multicolumn{8}{r}{{Continued on next page}} \\
            \midrule
            \endfoot

            \bottomrule
            \endlastfoot
            \rowcolor{LIGHTGREY} Equus c. &      Equus caballus &               $ 0.002$ &              $ 0.014$ &              $ 0.061$ &                                          $ 0.235$ &                         $ 0.519$ &                      $ 0.122$ \\
            Iran &          Bos taurus &               $ 0.002$ &              $ 0.013$ &              $ 0.007$ &                                          $ 1.908$ &                         $ 0.368$ &                      $ 0.702$ \\
            Uganda &          Bos taurus &               $ 0.005$ &              $ 0.013$ &              $ 0.008$ &                                          $ 1.667$ &                         $ 0.255$ &                      $ 0.425$ \\
            \rowcolor{LIGHTGREY} Australia &        Capra hircus &               $ 0.003$ &              $ 0.014$ &              $ 0.007$ &                                          $ 2.099$ &                         $ 0.147$ &                      $ 0.308$ \\
            \rowcolor{LIGHTGREY} France &        Capra hircus &               $ 0.004$ &              $ 0.014$ &              $ 0.007$ &                                          $ 2.043$ &                         $ 0.149$ &                      $ 0.304$ \\
            \rowcolor{LIGHTGREY} Iran (C. aegagrus) &        Capra hircus &               $ 0.004$ &              $ 0.014$ &              $ 0.013$ &                                          $ 1.069$ &                         $ 0.147$ &                      $ 0.157$ \\
            \rowcolor{LIGHTGREY} Iran &        Capra hircus &               $ 0.004$ &              $ 0.014$ &              $ 0.008$ &                                          $ 1.699$ &                         $ 0.152$ &                      $ 0.258$ \\
            \rowcolor{LIGHTGREY} Italy &        Capra hircus &               $ 0.004$ &              $ 0.014$ &              $ 0.006$ &                                          $ 2.160$ &                         $ 0.147$ &                      $ 0.316$ \\
            \rowcolor{LIGHTGREY} Morocco &        Capra hircus &               $ 0.004$ &              $ 0.014$ &              $ 0.006$ &                                          $ 2.141$ &                         $ 0.186$ &                      $ 0.398$ \\
            Iran &          Ovis aries &               $ 0.007$ &              $ 0.014$ &              $ 0.012$ &                                          $ 1.210$ &                         $ 0.158$ &                      $ 0.191$ \\
            Iran (O. orientalis) &          Ovis aries &               $ 0.009$ &              $ 0.014$ &              $ 0.012$ &                                          $ 1.144$ &                         $ 0.157$ &                      $ 0.180$ \\
            Iran (O. vignei) &          Ovis aries &               $ 0.007$ &              $ 0.014$ &              $ 0.022$ &                                          $ 0.631$ &                         $ 0.186$ &                      $ 0.118$ \\
            Various &          Ovis aries &               $ 0.008$ &              $ 0.014$ &              $ 0.012$ &                                          $ 1.211$ &                         $ 0.156$ &                      $ 0.188$ \\
            Morocco &          Ovis aries &               $ 0.008$ &              $ 0.014$ &              $ 0.009$ &                                          $ 1.536$ &                         $ 0.190$ &                      $ 0.292$ \\
            \rowcolor{LIGHTGREY} Barbados & Chlorocebus sabaeus &               $ 0.004$ &              $ 0.012$ &              $ 0.005$ &                                          $ 2.230$ &                         $ 0.282$ &                      $ 0.630$ \\
            \rowcolor{LIGHTGREY} Central Afr. Rep. & Chlorocebus sabaeus &               $ 0.006$ &              $ 0.012$ &              $ 0.014$ &                                          $ 0.849$ &                         $ 0.310$ &                      $ 0.263$ \\
            \rowcolor{LIGHTGREY} Ethiopia & Chlorocebus sabaeus &               $ 0.005$ &              $ 0.012$ &              $ 0.008$ &                                          $ 1.530$ &                         $ 0.298$ &                      $ 0.456$ \\
            \rowcolor{LIGHTGREY} Gambia & Chlorocebus sabaeus &               $ 0.005$ &              $ 0.012$ &              $ 0.008$ &                                          $ 1.384$ &                         $ 0.305$ &                      $ 0.422$ \\
            \rowcolor{LIGHTGREY} Kenya & Chlorocebus sabaeus &               $ 0.005$ &              $ 0.012$ &              $ 0.014$ &                                          $ 0.854$ &                         $ 0.296$ &                      $ 0.253$ \\
            \rowcolor{LIGHTGREY} Nevis & Chlorocebus sabaeus &               $ 0.003$ &              $ 0.012$ &              $ 0.005$ &                                          $ 2.438$ &                         $ 0.285$ &                      $ 0.696$ \\
            \rowcolor{LIGHTGREY} South Africa  & Chlorocebus sabaeus &               $ 0.006$ &              $ 0.012$ &              $ 0.011$ &                                          $ 1.047$ &                         $ 0.323$ &                      $ 0.338$ \\
            \rowcolor{LIGHTGREY} Saint Kitts & Chlorocebus sabaeus &               $ 0.004$ &              $ 0.012$ &              $ 0.006$ &                                          $ 2.070$ &                         $ 0.292$ &                      $ 0.604$ \\
            \rowcolor{LIGHTGREY} Zambia & Chlorocebus sabaeus &               $ 0.006$ &              $ 0.012$ &              $ 0.014$ &                                          $ 0.816$ &                         $ 0.309$ &                      $ 0.252$ \\
            African &        Homo sapiens &               $ 0.002$ &              $ 0.012$ &              $ 0.015$ &                                          $ 0.765$ &                         $ 0.371$ &                      $ 0.284$ \\
            Admixed American &        Homo sapiens &               $ 0.002$ &              $ 0.012$ &              $ 0.009$ &                                          $ 1.259$ &                         $ 0.368$ &                      $ 0.464$ \\
            East Asian &        Homo sapiens &               $ 0.002$ &              $ 0.012$ &              $ 0.012$ &                                          $ 0.973$ &                         $ 0.369$ &                      $ 0.358$ \\
            European &        Homo sapiens &               $ 0.002$ &              $ 0.012$ &              $ 0.009$ &                                          $ 1.356$ &                         $ 0.368$ &                      $ 0.499$ \\
            South Asian &        Homo sapiens &               $ 0.002$ &              $ 0.012$ &              $ 0.010$ &                                          $ 1.147$ &                         $ 0.369$ &                      $ 0.423$ \\
        \end{longtable*}
    \end{center}
    \begin{itemize}
        \item Watterson's $\thetaSyn$ is the observed genetic diversity calculated for synonymous changes.
        \item $\proba [ \SphyBen ]$ (eq.~5) is the probability for a new mutation a beneficial back-mutation.
        These mutations have a selection coefficient predicted at the phylogenetic-scale larger than 1, thus toward a more fit amino-acid.
        \item $\proba [ \SpopBen ]$ (eq.~15) is the probability for a mutation to be beneficial.
        These mutations have a selection coefficient at the population-scale larger than 1.
        \item $\proba [ \SpopBen \given \SphyBen]$ (eq.~12) is the probability for a mutation to be beneficial at the population scale, given it is predicted to be a beneficial back-mutation at the phylogenetic scale (the \textit{precision}).
        \item $\proba [ \SphyBen \given \SpopBen]$ (eq.~14) is the probability for a mutation to a beneficial back-mutations, given it is beneficial (the \textit{recall}).
        This probability is obtained using Bayes' formula.
    \end{itemize}
    \newpage

    \subsection{Precision and recall}
    \captionof{table}{Precision and recall}
    \begin{center}
        \begin{adjustbox}{width=\textwidth}
            \begin{tabular}{||l|l|r||r|r||r|r||r|r||}
                \toprule
                \multicolumn{3}{||c||}{} &
                \multicolumn{2}{c||}{\shortstack{\textbf{Deleterious mutations} \\ $\bm{\SpopDel \coloneqq \Spop<-1}$ \\ $\bm{\SphyDel \coloneqq \Sphy<-1}$}} &
                \multicolumn{2}{c||}{\shortstack{\textbf{Nearly-neutral mutations} \\ $\bm{\SpopNeu \coloneqq -1<\Spop<1}$ \\ $\bm{\SphyNeu \coloneqq -1<\Sphy<1}$}} &
                \multicolumn{2}{c||}{\shortstack{\textbf{Beneficial mutations} \\ $\bm{\SpopBen \coloneqq \Spop>1}$ \\ $\bm{\SphyBen \coloneqq \Sphy>1}$}}
                \\ \hline
                \textbf{Population} &
                \textbf{Species} &
                $\bm{\thetaSyn}$ &
                \makecell{\textbf{Precision} \\ $\bm{\proba [\SpopDel \given \SphyDel]}$} &
                \makecell{\textbf{Recall} \\ $\bm{\proba [\SphyDel \given \SpopDel]}$}           &
                \makecell{\textbf{Precision}      \\ $\bm{\proba [\SpopNeu \given \SphyNeu]}$}                                    &
                \makecell{\textbf{Recall}          \\ $\bm{\proba [\SphyNeu \given \SpopNeu]}$}                                  &
                \makecell{\textbf{Precision}          \\ $\bm{\proba [\SpopBen \given \SphyBen]}$}          &
                \makecell{\textbf{Recall}        \\ $\bm{\proba [\SphyBen \given \SpopBen]}$}
                \\   \midrule
                \rowcolor{LIGHTGREY} Equus c. & Equus caballus        & $ 0.002$ & $ 0.918$ & $ 0.957$ & $ 0.186$ & $ 0.213$ & $ 0.519$ & $ 0.122$ \\
                Iran               & Bos taurus        & $ 0.002$ & $ 0.924$ & $ 0.965$ & $ 0.565$ & $ 0.356$ & $ 0.368$ & $ 0.702$ \\
                Uganda              & Bos taurus        & $ 0.005$ & $ 0.943$ & $ 0.961$ & $ 0.516$ & $ 0.396$ & $ 0.255$ & $ 0.425$ \\
                \rowcolor{LIGHTGREY} Australia            & Capra hircus        & $ 0.003$ & $ 0.941$ & $ 0.964$ & $ 0.592$ & $ 0.434$ & $ 0.147$ & $ 0.308$ \\
                \rowcolor{LIGHTGREY} France                                    & Capra hircus          & $ 0.004$ & $ 0.940$ & $ 0.963$ & $ 0.571$ & $ 0.418$ & $ 0.149$ & $ 0.304$ \\
                \rowcolor{LIGHTGREY} Iran (C. aegagrus)                    & Capra hircus          & $ 0.004$ & $ 0.948$ & $ 0.960$ & $ 0.515$ & $ 0.446$ & $ 0.147$ & $ 0.157$ \\
                \rowcolor{LIGHTGREY} Iran                        & Capra hircus          & $ 0.004$ & $ 0.942$ & $ 0.961$ & $ 0.534$ & $ 0.410$ & $ 0.152$ & $ 0.258$ \\
                \rowcolor{LIGHTGREY} Italy                                 & Capra hircus          & $ 0.004$ & $ 0.940$ & $ 0.963$ & $ 0.579$ & $ 0.421$ & $ 0.147$ & $ 0.316$ \\
                \rowcolor{LIGHTGREY} Morocco                                 & Capra hircus          & $ 0.004$ & $ 0.939$ & $ 0.960$ & $ 0.549$ & $ 0.404$ & $ 0.186$ & $ 0.398$ \\
                Iran           & Ovis aries & $ 0.007$ & $ 0.957$ & $ 0.954$ & $ 0.412$ & $ 0.416$ & $ 0.158$ & $ 0.191$ \\
                Iran (O. orientalis)  & Ovis aries & $ 0.009$ & $ 0.961$ & $ 0.955$ & $ 0.413$ & $ 0.437$ & $ 0.157$ & $ 0.180$ \\
                Iran (O. vignei)           & Ovis aries & $ 0.007$ & $ 0.968$ & $ 0.951$ & $ 0.356$ & $ 0.516$ & $ 0.186$ & $ 0.118$ \\
                Various             & Ovis aries & $ 0.008$ & $ 0.958$ & $ 0.955$ & $ 0.428$ & $ 0.431$ & $ 0.156$ & $ 0.188$ \\
                Morocco              & Ovis aries & $ 0.008$ & $ 0.960$ & $ 0.953$ & $ 0.447$ & $ 0.460$ & $ 0.190$ & $ 0.292$ \\
                \rowcolor{LIGHTGREY} Barbados              & Chlorocebus sabaeus & $ 0.004$ & $ 0.931$ & $ 0.969$ & $ 0.628$ & $ 0.402$ & $ 0.282$ & $ 0.630$ \\
                \rowcolor{LIGHTGREY} Central Afr. Rep.       & Chlorocebus sabaeus & $ 0.006$ & $ 0.941$ & $ 0.964$ & $ 0.514$ & $ 0.408$ & $ 0.310$ & $ 0.263$ \\
                \rowcolor{LIGHTGREY} Ethiopia        & Chlorocebus sabaeus & $ 0.005$ & $ 0.933$ & $ 0.967$ & $ 0.586$ & $ 0.393$ & $ 0.298$ & $ 0.456$ \\
                \rowcolor{LIGHTGREY} Gambia             & Chlorocebus sabaeus & $ 0.005$ & $ 0.937$ & $ 0.967$ & $ 0.590$ & $ 0.414$ & $ 0.305$ & $ 0.422$ \\
                \rowcolor{LIGHTGREY} Kenya                                 & Chlorocebus sabaeus        & $ 0.005$ & $ 0.942$ & $ 0.965$ & $ 0.526$ & $ 0.418$ & $ 0.296$ & $ 0.253$ \\
                \rowcolor{LIGHTGREY} Nevis                        & Chlorocebus sabaeus        & $ 0.003$ & $ 0.930$ & $ 0.970$ & $ 0.637$ & $ 0.401$ & $ 0.285$ & $ 0.696$ \\
                \rowcolor{LIGHTGREY} South Africa                              & Chlorocebus sabaeus        & $ 0.006$ & $ 0.937$ & $ 0.964$ & $ 0.530$ & $ 0.393$ & $ 0.323$ & $ 0.338$ \\
                \rowcolor{LIGHTGREY} Saint Kitts                                & Chlorocebus sabaeus        & $ 0.004$ & $ 0.931$ & $ 0.969$ & $ 0.611$ & $ 0.393$ & $ 0.292$ & $ 0.604$ \\
                \rowcolor{LIGHTGREY} Zambia                             & Chlorocebus sabaeus        & $ 0.006$ & $ 0.941$ & $ 0.963$ & $ 0.506$ & $ 0.408$ & $ 0.309$ & $ 0.252$ \\
                African                             & Homo sapiens        & $ 0.002$ & $ 0.896$ & $ 0.968$ & $ 0.550$ & $ 0.297$ & $ 0.371$ & $ 0.284$ \\
                Admixed American                             & Homo sapiens        & $ 0.002$ & $ 0.890$ & $ 0.979$ & $ 0.710$ & $ 0.333$ & $ 0.368$ & $ 0.464$ \\
                East Asian                             & Homo sapiens        & $ 0.002$ & $ 0.890$ & $ 0.971$ & $ 0.578$ & $ 0.289$ & $ 0.369$ & $ 0.358$ \\
                European                             & Homo sapiens        & $ 0.002$ & $ 0.891$ & $ 0.980$ & $ 0.739$ & $ 0.345$ & $ 0.368$ & $ 0.499$ \\
                South Asian                             & Homo sapiens        & $ 0.002$ & $ 0.895$ & $ 0.981$ & $ 0.740$ & $ 0.356$ & $ 0.369$ & $ 0.423$ \\
                \bottomrule
            \end{tabular}
        \end{adjustbox}
    \end{center}
    \begin{itemize}
        \item Watterson's $\thetaSyn$ is the observed genetic diversity calculated for synonymous changes.
        \item \textit{Precision} is the estimation of the selection coefficient at population scale ($\Spop$) given that $\Sphy$ is known.
        \item  \textit{Recall} is the estimation of $\Sphy$ given selection coefficient at the population scale ($\Spop$) is known.
        \item \textit{Recall} for beneficial mutations ($\proba [\SphyBen \given \SpopBen]$) is thus the proportion of beneficial back-mutations among all beneficial mutations.
    \end{itemize}

    Altogether, comparing this table to tables 1 and S3, we acknowledge that the exact proportion of beneficial back-mutations among all beneficial ones is different whether we included or not substitutions in the terminal lineage for the estimation of $\Spop$.
    However, we can still see that beneficial back-mutations are positively selected compared to neutral and deleterious ones, and this result is not sensitive to inclusion of substitutions in the terminal lineage.

    \newpage
    \section{Discrete distribution of $\Spop$ at the population scale}

    \subsection{polyDFE model D (discrete distribution)}
    With polyDFE model D, the DFE of non-synonymous mutations is given as a discrete DFE of $K$ categories (instead of a continuous distribution in model C), where the selection coefficients of each category $i$ ($1 \leq i \leq K$) are fixed parameters $\Spop_i$, and each value $\Spop_i$ has a probability $p_i$ (estimated), with $\sum_{i=1}^{K} p_i =1$.
    We used $K=6$ with $\Spop_1 = -500$, $\Spop_2 = -4$, $\Spop_3 =-1$, $\Spop_4 = 0$, $\Spop_5 = 1$, $\Spop_6 = 4$.\\
    For each class of selection $\Sphyclass$, the parameters $p_i$ ($i \in \{ 1 \leq i \leq 6 \}$) were used to compute $\proba [ \SpopDel \given  \Sphyclass] $, $\proba [ \SpopNeu \given \Sphyclass]$, and $\proba [ \SpopBen \given \Sphyclass]$ as:
    \begin{align}
        \proba [ \SpopDel \given  \Sphyclass] &= \proba [ \Spop < -1 \given \Sphyclass ] = p_1 + p_2 \label{eq:polyProbaDel} \\
        \proba [ \SpopNeu \given \Sphyclass] &= \proba [ -1 < \Spop < 1 \given \Sphyclass ] = p_3 + p_4 + p_5,  \\
        \proba [ \SpopBen \given \Sphyclass] &= p_6. \label{eq:polyProbaAdv}
    \end{align}

    \subsection{Probabilities of beneficial back-mutations}
    \begin{center}
        \captionof{table}{Probability of beneficial back-mutations among all beneficial ones.}
        \scriptsize
        \begin{longtable*}{|l|l|r|r|r|r|r|r|}
            \toprule
            Population & Species & Watterson's $\thetaSyn$ & $\proba[\SphyBen]$ & $\proba [ \SpopBen ]$ & $\frac{\proba[\SphyBen]}{\proba[ \SpopBen ]}$ & $\proba [ \SpopBen \given \SphyBen]$ & $\proba[\SphyBen\given \SpopBen ]$ \\
            \midrule
            \endhead
            \midrule
            \multicolumn{8}{r}{{Continued on next page}} \\
            \midrule
            \endfoot

            \bottomrule
            \endlastfoot
            \rowcolor{LIGHTGREY} Equus c. &      Equus caballus &               $ 0.002$ &              $ 0.014$ &              $ 0.006$ &                                          $ 2.230$ &                         $ 0.141$ &                      $ 0.316$ \\
            Iran &          Bos taurus &               $ 0.002$ &              $ 0.013$ &              $ 0.006$ &                                          $ 2.200$ &                         $ 0.136$ &                      $ 0.298$ \\
            Uganda &          Bos taurus &               $ 0.005$ &              $ 0.013$ &              $ 0.011$ &                                          $ 1.177$ &                         $ 0.138$ &                      $ 0.162$ \\
            \rowcolor{LIGHTGREY} Australia &        Capra hircus &               $ 0.003$ &              $ 0.014$ &              $ 0.002$ &                                          $ 6.467$ &                         $ 0.135$ &                      $ 0.871$ \\
            \rowcolor{LIGHTGREY} France &        Capra hircus &               $ 0.004$ &              $ 0.014$ &              $ 0.002$ &                                          $ 6.407$ &                         $ 0.132$ &                      $ 0.845$ \\
            \rowcolor{LIGHTGREY} Iran (C. aegagrus) &        Capra hircus &               $ 0.004$ &              $ 0.014$ &              $ 0.002$ &                                          $ 5.993$ &                         $ 0.129$ &                      $ 0.772$ \\
            \rowcolor{LIGHTGREY} Iran &        Capra hircus &               $ 0.004$ &              $ 0.014$ &              $ 0.004$ &                                          $ 3.656$ &                         $ 0.138$ &                      $ 0.506$ \\
            \rowcolor{LIGHTGREY} Italy &        Capra hircus &               $ 0.004$ &              $ 0.014$ &              $ 0.002$ &                                          $ 6.216$ &                         $ 0.127$ &                      $ 0.787$ \\
            \rowcolor{LIGHTGREY} Morocco &        Capra hircus &               $ 0.004$ &              $ 0.014$ &              $ 0.002$ &                                          $ 6.019$ &                         $ 0.141$ &                      $ 0.848$ \\
            Iran &          Ovis aries &               $ 0.007$ &              $ 0.014$ &              $ 0.024$ &                                          $ 0.588$ &                         $ 0.148$ &                      $ 0.087$ \\
            Iran (O. orientalis) &          Ovis aries &               $ 0.009$ &              $ 0.014$ &              $ 0.024$ &                                          $ 0.569$ &                         $ 0.143$ &                      $ 0.081$ \\
            Iran (O. vignei) &          Ovis aries &               $ 0.007$ &              $ 0.014$ &              $ 0.024$ &                                          $ 0.587$ &                         $ 0.127$ &                      $ 0.075$ \\
            Various &          Ovis aries &               $ 0.008$ &              $ 0.014$ &              $ 0.024$ &                                          $ 0.588$ &                         $ 0.146$ &                      $ 0.086$ \\
            Morocco &          Ovis aries &               $ 0.008$ &              $ 0.014$ &              $ 0.025$ &                                          $ 0.564$ &                         $ 0.145$ &                      $ 0.082$ \\
            \rowcolor{LIGHTGREY} Barbados & Chlorocebus sabaeus &               $ 0.004$ &              $ 0.012$ &              $ 0.002$ &                                          $ 5.689$ &                         $ 0.170$ &                      $ 0.968$ \\
            \rowcolor{LIGHTGREY} Central Afr. Rep. & Chlorocebus sabaeus &               $ 0.006$ &              $ 0.012$ &              $ 0.005$ &                                          $ 2.142$ &                         $ 0.156$ &                      $ 0.334$ \\
            \rowcolor{LIGHTGREY} Ethiopia & Chlorocebus sabaeus &               $ 0.005$ &              $ 0.012$ &              $ 0.002$ &                                          $ 5.910$ &                         $ 0.160$ &                      $ 0.945$ \\
            \rowcolor{LIGHTGREY} Gambia & Chlorocebus sabaeus &               $ 0.005$ &              $ 0.012$ &              $ 0.006$ &                                          $ 1.850$ &                         $ 0.157$ &                      $ 0.290$ \\
            \rowcolor{LIGHTGREY} Kenya & Chlorocebus sabaeus &               $ 0.005$ &              $ 0.012$ &              $ 0.004$ &                                          $ 2.802$ &                         $ 0.163$ &                      $ 0.456$ \\
            \rowcolor{LIGHTGREY} Nevis & Chlorocebus sabaeus &               $ 0.003$ &              $ 0.012$ &              $ 0.002$ &                                          $ 5.396$ &                         $ 0.169$ &                      $ 0.910$ \\
            \rowcolor{LIGHTGREY} South Africa & Chlorocebus sabaeus &               $ 0.006$ &              $ 0.012$ &              $ 0.005$ &                                          $ 2.154$ &                         $ 0.153$ &                      $ 0.329$ \\
            \rowcolor{LIGHTGREY} Saint Kitts & Chlorocebus sabaeus &               $ 0.004$ &              $ 0.012$ &              $ 0.002$ &                                          $ 5.094$ &                         $ 0.164$ &                      $ 0.834$ \\
            \rowcolor{LIGHTGREY} Zambia & Chlorocebus sabaeus &               $ 0.006$ &              $ 0.012$ &              $ 0.006$ &                                          $ 1.952$ &                         $ 0.153$ &                      $ 0.298$ \\
            African &        Homo sapiens &               $ 0.002$ &              $ 0.012$ &              $ 0.014$ &                                          $ 0.869$ &                         $ 0.141$ &                      $ 0.123$ \\
            Admixed American &        Homo sapiens &               $ 0.002$ &              $ 0.012$ &              $ 0.010$ &                                          $ 1.130$ &                         $ 0.161$ &                      $ 0.182$ \\
            East Asian &        Homo sapiens &               $ 0.002$ &              $ 0.012$ &              $ 0.007$ &                                          $ 1.602$ &                         $ 0.162$ &                      $ 0.259$ \\
            European &        Homo sapiens &               $ 0.002$ &              $ 0.012$ &              $ 0.009$ &                                          $ 1.336$ &                         $ 0.163$ &                      $ 0.217$ \\
            South Asian &        Homo sapiens &               $ 0.002$ &              $ 0.012$ &              $ 0.009$ &                                          $ 1.342$ &                         $ 0.163$ &                      $ 0.219$ \\
        \end{longtable*}
    \end{center}
    \begin{itemize}
        \item Watterson's $\thetaSyn$ is the observed genetic diversity calculated for synonymous changes.
        \item $\proba [ \SphyBen ]$ (eq.~5) is the probability for a new mutation a beneficial back-mutation.
        These mutations have a selection coefficient predicted at the phylogenetic-scale larger than 1, thus toward a more fit amino-acid.
        \item $\proba [ \SpopBen ]$ (eq.~15) is the probability for a mutation to be beneficial.
        These mutations have a selection coefficient at the population-scale larger than 1.
        \item $\proba [ \SpopBen \given \SphyBen]$ (eq.~12) is the probability for a mutation to be beneficial at the population scale, given it is predicted to be a beneficial back-mutation at the phylogenetic scale (the \textit{precision}).
        \item $\proba [ \SphyBen \given \SpopBen]$ (eq.~14) is the probability for a mutation to a beneficial back-mutations, given it is beneficial (the \textit{recall}).
        This probability is obtained using Bayes' formula.
    \end{itemize}
    \newpage

    \subsection{Precision and recall}
    \captionof{table}{Precision and recall}
    \begin{center}
        \begin{adjustbox}{width=\textwidth}
            \begin{tabular}{||l|l|r||r|r||r|r||r|r||}
                \toprule
                \multicolumn{3}{||c||}{} &
                \multicolumn{2}{c||}{\shortstack{\textbf{Deleterious mutations} \\ $\bm{\SpopDel \coloneqq \Spop<-1}$ \\ $\bm{\SphyDel \coloneqq \Sphy<-1}$}} &
                \multicolumn{2}{c||}{\shortstack{\textbf{Nearly-neutral mutations} \\ $\bm{\SpopNeu \coloneqq -1<\Spop<1}$ \\ $\bm{\SphyNeu \coloneqq -1<\Sphy<1}$}} &
                \multicolumn{2}{c||}{\shortstack{\textbf{Beneficial mutations} \\ $\bm{\SpopBen \coloneqq \Spop>1}$ \\ $\bm{\SphyBen \coloneqq \Sphy>1}$}}
                \\ \hline
                \textbf{Population} &
                \textbf{Species} &
                $\bm{\thetaSyn}$ &
                \makecell{\textbf{Precision} \\ $\bm{\proba [\SpopDel \given \SphyDel]}$} &
                \makecell{\textbf{Recall} \\ $\bm{\proba [\SphyDel \given \SpopDel]}$}           &
                \makecell{\textbf{Precision}      \\ $\bm{\proba [\SpopNeu \given \SphyNeu]}$}                                    &
                \makecell{\textbf{Recall}          \\ $\bm{\proba [\SphyNeu \given \SpopNeu]}$}                                  &
                \makecell{\textbf{Precision}          \\ $\bm{\proba [\SpopBen \given \SphyBen]}$}          &
                \makecell{\textbf{Recall}        \\ $\bm{\proba [\SphyBen \given \SpopBen]}$}
                \\   \midrule
                \rowcolor{LIGHTGREY} Equus c. & Equus caballus        & $ 0.002$ & $ 0.895$ & $ 0.969$ & $ 0.645$ & $ 0.315$ & $ 0.141$ & $ 0.316$ \\
                Iran               & Bos taurus        & $ 0.002$ & $ 0.878$ & $ 0.968$ & $ 0.632$ & $ 0.286$ & $ 0.136$ & $ 0.298$ \\
                Uganda              & Bos taurus        & $ 0.005$ & $ 0.951$ & $ 0.963$ & $ 0.597$ & $ 0.508$ & $ 0.138$ & $ 0.162$ \\
                \rowcolor{LIGHTGREY} Australia            & Capra hircus        & $ 0.003$ & $ 0.895$ & $ 0.981$ & $ 0.830$ & $ 0.379$ & $ 0.135$ & $ 0.871$ \\
                \rowcolor{LIGHTGREY} France                                    & Capra hircus          & $ 0.004$ & $ 0.896$ & $ 0.978$ & $ 0.801$ & $ 0.373$ & $ 0.132$ & $ 0.845$ \\
                \rowcolor{LIGHTGREY} Iran (C. aegagrus)                    & Capra hircus          & $ 0.004$ & $ 0.913$ & $ 0.971$ & $ 0.711$ & $ 0.383$ & $ 0.129$ & $ 0.772$ \\
                \rowcolor{LIGHTGREY} Iran                        & Capra hircus          & $ 0.004$ & $ 0.949$ & $ 0.968$ & $ 0.655$ & $ 0.481$ & $ 0.138$ & $ 0.506$ \\
                \rowcolor{LIGHTGREY} Italy                                 & Capra hircus          & $ 0.004$ & $ 0.894$ & $ 0.979$ & $ 0.814$ & $ 0.373$ & $ 0.127$ & $ 0.787$ \\
                \rowcolor{LIGHTGREY} Morocco                                 & Capra hircus          & $ 0.004$ & $ 0.951$ & $ 0.972$ & $ 0.709$ & $ 0.504$ & $ 0.141$ & $ 0.848$ \\
                Iran           & Ovis aries & $ 0.007$ & $ 0.980$ & $ 0.956$ & $ 0.456$ & $ 0.799$ & $ 0.148$ & $ 0.087$ \\
                Iran (O. orientalis)  & Ovis aries & $ 0.009$ & $ 0.979$ & $ 0.955$ & $ 0.454$ & $ 0.798$ & $ 0.143$ & $ 0.081$ \\
                Iran (O. vignei)           & Ovis aries & $ 0.007$ & $ 0.974$ & $ 0.957$ & $ 0.483$ & $ 0.729$ & $ 0.127$ & $ 0.075$ \\
                Various             & Ovis aries & $ 0.008$ & $ 0.979$ & $ 0.958$ & $ 0.500$ & $ 0.813$ & $ 0.146$ & $ 0.086$ \\
                Morocco              & Ovis aries & $ 0.008$ & $ 0.980$ & $ 0.952$ & $ 0.392$ & $ 0.774$ & $ 0.145$ & $ 0.082$ \\
                \rowcolor{LIGHTGREY} Barbados              & Chlorocebus sabaeus & $ 0.004$ & $ 0.880$ & $ 0.984$ & $ 0.854$ & $ 0.354$ & $ 0.170$ & $ 0.968$ \\
                \rowcolor{LIGHTGREY} Central Afr. Rep.       & Chlorocebus sabaeus & $ 0.006$ & $ 0.932$ & $ 0.968$ & $ 0.634$ & $ 0.413$ & $ 0.156$ & $ 0.334$ \\
                \rowcolor{LIGHTGREY} Ethiopia        & Chlorocebus sabaeus & $ 0.005$ & $ 0.877$ & $ 0.980$ & $ 0.810$ & $ 0.337$ & $ 0.160$ & $ 0.945$ \\
                \rowcolor{LIGHTGREY} Gambia             & Chlorocebus sabaeus & $ 0.005$ & $ 0.947$ & $ 0.977$ & $ 0.745$ & $ 0.518$ & $ 0.157$ & $ 0.290$ \\
                \rowcolor{LIGHTGREY} Kenya                                 & Chlorocebus sabaeus        & $ 0.005$ & $ 0.903$ & $ 0.969$ & $ 0.664$ & $ 0.343$ & $ 0.163$ & $ 0.456$ \\
                \rowcolor{LIGHTGREY} Nevis                        & Chlorocebus sabaeus        & $ 0.003$ & $ 0.880$ & $ 0.984$ & $ 0.861$ & $ 0.357$ & $ 0.169$ & $ 0.910$ \\
                \rowcolor{LIGHTGREY} South Africa                              & Chlorocebus sabaeus        & $ 0.006$ & $ 0.931$ & $ 0.967$ & $ 0.615$ & $ 0.401$ & $ 0.153$ & $ 0.329$ \\
                \rowcolor{LIGHTGREY} Saint Kitts                                & Chlorocebus sabaeus        & $ 0.004$ & $ 0.883$ & $ 0.982$ & $ 0.834$ & $ 0.355$ & $ 0.164$ & $ 0.834$ \\
                \rowcolor{LIGHTGREY} Zambia                             & Chlorocebus sabaeus        & $ 0.006$ & $ 0.896$ & $ 0.969$ & $ 0.659$ & $ 0.332$ & $ 0.153$ & $ 0.298$ \\
                African                             & Homo sapiens        & $ 0.002$ & $ 0.946$ & $ 0.971$ & $ 0.630$ & $ 0.486$ & $ 0.141$ & $ 0.123$ \\
                Admixed American                             & Homo sapiens        & $ 0.002$ & $ 0.929$ & $ 0.972$ & $ 0.636$ & $ 0.410$ & $ 0.161$ & $ 0.182$ \\
                East Asian                             & Homo sapiens        & $ 0.002$ & $ 0.937$ & $ 0.973$ & $ 0.647$ & $ 0.429$ & $ 0.162$ & $ 0.259$ \\
                European                             & Homo sapiens        & $ 0.002$ & $ 0.937$ & $ 0.973$ & $ 0.643$ & $ 0.433$ & $ 0.163$ & $ 0.217$ \\
                South Asian                             & Homo sapiens        & $ 0.002$ & $ 0.939$ & $ 0.972$ & $ 0.639$ & $ 0.439$ & $ 0.163$ & $ 0.219$ \\
                \bottomrule
            \end{tabular}
        \end{adjustbox}
    \end{center}
    \begin{itemize}
        \item Watterson's $\thetaSyn$ is the observed genetic diversity calculated for synonymous changes.
        \item \textit{Precision} is the estimation of the selection coefficient at population scale ($\Spop$) given that $\Sphy$ is known.
        \item \textit{Recall} is the estimation of $\Sphy$ given selection coefficient at the population scale ($\Spop$) is known.
        \item \textit{Recall} for beneficial mutations ($\proba [\SphyBen \given \SpopBen]$) is thus the proportion of beneficial back-mutations among all beneficial mutations.
    \end{itemize}

    Altogether, comparing this table to tables 1 and S3, we acknowledge that the exact proportion of beneficial back-mutations among all beneficial ones is dependent on the model used to estimate the DFE.
    However, we can still see that beneficial back-mutations are positively selected compared to neutral and deleterious ones, and this result is not sensitive to the underlying DFE at the population scale.

    %\section{References}
    %\printbibliography[heading=none]

    %    \section{Distribution of fitness effects}
    %    \captionof{figure}{~}
    %    \begin{minipage}{0.49\linewidth}
    %        \includegraphics[width=\linewidth, page=1]{artworks/2bins.Yeast.experimentalDFE.pdf}
    %    \end{minipage}
    %    \begin{minipage}{0.49\linewidth}
    %        \includegraphics[width=\linewidth, page=1]{artworks/2bins.AFR.histogramDFE.pdf}
    %    \end{minipage}
    %    Left panel: distribution of fitness effects experimentally determined in Yeast, data obtained in \textcite{shen_synonymous_2022}.
    %    Right panel:  distribution of fitness effects ($\Sphy$) as predicted with mutation-selection codon models, from reconstructed ancestral genome in \textit{Homo sapiens}.
    %    The DFE predicted with mutation-selection codon models has a negative mode, consistently with experimental DFEs, meaning that most mutations are effectively deleterious.
\end{document}