### 2023-10-14 manuscript/supp-figure3.tex

%TC:ignore
\documentclass{article}

\usepackage{xcolor, colortbl}
\definecolor{BLUELINK}{HTML}{0645AD}
\definecolor{DARKBLUELINK}{HTML}{0B0080}
\PassOptionsToPackage{hyphens}{url}
\usepackage[colorlinks=false]{hyperref}
% for linking between references, figures, TOC, etc in the pdf document
\hypersetup{colorlinks,
    linkcolor=DARKBLUELINK,
    anchorcolor=DARKBLUELINK,
    citecolor=DARKBLUELINK,
    filecolor=DARKBLUELINK,
    menucolor=DARKBLUELINK,
    urlcolor=BLUELINK
} % Color citation links in purple
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{naturalnames}{hyperref}

\usepackage[margin=60pt]{geometry}
\usepackage{amssymb,amsfonts,amsmath,amsthm,mathtools}
\usepackage{lmodern}
\usepackage{bm,bbold}
\usepackage{verbatim}
\usepackage{float}
\usepackage{listings, enumerate, enumitem}
\usepackage[export]{adjustbox}
\usepackage{tabu}
\usepackage{longtable}
\tabulinesep=0.6mm
\newcommand\cellwidth{\TX@col@width}
\usepackage{hhline}
\setlength{\arrayrulewidth}{1.2pt}
\usepackage{multicol,multirow,array}
\usepackage{etoolbox}
\AtBeginEnvironment{tabu}{\footnotesize}
\usepackage{booktabs}
\usepackage{graphicx}
\pdfinclusioncopyfonts=1

\newcommand{\UniDimArray}[1]{\bm{#1}}
\newcommand{\der}{\text{d}}
\newcommand{\e}{\text{e}}
\newcommand{\Ne}{N_{\text{e}}}
\newcommand{\proba}{\mathbb{P}}
\newcommand{\pfix}{\proba_{\text{fix}}}
\newcommand{\dn}{d_N}
\newcommand{\ds}{d_S}
\newcommand{\dnds}{\dn / \ds}
\newcommand{\Sphy}{S_{0}}
\newcommand{\SphyDel}{\mathcal{D}_0}
\newcommand{\SphyNeu}{\mathcal{N}_0}
\newcommand{\SphyBen}{\mathcal{B}_0}
\newcommand{\Sphyclass}{x}
\newcommand{\SphyclassAlt}{y}
\newcommand{\given}{\mid}
\newcommand{\Spop}{S}
\newcommand{\SpopDel}{\mathcal{D}}
\newcommand{\SpopNeu}{\mathcal{N}}
\newcommand{\SpopBen}{\mathcal{B}}
\newcommand{\ProbaPopDel}{\proba [ \SpopDel]}
\newcommand{\ProbaPopNeu}{\proba [ \SpopNeu ]}
\newcommand{\ProbaPopBen}{\proba [ \SpopBen ]}
\newcommand{\AdvMean}{\beta_b}
\newcommand{\DelMean}{\beta_d}
\newcommand{\thetaSyn}{\theta_{\text{S}}}

\title{Replicability of figure 3 across populations}
\date{~}
\author{~}

\begin{document}
    \maketitle
    \tableofcontents

    \newpage

    (Top-left) Distribution of scaled selection coefficients ($\Sphy$), predicted for all possible non-synonymous DNA mutations away from the ancestral exome.
    Mutations are divided into three classes of selection: deleterious ($\SphyDel$), nearly-neutral ($\SphyNeu$) and beneficial ($\SphyBen$, supposedly beneficial back-mutations)

    (Top-right) Distribution of scaled selection coefficients ($\Sphy$) for all observed substitutions along the terminal branch of the tree.
    If there are fewer substitutions than expected, this class is thus undergoing purifying selection, as is the case for $\SphyDel$.

    (Bottom-left) The site-frequency spectrum (SFS) for a random sample of 16 alleles (means in solid lines and standard deviations in color shades) for each class of selection and for synonymous mutations, supposedly neutral (black).
    The SFS represents the proportion of mutations (y-axis) with a given number of derived alleles in the population (x-axis).
    At high frequencies, deleterious mutations are underrepresented.

    (Bottom-right) Proportion of beneficial $\ProbaPopDel$, nearly-neutral $\ProbaPopNeu$, and deleterious mutations $\ProbaPopBen$ estimated at the population scale for each class of selection at the phylogenetic scale.
    Proportions depicted here are not weighted by their mutational opportunities.
    \section{Bos taurus}

    \subsection{Iran}

    \begin{minipage}{0.49\linewidth}
        \includegraphics[width=\linewidth, page=1]{/Users/tlatrille/Documents/SelCoeff/data_processed/opportunities_bins3Mask0.9/DFE.Bos_taurus.IRBT.MutSel.pdf}
    \end{minipage}
    \begin{minipage}{0.49\linewidth}
        \includegraphics[width=\linewidth, page=1]{/Users/tlatrille/Documents/SelCoeff/data_processed/ancestral_bins3Mask0.9/Bos_taurus.IRBT.pdf}
    \end{minipage}
    \\
    \begin{minipage}{0.49\linewidth}
        \includegraphics[width=\linewidth, page=1]{/Users/tlatrille/Documents/SelCoeff/experiments/3bins-mC-nodiv/analysis/Bos_taurus.IRBT.MutSel-sfs.normalize.pdf}
    \end{minipage}
    \begin{minipage}{0.49\linewidth}
        \includegraphics[width=\linewidth, page=1]{/Users/tlatrille/Documents/SelCoeff/experiments/3bins-mC-nodiv/analysis/Bos_taurus.IRBT.MutSel.polyDFE_C.pdf}
    \end{minipage}
    \\
    \subsection{Uganda}

    \begin{minipage}{0.49\linewidth}
        \includegraphics[width=\linewidth, page=1]{/Users/tlatrille/Documents/SelCoeff/data_processed/opportunities_bins3Mask0.9/DFE.Bos_taurus.UGBT.MutSel.pdf}
    \end{minipage}
    \begin{minipage}{0.49\linewidth}
        \includegraphics[width=\linewidth, page=1]{/Users/tlatrille/Documents/SelCoeff/data_processed/ancestral_bins3Mask0.9/Bos_taurus.UGBT.pdf}
    \end{minipage}
    \\
    \begin{minipage}{0.49\linewidth}
        \includegraphics[width=\linewidth, page=1]{/Users/tlatrille/Documents/SelCoeff/experiments/3bins-mC-nodiv/analysis/Bos_taurus.UGBT.MutSel-sfs.normalize.pdf}
    \end{minipage}
    \begin{minipage}{0.49\linewidth}
        \includegraphics[width=\linewidth, page=1]{/Users/tlatrille/Documents/SelCoeff/experiments/3bins-mC-nodiv/analysis/Bos_taurus.UGBT.MutSel.polyDFE_C.pdf}
    \end{minipage}
    \\
    \section{Capra}

    \subsection{Australia}

    \begin{minipage}{0.49\linewidth}
        \includegraphics[width=\linewidth, page=1]{/Users/tlatrille/Documents/SelCoeff/data_processed/opportunities_bins3Mask0.9/DFE.Capra_hircus.AUCH.MutSel.pdf}
    \end{minipage}
    \begin{minipage}{0.49\linewidth}
        \includegraphics[width=\linewidth, page=1]{/Users/tlatrille/Documents/SelCoeff/data_processed/ancestral_bins3Mask0.9/Capra_hircus.AUCH.pdf}
    \end{minipage}
    \\
    \begin{minipage}{0.49\linewidth}
        \includegraphics[width=\linewidth, page=1]{/Users/tlatrille/Documents/SelCoeff/experiments/3bins-mC-nodiv/analysis/Capra_hircus.AUCH.MutSel-sfs.normalize.pdf}
    \end{minipage}
    \begin{minipage}{0.49\linewidth}
        \includegraphics[width=\linewidth, page=1]{/Users/tlatrille/Documents/SelCoeff/experiments/3bins-mC-nodiv/analysis/Capra_hircus.AUCH.MutSel.polyDFE_C.pdf}
    \end{minipage}
    \\
    \subsection{France}

    \begin{minipage}{0.49\linewidth}
        \includegraphics[width=\linewidth, page=1]{/Users/tlatrille/Documents/SelCoeff/data_processed/opportunities_bins3Mask0.9/DFE.Capra_hircus.FRCH.MutSel.pdf}
    \end{minipage}
    \begin{minipage}{0.49\linewidth}
        \includegraphics[width=\linewidth, page=1]{/Users/tlatrille/Documents/SelCoeff/data_processed/ancestral_bins3Mask0.9/Capra_hircus.FRCH.pdf}
    \end{minipage}
    \\
    \begin{minipage}{0.49\linewidth}
        \includegraphics[width=\linewidth, page=1]{/Users/tlatrille/Documents/SelCoeff/experiments/3bins-mC-nodiv/analysis/Capra_hircus.FRCH.MutSel-sfs.normalize.pdf}
    \end{minipage}
    \begin{minipage}{0.49\linewidth}
        \includegraphics[width=\linewidth, page=1]{/Users/tlatrille/Documents/SelCoeff/experiments/3bins-mC-nodiv/analysis/Capra_hircus.FRCH.MutSel.polyDFE_C.pdf}
    \end{minipage}
    \\
    \subsection{Iran (C. aegagrus)}

    \begin{minipage}{0.49\linewidth}
        \includegraphics[width=\linewidth, page=1]{/Users/tlatrille/Documents/SelCoeff/data_processed/opportunities_bins3Mask0.9/DFE.Capra_hircus.IRCA.MutSel.pdf}
    \end{minipage}
    \begin{minipage}{0.49\linewidth}
        \includegraphics[width=\linewidth, page=1]{/Users/tlatrille/Documents/SelCoeff/data_processed/ancestral_bins3Mask0.9/Capra_hircus.IRCA.pdf}
    \end{minipage}
    \\
    \begin{minipage}{0.49\linewidth}
        \includegraphics[width=\linewidth, page=1]{/Users/tlatrille/Documents/SelCoeff/experiments/3bins-mC-nodiv/analysis/Capra_hircus.IRCA.MutSel-sfs.normalize.pdf}
    \end{minipage}
    \begin{minipage}{0.49\linewidth}
        \includegraphics[width=\linewidth, page=1]{/Users/tlatrille/Documents/SelCoeff/experiments/3bins-mC-nodiv/analysis/Capra_hircus.IRCA.MutSel.polyDFE_C.pdf}
    \end{minipage}
    \\
    \subsection{Iran}

    \begin{minipage}{0.49\linewidth}
        \includegraphics[width=\linewidth, page=1]{/Users/tlatrille/Documents/SelCoeff/data_processed/opportunities_bins3Mask0.9/DFE.Capra_hircus.IRCH.MutSel.pdf}
    \end{minipage}
    \begin{minipage}{0.49\linewidth}
        \includegraphics[width=\linewidth, page=1]{/Users/tlatrille/Documents/SelCoeff/data_processed/ancestral_bins3Mask0.9/Capra_hircus.IRCH.pdf}
    \end{minipage}
    \\
    \begin{minipage}{0.49\linewidth}
        \includegraphics[width=\linewidth, page=1]{/Users/tlatrille/Documents/SelCoeff/experiments/3bins-mC-nodiv/analysis/Capra_hircus.IRCH.MutSel-sfs.normalize.pdf}
    \end{minipage}
    \begin{minipage}{0.49\linewidth}
        \includegraphics[width=\linewidth, page=1]{/Users/tlatrille/Documents/SelCoeff/experiments/3bins-mC-nodiv/analysis/Capra_hircus.IRCH.MutSel.polyDFE_C.pdf}
    \end{minipage}
    \\
    \subsection{Italy}

    \begin{minipage}{0.49\linewidth}
        \includegraphics[width=\linewidth, page=1]{/Users/tlatrille/Documents/SelCoeff/data_processed/opportunities_bins3Mask0.9/DFE.Capra_hircus.ITCH.MutSel.pdf}
    \end{minipage}
    \begin{minipage}{0.49\linewidth}
        \includegraphics[width=\linewidth, page=1]{/Users/tlatrille/Documents/SelCoeff/data_processed/ancestral_bins3Mask0.9/Capra_hircus.ITCH.pdf}
    \end{minipage}
    \\
    \begin{minipage}{0.49\linewidth}
        \includegraphics[width=\linewidth, page=1]{/Users/tlatrille/Documents/SelCoeff/experiments/3bins-mC-nodiv/analysis/Capra_hircus.ITCH.MutSel-sfs.normalize.pdf}
    \end{minipage}
    \begin{minipage}{0.49\linewidth}
        \includegraphics[width=\linewidth, page=1]{/Users/tlatrille/Documents/SelCoeff/experiments/3bins-mC-nodiv/analysis/Capra_hircus.ITCH.MutSel.polyDFE_C.pdf}
    \end{minipage}
    \\
    \subsection{Morocco}

    \begin{minipage}{0.49\linewidth}
        \includegraphics[width=\linewidth, page=1]{/Users/tlatrille/Documents/SelCoeff/data_processed/opportunities_bins3Mask0.9/DFE.Capra_hircus.MOCH.MutSel.pdf}
    \end{minipage}
    \begin{minipage}{0.49\linewidth}
        \includegraphics[width=\linewidth, page=1]{/Users/tlatrille/Documents/SelCoeff/data_processed/ancestral_bins3Mask0.9/Capra_hircus.MOCH.pdf}
    \end{minipage}
    \\
    \begin{minipage}{0.49\linewidth}
        \includegraphics[width=\linewidth, page=1]{/Users/tlatrille/Documents/SelCoeff/experiments/3bins-mC-nodiv/analysis/Capra_hircus.MOCH.MutSel-sfs.normalize.pdf}
    \end{minipage}
    \begin{minipage}{0.49\linewidth}
        \includegraphics[width=\linewidth, page=1]{/Users/tlatrille/Documents/SelCoeff/experiments/3bins-mC-nodiv/analysis/Capra_hircus.MOCH.MutSel.polyDFE_C.pdf}
    \end{minipage}
    \\
    \section{Chlorocebus sabaeus}

    \subsection{Barbados}

    \begin{minipage}{0.49\linewidth}
        \includegraphics[width=\linewidth, page=1]{/Users/tlatrille/Documents/SelCoeff/data_processed/opportunities_bins3Mask0.9/DFE.Chlorocebus_sabaeus.Barbados.MutSel.pdf}
    \end{minipage}
    \begin{minipage}{0.49\linewidth}
        \includegraphics[width=\linewidth, page=1]{/Users/tlatrille/Documents/SelCoeff/data_processed/ancestral_bins3Mask0.9/Chlorocebus_sabaeus.Barbados.pdf}
    \end{minipage}
    \\
    \begin{minipage}{0.49\linewidth}
        \includegraphics[width=\linewidth, page=1]{/Users/tlatrille/Documents/SelCoeff/experiments/3bins-mC-nodiv/analysis/Chlorocebus_sabaeus.Barbados.MutSel-sfs.normalize.pdf}
    \end{minipage}
    \begin{minipage}{0.49\linewidth}
        \includegraphics[width=\linewidth, page=1]{/Users/tlatrille/Documents/SelCoeff/experiments/3bins-mC-nodiv/analysis/Chlorocebus_sabaeus.Barbados.MutSel.polyDFE_C.pdf}
    \end{minipage}
    \\
    \subsection{Central Afr. Rep.}

    \begin{minipage}{0.49\linewidth}
        \includegraphics[width=\linewidth, page=1]{/Users/tlatrille/Documents/SelCoeff/data_processed/opportunities_bins3Mask0.9/DFE.Chlorocebus_sabaeus.Central_African_Republic.MutSel.pdf}
    \end{minipage}
    \begin{minipage}{0.49\linewidth}
        \includegraphics[width=\linewidth, page=1]{/Users/tlatrille/Documents/SelCoeff/data_processed/ancestral_bins3Mask0.9/Chlorocebus_sabaeus.Central_African_Republic.pdf}
    \end{minipage}
    \\
    \begin{minipage}{0.49\linewidth}
        \includegraphics[width=\linewidth, page=1]{/Users/tlatrille/Documents/SelCoeff/experiments/3bins-mC-nodiv/analysis/Chlorocebus_sabaeus.Central_African_Republic.MutSel-sfs.normalize.pdf}
    \end{minipage}
    \begin{minipage}{0.49\linewidth}
        \includegraphics[width=\linewidth, page=1]{/Users/tlatrille/Documents/SelCoeff/experiments/3bins-mC-nodiv/analysis/Chlorocebus_sabaeus.Central_African_Republic.MutSel.polyDFE_C.pdf}
    \end{minipage}
    \\
    \subsection{Ethiopia}

    \begin{minipage}{0.49\linewidth}
        \includegraphics[width=\linewidth, page=1]{/Users/tlatrille/Documents/SelCoeff/data_processed/opportunities_bins3Mask0.9/DFE.Chlorocebus_sabaeus.Ethiopia.MutSel.pdf}
    \end{minipage}
    \begin{minipage}{0.49\linewidth}
        \includegraphics[width=\linewidth, page=1]{/Users/tlatrille/Documents/SelCoeff/data_processed/ancestral_bins3Mask0.9/Chlorocebus_sabaeus.Ethiopia.pdf}
    \end{minipage}
    \\
    \begin{minipage}{0.49\linewidth}
        \includegraphics[width=\linewidth, page=1]{/Users/tlatrille/Documents/SelCoeff/experiments/3bins-mC-nodiv/analysis/Chlorocebus_sabaeus.Ethiopia.MutSel-sfs.normalize.pdf}
    \end{minipage}
    \begin{minipage}{0.49\linewidth}
        \includegraphics[width=\linewidth, page=1]{/Users/tlatrille/Documents/SelCoeff/experiments/3bins-mC-nodiv/analysis/Chlorocebus_sabaeus.Ethiopia.MutSel.polyDFE_C.pdf}
    \end{minipage}
    \\
    \subsection{Gambia}

    \begin{minipage}{0.49\linewidth}
        \includegraphics[width=\linewidth, page=1]{/Users/tlatrille/Documents/SelCoeff/data_processed/opportunities_bins3Mask0.9/DFE.Chlorocebus_sabaeus.Gambia.MutSel.pdf}
    \end{minipage}
    \begin{minipage}{0.49\linewidth}
        \includegraphics[width=\linewidth, page=1]{/Users/tlatrille/Documents/SelCoeff/data_processed/ancestral_bins3Mask0.9/Chlorocebus_sabaeus.Gambia.pdf}
    \end{minipage}
    \\
    \begin{minipage}{0.49\linewidth}
        \includegraphics[width=\linewidth, page=1]{/Users/tlatrille/Documents/SelCoeff/experiments/3bins-mC-nodiv/analysis/Chlorocebus_sabaeus.Gambia.MutSel-sfs.normalize.pdf}
    \end{minipage}
    \begin{minipage}{0.49\linewidth}
        \includegraphics[width=\linewidth, page=1]{/Users/tlatrille/Documents/SelCoeff/experiments/3bins-mC-nodiv/analysis/Chlorocebus_sabaeus.Gambia.MutSel.polyDFE_C.pdf}
    \end{minipage}
    \\
    \subsection{Kenya}

    \begin{minipage}{0.49\linewidth}
        \includegraphics[width=\linewidth, page=1]{/Users/tlatrille/Documents/SelCoeff/data_processed/opportunities_bins3Mask0.9/DFE.Chlorocebus_sabaeus.Kenya.MutSel.pdf}
    \end{minipage}
    \begin{minipage}{0.49\linewidth}
        \includegraphics[width=\linewidth, page=1]{/Users/tlatrille/Documents/SelCoeff/data_processed/ancestral_bins3Mask0.9/Chlorocebus_sabaeus.Kenya.pdf}
    \end{minipage}
    \\
    \begin{minipage}{0.49\linewidth}
        \includegraphics[width=\linewidth, page=1]{/Users/tlatrille/Documents/SelCoeff/experiments/3bins-mC-nodiv/analysis/Chlorocebus_sabaeus.Kenya.MutSel-sfs.normalize.pdf}
    \end{minipage}
    \begin{minipage}{0.49\linewidth}
        \includegraphics[width=\linewidth, page=1]{/Users/tlatrille/Documents/SelCoeff/experiments/3bins-mC-nodiv/analysis/Chlorocebus_sabaeus.Kenya.MutSel.polyDFE_C.pdf}
    \end{minipage}
    \\
    \subsection{Nevis}

    \begin{minipage}{0.49\linewidth}
        \includegraphics[width=\linewidth, page=1]{/Users/tlatrille/Documents/SelCoeff/data_processed/opportunities_bins3Mask0.9/DFE.Chlorocebus_sabaeus.Nevis.MutSel.pdf}
    \end{minipage}
    \begin{minipage}{0.49\linewidth}
        \includegraphics[width=\linewidth, page=1]{/Users/tlatrille/Documents/SelCoeff/data_processed/ancestral_bins3Mask0.9/Chlorocebus_sabaeus.Nevis.pdf}
    \end{minipage}
    \\
    \begin{minipage}{0.49\linewidth}
        \includegraphics[width=\linewidth, page=1]{/Users/tlatrille/Documents/SelCoeff/experiments/3bins-mC-nodiv/analysis/Chlorocebus_sabaeus.Nevis.MutSel-sfs.normalize.pdf}
    \end{minipage}
    \begin{minipage}{0.49\linewidth}
        \includegraphics[width=\linewidth, page=1]{/Users/tlatrille/Documents/SelCoeff/experiments/3bins-mC-nodiv/analysis/Chlorocebus_sabaeus.Nevis.MutSel.polyDFE_C.pdf}
    \end{minipage}
    \\
    \subsection{Saint Kitts}

    \begin{minipage}{0.49\linewidth}
        \includegraphics[width=\linewidth, page=1]{/Users/tlatrille/Documents/SelCoeff/data_processed/opportunities_bins3Mask0.9/DFE.Chlorocebus_sabaeus.Saint_Kitts.MutSel.pdf}
    \end{minipage}
    \begin{minipage}{0.49\linewidth}
        \includegraphics[width=\linewidth, page=1]{/Users/tlatrille/Documents/SelCoeff/data_processed/ancestral_bins3Mask0.9/Chlorocebus_sabaeus.Saint_Kitts.pdf}
    \end{minipage}
    \\
    \begin{minipage}{0.49\linewidth}
        \includegraphics[width=\linewidth, page=1]{/Users/tlatrille/Documents/SelCoeff/experiments/3bins-mC-nodiv/analysis/Chlorocebus_sabaeus.Saint_Kitts.MutSel-sfs.normalize.pdf}
    \end{minipage}
    \begin{minipage}{0.49\linewidth}
        \includegraphics[width=\linewidth, page=1]{/Users/tlatrille/Documents/SelCoeff/experiments/3bins-mC-nodiv/analysis/Chlorocebus_sabaeus.Saint_Kitts.MutSel.polyDFE_C.pdf}
    \end{minipage}
    \\
    \subsection{South Africa}

    \begin{minipage}{0.49\linewidth}
        \includegraphics[width=\linewidth, page=1]{/Users/tlatrille/Documents/SelCoeff/data_processed/opportunities_bins3Mask0.9/DFE.Chlorocebus_sabaeus.South_Africa.MutSel.pdf}
    \end{minipage}
    \begin{minipage}{0.49\linewidth}
        \includegraphics[width=\linewidth, page=1]{/Users/tlatrille/Documents/SelCoeff/data_processed/ancestral_bins3Mask0.9/Chlorocebus_sabaeus.South_Africa.pdf}
    \end{minipage}
    \\
    \begin{minipage}{0.49\linewidth}
        \includegraphics[width=\linewidth, page=1]{/Users/tlatrille/Documents/SelCoeff/experiments/3bins-mC-nodiv/analysis/Chlorocebus_sabaeus.South_Africa.MutSel-sfs.normalize.pdf}
    \end{minipage}
    \begin{minipage}{0.49\linewidth}
        \includegraphics[width=\linewidth, page=1]{/Users/tlatrille/Documents/SelCoeff/experiments/3bins-mC-nodiv/analysis/Chlorocebus_sabaeus.South_Africa.MutSel.polyDFE_C.pdf}
    \end{minipage}
    \\
    \subsection{Zambia}

    \begin{minipage}{0.49\linewidth}
        \includegraphics[width=\linewidth, page=1]{/Users/tlatrille/Documents/SelCoeff/data_processed/opportunities_bins3Mask0.9/DFE.Chlorocebus_sabaeus.Zambia.MutSel.pdf}
    \end{minipage}
    \begin{minipage}{0.49\linewidth}
        \includegraphics[width=\linewidth, page=1]{/Users/tlatrille/Documents/SelCoeff/data_processed/ancestral_bins3Mask0.9/Chlorocebus_sabaeus.Zambia.pdf}
    \end{minipage}
    \\
    \begin{minipage}{0.49\linewidth}
        \includegraphics[width=\linewidth, page=1]{/Users/tlatrille/Documents/SelCoeff/experiments/3bins-mC-nodiv/analysis/Chlorocebus_sabaeus.Zambia.MutSel-sfs.normalize.pdf}
    \end{minipage}
    \begin{minipage}{0.49\linewidth}
        \includegraphics[width=\linewidth, page=1]{/Users/tlatrille/Documents/SelCoeff/experiments/3bins-mC-nodiv/analysis/Chlorocebus_sabaeus.Zambia.MutSel.polyDFE_C.pdf}
    \end{minipage}
    \\
    \section{Equus caballus}

    \begin{minipage}{0.49\linewidth}
        \includegraphics[width=\linewidth, page=1]{/Users/tlatrille/Documents/SelCoeff/data_processed/opportunities_bins3Mask0.9/DFE.Equus_caballus.up.MutSel.pdf}
    \end{minipage}
    \begin{minipage}{0.49\linewidth}
        \includegraphics[width=\linewidth, page=1]{/Users/tlatrille/Documents/SelCoeff/data_processed/ancestral_bins3Mask0.9/Equus_caballus.up.pdf}
    \end{minipage}
    \\
    \begin{minipage}{0.49\linewidth}
        \includegraphics[width=\linewidth, page=1]{/Users/tlatrille/Documents/SelCoeff/experiments/3bins-mC-nodiv/analysis/Equus_caballus.up.MutSel-sfs.normalize.pdf}
    \end{minipage}
    \begin{minipage}{0.49\linewidth}
        \includegraphics[width=\linewidth, page=1]{/Users/tlatrille/Documents/SelCoeff/experiments/3bins-mC-nodiv/analysis/Equus_caballus.up.MutSel.polyDFE_C.pdf}
    \end{minipage}
    \\
    \section{Homo sapiens}

    \subsection{African}

    \begin{minipage}{0.49\linewidth}
        \includegraphics[width=\linewidth, page=1]{/Users/tlatrille/Documents/SelCoeff/data_processed/opportunities_bins3Mask0.9/DFE.Homo_sapiens.AFR.MutSel.pdf}
    \end{minipage}
    \begin{minipage}{0.49\linewidth}
        \includegraphics[width=\linewidth, page=1]{/Users/tlatrille/Documents/SelCoeff/data_processed/ancestral_bins3Mask0.9/Homo_sapiens.AFR.pdf}
    \end{minipage}
    \\
    \begin{minipage}{0.49\linewidth}
        \includegraphics[width=\linewidth, page=1]{/Users/tlatrille/Documents/SelCoeff/experiments/3bins-mC-nodiv/analysis/Homo_sapiens.AFR.MutSel-sfs.normalize.pdf}
    \end{minipage}
    \begin{minipage}{0.49\linewidth}
        \includegraphics[width=\linewidth, page=1]{/Users/tlatrille/Documents/SelCoeff/experiments/3bins-mC-nodiv/analysis/Homo_sapiens.AFR.MutSel.polyDFE_C.pdf}
    \end{minipage}
    \\
    \subsection{Admixed American}

    \begin{minipage}{0.49\linewidth}
        \includegraphics[width=\linewidth, page=1]{/Users/tlatrille/Documents/SelCoeff/data_processed/opportunities_bins3Mask0.9/DFE.Homo_sapiens.AMR.MutSel.pdf}
    \end{minipage}
    \begin{minipage}{0.49\linewidth}
        \includegraphics[width=\linewidth, page=1]{/Users/tlatrille/Documents/SelCoeff/data_processed/ancestral_bins3Mask0.9/Homo_sapiens.AMR.pdf}
    \end{minipage}
    \\
    \begin{minipage}{0.49\linewidth}
        \includegraphics[width=\linewidth, page=1]{/Users/tlatrille/Documents/SelCoeff/experiments/3bins-mC-nodiv/analysis/Homo_sapiens.AMR.MutSel-sfs.normalize.pdf}
    \end{minipage}
    \begin{minipage}{0.49\linewidth}
        \includegraphics[width=\linewidth, page=1]{/Users/tlatrille/Documents/SelCoeff/experiments/3bins-mC-nodiv/analysis/Homo_sapiens.AMR.MutSel.polyDFE_C.pdf}
    \end{minipage}
    \\
    \subsection{East Asian}

    \begin{minipage}{0.49\linewidth}
        \includegraphics[width=\linewidth, page=1]{/Users/tlatrille/Documents/SelCoeff/data_processed/opportunities_bins3Mask0.9/DFE.Homo_sapiens.EAS.MutSel.pdf}
    \end{minipage}
    \begin{minipage}{0.49\linewidth}
        \includegraphics[width=\linewidth, page=1]{/Users/tlatrille/Documents/SelCoeff/data_processed/ancestral_bins3Mask0.9/Homo_sapiens.EAS.pdf}
    \end{minipage}
    \\
    \begin{minipage}{0.49\linewidth}
        \includegraphics[width=\linewidth, page=1]{/Users/tlatrille/Documents/SelCoeff/experiments/3bins-mC-nodiv/analysis/Homo_sapiens.EAS.MutSel-sfs.normalize.pdf}
    \end{minipage}
    \begin{minipage}{0.49\linewidth}
        \includegraphics[width=\linewidth, page=1]{/Users/tlatrille/Documents/SelCoeff/experiments/3bins-mC-nodiv/analysis/Homo_sapiens.EAS.MutSel.polyDFE_C.pdf}
    \end{minipage}
    \\
    \subsection{European}

    \begin{minipage}{0.49\linewidth}
        \includegraphics[width=\linewidth, page=1]{/Users/tlatrille/Documents/SelCoeff/data_processed/opportunities_bins3Mask0.9/DFE.Homo_sapiens.EUR.MutSel.pdf}
    \end{minipage}
    \begin{minipage}{0.49\linewidth}
        \includegraphics[width=\linewidth, page=1]{/Users/tlatrille/Documents/SelCoeff/data_processed/ancestral_bins3Mask0.9/Homo_sapiens.EUR.pdf}
    \end{minipage}
    \\
    \begin{minipage}{0.49\linewidth}
        \includegraphics[width=\linewidth, page=1]{/Users/tlatrille/Documents/SelCoeff/experiments/3bins-mC-nodiv/analysis/Homo_sapiens.EUR.MutSel-sfs.normalize.pdf}
    \end{minipage}
    \begin{minipage}{0.49\linewidth}
        \includegraphics[width=\linewidth, page=1]{/Users/tlatrille/Documents/SelCoeff/experiments/3bins-mC-nodiv/analysis/Homo_sapiens.EUR.MutSel.polyDFE_C.pdf}
    \end{minipage}
    \\
    \subsection{South Asian}

    \begin{minipage}{0.49\linewidth}
        \includegraphics[width=\linewidth, page=1]{/Users/tlatrille/Documents/SelCoeff/data_processed/opportunities_bins3Mask0.9/DFE.Homo_sapiens.SAS.MutSel.pdf}
    \end{minipage}
    \begin{minipage}{0.49\linewidth}
        \includegraphics[width=\linewidth, page=1]{/Users/tlatrille/Documents/SelCoeff/data_processed/ancestral_bins3Mask0.9/Homo_sapiens.SAS.pdf}
    \end{minipage}
    \\
    \begin{minipage}{0.49\linewidth}
        \includegraphics[width=\linewidth, page=1]{/Users/tlatrille/Documents/SelCoeff/experiments/3bins-mC-nodiv/analysis/Homo_sapiens.SAS.MutSel-sfs.normalize.pdf}
    \end{minipage}
    \begin{minipage}{0.49\linewidth}
        \includegraphics[width=\linewidth, page=1]{/Users/tlatrille/Documents/SelCoeff/experiments/3bins-mC-nodiv/analysis/Homo_sapiens.SAS.MutSel.polyDFE_C.pdf}
    \end{minipage}
    \\
    \section{Ovis}

    \subsection{Iran}

    \begin{minipage}{0.49\linewidth}
        \includegraphics[width=\linewidth, page=1]{/Users/tlatrille/Documents/SelCoeff/data_processed/opportunities_bins3Mask0.9/DFE.Ovis_aries.IROA.MutSel.pdf}
    \end{minipage}
    \begin{minipage}{0.49\linewidth}
        \includegraphics[width=\linewidth, page=1]{/Users/tlatrille/Documents/SelCoeff/data_processed/ancestral_bins3Mask0.9/Ovis_aries.IROA.pdf}
    \end{minipage}
    \\
    \begin{minipage}{0.49\linewidth}
        \includegraphics[width=\linewidth, page=1]{/Users/tlatrille/Documents/SelCoeff/experiments/3bins-mC-nodiv/analysis/Ovis_aries.IROA.MutSel-sfs.normalize.pdf}
    \end{minipage}
    \begin{minipage}{0.49\linewidth}
        \includegraphics[width=\linewidth, page=1]{/Users/tlatrille/Documents/SelCoeff/experiments/3bins-mC-nodiv/analysis/Ovis_aries.IROA.MutSel.polyDFE_C.pdf}
    \end{minipage}
    \\
    \subsection{Iran (O. orientalis)}

    \begin{minipage}{0.49\linewidth}
        \includegraphics[width=\linewidth, page=1]{/Users/tlatrille/Documents/SelCoeff/data_processed/opportunities_bins3Mask0.9/DFE.Ovis_aries.IROO.MutSel.pdf}
    \end{minipage}
    \begin{minipage}{0.49\linewidth}
        \includegraphics[width=\linewidth, page=1]{/Users/tlatrille/Documents/SelCoeff/data_processed/ancestral_bins3Mask0.9/Ovis_aries.IROO.pdf}
    \end{minipage}
    \\
    \begin{minipage}{0.49\linewidth}
        \includegraphics[width=\linewidth, page=1]{/Users/tlatrille/Documents/SelCoeff/experiments/3bins-mC-nodiv/analysis/Ovis_aries.IROO.MutSel-sfs.normalize.pdf}
    \end{minipage}
    \begin{minipage}{0.49\linewidth}
        \includegraphics[width=\linewidth, page=1]{/Users/tlatrille/Documents/SelCoeff/experiments/3bins-mC-nodiv/analysis/Ovis_aries.IROO.MutSel.polyDFE_C.pdf}
    \end{minipage}
    \\
    \subsection{Iran (O. vignei)}

    \begin{minipage}{0.49\linewidth}
        \includegraphics[width=\linewidth, page=1]{/Users/tlatrille/Documents/SelCoeff/data_processed/opportunities_bins3Mask0.9/DFE.Ovis_aries.IROV.MutSel.pdf}
    \end{minipage}
    \begin{minipage}{0.49\linewidth}
        \includegraphics[width=\linewidth, page=1]{/Users/tlatrille/Documents/SelCoeff/data_processed/ancestral_bins3Mask0.9/Ovis_aries.IROV.pdf}
    \end{minipage}
    \\
    \begin{minipage}{0.49\linewidth}
        \includegraphics[width=\linewidth, page=1]{/Users/tlatrille/Documents/SelCoeff/experiments/3bins-mC-nodiv/analysis/Ovis_aries.IROV.MutSel-sfs.normalize.pdf}
    \end{minipage}
    \begin{minipage}{0.49\linewidth}
        \includegraphics[width=\linewidth, page=1]{/Users/tlatrille/Documents/SelCoeff/experiments/3bins-mC-nodiv/analysis/Ovis_aries.IROV.MutSel.polyDFE_C.pdf}
    \end{minipage}
    \\
    \subsection{Various}

    \begin{minipage}{0.49\linewidth}
        \includegraphics[width=\linewidth, page=1]{/Users/tlatrille/Documents/SelCoeff/data_processed/opportunities_bins3Mask0.9/DFE.Ovis_aries.ISGC.MutSel.pdf}
    \end{minipage}
    \begin{minipage}{0.49\linewidth}
        \includegraphics[width=\linewidth, page=1]{/Users/tlatrille/Documents/SelCoeff/data_processed/ancestral_bins3Mask0.9/Ovis_aries.ISGC.pdf}
    \end{minipage}
    \\
    \begin{minipage}{0.49\linewidth}
        \includegraphics[width=\linewidth, page=1]{/Users/tlatrille/Documents/SelCoeff/experiments/3bins-mC-nodiv/analysis/Ovis_aries.ISGC.MutSel-sfs.normalize.pdf}
    \end{minipage}
    \begin{minipage}{0.49\linewidth}
        \includegraphics[width=\linewidth, page=1]{/Users/tlatrille/Documents/SelCoeff/experiments/3bins-mC-nodiv/analysis/Ovis_aries.ISGC.MutSel.polyDFE_C.pdf}
    \end{minipage}
    \\
    \subsection{Morocco}

    \begin{minipage}{0.49\linewidth}
        \includegraphics[width=\linewidth, page=1]{/Users/tlatrille/Documents/SelCoeff/data_processed/opportunities_bins3Mask0.9/DFE.Ovis_aries.MOOA.MutSel.pdf}
    \end{minipage}
    \begin{minipage}{0.49\linewidth}
        \includegraphics[width=\linewidth, page=1]{/Users/tlatrille/Documents/SelCoeff/data_processed/ancestral_bins3Mask0.9/Ovis_aries.MOOA.pdf}
    \end{minipage}
    \\
    \begin{minipage}{0.49\linewidth}
        \includegraphics[width=\linewidth, page=1]{/Users/tlatrille/Documents/SelCoeff/experiments/3bins-mC-nodiv/analysis/Ovis_aries.MOOA.MutSel-sfs.normalize.pdf}
    \end{minipage}
    \begin{minipage}{0.49\linewidth}
        \includegraphics[width=\linewidth, page=1]{/Users/tlatrille/Documents/SelCoeff/experiments/3bins-mC-nodiv/analysis/Ovis_aries.MOOA.MutSel.polyDFE_C.pdf}
    \end{minipage}
    \\

\end{document}