### 2024-03-04 manuscript/supp-figure3.tex

%TC:ignore
\documentclass{article}

\usepackage{xcolor, colortbl}
\definecolor{BLUELINK}{HTML}{0645AD}
\definecolor{DARKBLUELINK}{HTML}{0B0080}
\PassOptionsToPackage{hyphens}{url}
\usepackage[colorlinks=false]{hyperref}
% for linking between references, figures, TOC, etc in the pdf document
\hypersetup{colorlinks,
    linkcolor=DARKBLUELINK,
    anchorcolor=DARKBLUELINK,
    citecolor=DARKBLUELINK,
    filecolor=DARKBLUELINK,
    menucolor=DARKBLUELINK,
    urlcolor=BLUELINK
} % Color citation links in purple
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{naturalnames}{hyperref}

\usepackage[margin=60pt]{geometry}
\usepackage{amssymb,amsfonts,amsmath,amsthm,mathtools}
\usepackage{lmodern}
\usepackage{bm,bbold}
\usepackage{verbatim}
\usepackage{float}
\usepackage{listings, enumerate, enumitem}
\usepackage[export]{adjustbox}
\usepackage{tabu}
\usepackage{longtable}
\tabulinesep=0.6mm
\newcommand\cellwidth{\TX@col@width}
\usepackage{hhline}
\setlength{\arrayrulewidth}{1.2pt}
\usepackage{multicol,multirow,array}
\usepackage{etoolbox}
\AtBeginEnvironment{tabu}{\footnotesize}
\usepackage{booktabs}
\usepackage{graphicx}
\pdfinclusioncopyfonts=1

\title{Replicability across populations (S7.2)}
\date{~}
\author{~}

\begin{document}
    \maketitle
    \tableofcontents
    \newpage

    \section{Bos taurus}

    \subsection{Iran}

    \begin{minipage}{0.32\linewidth}
        \flushleft {\tiny A: Expected DFE for mutations}
        \includegraphics[width=\linewidth, page=1]{/Users/tlatrille/Documents/SelCoeff/data_processed/opportunities_bins3Mask0.9/DFE.Bos_taurus.IRBT.MutSel.pdf}
    \end{minipage}
    \begin{minipage}{0.32\linewidth}
        \flushleft {\tiny B: Expected DFE for substitutions}
        \includegraphics[width=\linewidth, page=1]{/Users/tlatrille/Documents/SelCoeff/data_processed/opportunities_bins3Mask0.9/DFE.Bos_taurus.IRBT.MutSel.Flow.pdf}
    \end{minipage}
    \begin{minipage}{0.32\linewidth}
        \flushleft {\tiny C: Observed DFE for substitutions}
        \includegraphics[width=\linewidth, page=1]{/Users/tlatrille/Documents/SelCoeff/supp/hist/subs.Bos_taurus.IRBT.bins3Mask0.9.pdf}
    \end{minipage}
    \\
    \begin{minipage}{0.32\linewidth}
        \flushleft {\tiny D: Observed DFE for polymorphisms}
        \includegraphics[width=\linewidth, page=1]{/Users/tlatrille/Documents/SelCoeff/supp/hist/poly.Bos_taurus.IRBT.bins3Mask0.9.pdf}
    \end{minipage}
    \begin{minipage}{0.32\linewidth}
        \flushleft {\tiny E: Site frequency spectrum}
        \includegraphics[width=\linewidth, page=1]{/Users/tlatrille/Documents/SelCoeff/experiments/3bins-mC-nodiv/analysis/Bos_taurus.IRBT.MutSel-sfs.normalize.pdf}
    \end{minipage}
    \begin{minipage}{0.32\linewidth}
        \flushleft {\tiny F: $S$ as function of $S_0$ for each class}
        \includegraphics[width=\linewidth, page=1]{/Users/tlatrille/Documents/SelCoeff/experiments/3bins-mC-nodiv/analysis/Bos_taurus.IRBT.MutSel.polyDFE_C.pdf}
    \end{minipage}
    \subsection{Uganda}

    \begin{minipage}{0.32\linewidth}
        \flushleft {\tiny A: Expected DFE for mutations}
        \includegraphics[width=\linewidth, page=1]{/Users/tlatrille/Documents/SelCoeff/data_processed/opportunities_bins3Mask0.9/DFE.Bos_taurus.UGBT.MutSel.pdf}
    \end{minipage}
    \begin{minipage}{0.32\linewidth}
        \flushleft {\tiny B: Expected DFE for substitutions}
        \includegraphics[width=\linewidth, page=1]{/Users/tlatrille/Documents/SelCoeff/data_processed/opportunities_bins3Mask0.9/DFE.Bos_taurus.UGBT.MutSel.Flow.pdf}
    \end{minipage}
    \begin{minipage}{0.32\linewidth}
        \flushleft {\tiny C: Observed DFE for substitutions}
        \includegraphics[width=\linewidth, page=1]{/Users/tlatrille/Documents/SelCoeff/supp/hist/subs.Bos_taurus.UGBT.bins3Mask0.9.pdf}
    \end{minipage}
    \\
    \begin{minipage}{0.32\linewidth}
        \flushleft {\tiny D: Observed DFE for polymorphisms}
        \includegraphics[width=\linewidth, page=1]{/Users/tlatrille/Documents/SelCoeff/supp/hist/poly.Bos_taurus.UGBT.bins3Mask0.9.pdf}
    \end{minipage}
    \begin{minipage}{0.32\linewidth}
        \flushleft {\tiny E: Site frequency spectrum}
        \includegraphics[width=\linewidth, page=1]{/Users/tlatrille/Documents/SelCoeff/experiments/3bins-mC-nodiv/analysis/Bos_taurus.UGBT.MutSel-sfs.normalize.pdf}
    \end{minipage}
    \begin{minipage}{0.32\linewidth}
        \flushleft {\tiny F: $S$ as function of $S_0$ for each class}
        \includegraphics[width=\linewidth, page=1]{/Users/tlatrille/Documents/SelCoeff/experiments/3bins-mC-nodiv/analysis/Bos_taurus.UGBT.MutSel.polyDFE_C.pdf}
    \end{minipage}
    \section{Capra}

    \subsection{Australia}

    \begin{minipage}{0.32\linewidth}
        \flushleft {\tiny A: Expected DFE for mutations}
        \includegraphics[width=\linewidth, page=1]{/Users/tlatrille/Documents/SelCoeff/data_processed/opportunities_bins3Mask0.9/DFE.Capra_hircus.AUCH.MutSel.pdf}
    \end{minipage}
    \begin{minipage}{0.32\linewidth}
        \flushleft {\tiny B: Expected DFE for substitutions}
        \includegraphics[width=\linewidth, page=1]{/Users/tlatrille/Documents/SelCoeff/data_processed/opportunities_bins3Mask0.9/DFE.Capra_hircus.AUCH.MutSel.Flow.pdf}
    \end{minipage}
    \begin{minipage}{0.32\linewidth}
        \flushleft {\tiny C: Observed DFE for substitutions}
        \includegraphics[width=\linewidth, page=1]{/Users/tlatrille/Documents/SelCoeff/supp/hist/subs.Capra_hircus.AUCH.bins3Mask0.9.pdf}
    \end{minipage}
    \\
    \begin{minipage}{0.32\linewidth}
        \flushleft {\tiny D: Observed DFE for polymorphisms}
        \includegraphics[width=\linewidth, page=1]{/Users/tlatrille/Documents/SelCoeff/supp/hist/poly.Capra_hircus.AUCH.bins3Mask0.9.pdf}
    \end{minipage}
    \begin{minipage}{0.32\linewidth}
        \flushleft {\tiny E: Site frequency spectrum}
        \includegraphics[width=\linewidth, page=1]{/Users/tlatrille/Documents/SelCoeff/experiments/3bins-mC-nodiv/analysis/Capra_hircus.AUCH.MutSel-sfs.normalize.pdf}
    \end{minipage}
    \begin{minipage}{0.32\linewidth}
        \flushleft {\tiny F: $S$ as function of $S_0$ for each class}
        \includegraphics[width=\linewidth, page=1]{/Users/tlatrille/Documents/SelCoeff/experiments/3bins-mC-nodiv/analysis/Capra_hircus.AUCH.MutSel.polyDFE_C.pdf}
    \end{minipage}
    \subsection{France}

    \begin{minipage}{0.32\linewidth}
        \flushleft {\tiny A: Expected DFE for mutations}
        \includegraphics[width=\linewidth, page=1]{/Users/tlatrille/Documents/SelCoeff/data_processed/opportunities_bins3Mask0.9/DFE.Capra_hircus.FRCH.MutSel.pdf}
    \end{minipage}
    \begin{minipage}{0.32\linewidth}
        \flushleft {\tiny B: Expected DFE for substitutions}
        \includegraphics[width=\linewidth, page=1]{/Users/tlatrille/Documents/SelCoeff/data_processed/opportunities_bins3Mask0.9/DFE.Capra_hircus.FRCH.MutSel.Flow.pdf}
    \end{minipage}
    \begin{minipage}{0.32\linewidth}
        \flushleft {\tiny C: Observed DFE for substitutions}
        \includegraphics[width=\linewidth, page=1]{/Users/tlatrille/Documents/SelCoeff/supp/hist/subs.Capra_hircus.FRCH.bins3Mask0.9.pdf}
    \end{minipage}
    \\
    \begin{minipage}{0.32\linewidth}
        \flushleft {\tiny D: Observed DFE for polymorphisms}
        \includegraphics[width=\linewidth, page=1]{/Users/tlatrille/Documents/SelCoeff/supp/hist/poly.Capra_hircus.FRCH.bins3Mask0.9.pdf}
    \end{minipage}
    \begin{minipage}{0.32\linewidth}
        \flushleft {\tiny E: Site frequency spectrum}
        \includegraphics[width=\linewidth, page=1]{/Users/tlatrille/Documents/SelCoeff/experiments/3bins-mC-nodiv/analysis/Capra_hircus.FRCH.MutSel-sfs.normalize.pdf}
    \end{minipage}
    \begin{minipage}{0.32\linewidth}
        \flushleft {\tiny F: $S$ as function of $S_0$ for each class}
        \includegraphics[width=\linewidth, page=1]{/Users/tlatrille/Documents/SelCoeff/experiments/3bins-mC-nodiv/analysis/Capra_hircus.FRCH.MutSel.polyDFE_C.pdf}
    \end{minipage}
    \subsection{Iran (C. aegagrus)}

    \begin{minipage}{0.32\linewidth}
        \flushleft {\tiny A: Expected DFE for mutations}
        \includegraphics[width=\linewidth, page=1]{/Users/tlatrille/Documents/SelCoeff/data_processed/opportunities_bins3Mask0.9/DFE.Capra_hircus.IRCA.MutSel.pdf}
    \end{minipage}
    \begin{minipage}{0.32\linewidth}
        \flushleft {\tiny B: Expected DFE for substitutions}
        \includegraphics[width=\linewidth, page=1]{/Users/tlatrille/Documents/SelCoeff/data_processed/opportunities_bins3Mask0.9/DFE.Capra_hircus.IRCA.MutSel.Flow.pdf}
    \end{minipage}
    \begin{minipage}{0.32\linewidth}
        \flushleft {\tiny C: Observed DFE for substitutions}
        \includegraphics[width=\linewidth, page=1]{/Users/tlatrille/Documents/SelCoeff/supp/hist/subs.Capra_hircus.IRCA.bins3Mask0.9.pdf}
    \end{minipage}
    \\
    \begin{minipage}{0.32\linewidth}
        \flushleft {\tiny D: Observed DFE for polymorphisms}
        \includegraphics[width=\linewidth, page=1]{/Users/tlatrille/Documents/SelCoeff/supp/hist/poly.Capra_hircus.IRCA.bins3Mask0.9.pdf}
    \end{minipage}
    \begin{minipage}{0.32\linewidth}
        \flushleft {\tiny E: Site frequency spectrum}
        \includegraphics[width=\linewidth, page=1]{/Users/tlatrille/Documents/SelCoeff/experiments/3bins-mC-nodiv/analysis/Capra_hircus.IRCA.MutSel-sfs.normalize.pdf}
    \end{minipage}
    \begin{minipage}{0.32\linewidth}
        \flushleft {\tiny F: $S$ as function of $S_0$ for each class}
        \includegraphics[width=\linewidth, page=1]{/Users/tlatrille/Documents/SelCoeff/experiments/3bins-mC-nodiv/analysis/Capra_hircus.IRCA.MutSel.polyDFE_C.pdf}
    \end{minipage}
    \subsection{Iran}

    \begin{minipage}{0.32\linewidth}
        \flushleft {\tiny A: Expected DFE for mutations}
        \includegraphics[width=\linewidth, page=1]{/Users/tlatrille/Documents/SelCoeff/data_processed/opportunities_bins3Mask0.9/DFE.Capra_hircus.IRCH.MutSel.pdf}
    \end{minipage}
    \begin{minipage}{0.32\linewidth}
        \flushleft {\tiny B: Expected DFE for substitutions}
        \includegraphics[width=\linewidth, page=1]{/Users/tlatrille/Documents/SelCoeff/data_processed/opportunities_bins3Mask0.9/DFE.Capra_hircus.IRCH.MutSel.Flow.pdf}
    \end{minipage}
    \begin{minipage}{0.32\linewidth}
        \flushleft {\tiny C: Observed DFE for substitutions}
        \includegraphics[width=\linewidth, page=1]{/Users/tlatrille/Documents/SelCoeff/supp/hist/subs.Capra_hircus.IRCH.bins3Mask0.9.pdf}
    \end{minipage}
    \\
    \begin{minipage}{0.32\linewidth}
        \flushleft {\tiny D: Observed DFE for polymorphisms}
        \includegraphics[width=\linewidth, page=1]{/Users/tlatrille/Documents/SelCoeff/supp/hist/poly.Capra_hircus.IRCH.bins3Mask0.9.pdf}
    \end{minipage}
    \begin{minipage}{0.32\linewidth}
        \flushleft {\tiny E: Site frequency spectrum}
        \includegraphics[width=\linewidth, page=1]{/Users/tlatrille/Documents/SelCoeff/experiments/3bins-mC-nodiv/analysis/Capra_hircus.IRCH.MutSel-sfs.normalize.pdf}
    \end{minipage}
    \begin{minipage}{0.32\linewidth}
        \flushleft {\tiny F: $S$ as function of $S_0$ for each class}
        \includegraphics[width=\linewidth, page=1]{/Users/tlatrille/Documents/SelCoeff/experiments/3bins-mC-nodiv/analysis/Capra_hircus.IRCH.MutSel.polyDFE_C.pdf}
    \end{minipage}
    \subsection{Italy}

    \begin{minipage}{0.32\linewidth}
        \flushleft {\tiny A: Expected DFE for mutations}
        \includegraphics[width=\linewidth, page=1]{/Users/tlatrille/Documents/SelCoeff/data_processed/opportunities_bins3Mask0.9/DFE.Capra_hircus.ITCH.MutSel.pdf}
    \end{minipage}
    \begin{minipage}{0.32\linewidth}
        \flushleft {\tiny B: Expected DFE for substitutions}
        \includegraphics[width=\linewidth, page=1]{/Users/tlatrille/Documents/SelCoeff/data_processed/opportunities_bins3Mask0.9/DFE.Capra_hircus.ITCH.MutSel.Flow.pdf}
    \end{minipage}
    \begin{minipage}{0.32\linewidth}
        \flushleft {\tiny C: Observed DFE for substitutions}
        \includegraphics[width=\linewidth, page=1]{/Users/tlatrille/Documents/SelCoeff/supp/hist/subs.Capra_hircus.ITCH.bins3Mask0.9.pdf}
    \end{minipage}
    \\
    \begin{minipage}{0.32\linewidth}
        \flushleft {\tiny D: Observed DFE for polymorphisms}
        \includegraphics[width=\linewidth, page=1]{/Users/tlatrille/Documents/SelCoeff/supp/hist/poly.Capra_hircus.ITCH.bins3Mask0.9.pdf}
    \end{minipage}
    \begin{minipage}{0.32\linewidth}
        \flushleft {\tiny E: Site frequency spectrum}
        \includegraphics[width=\linewidth, page=1]{/Users/tlatrille/Documents/SelCoeff/experiments/3bins-mC-nodiv/analysis/Capra_hircus.ITCH.MutSel-sfs.normalize.pdf}
    \end{minipage}
    \begin{minipage}{0.32\linewidth}
        \flushleft {\tiny F: $S$ as function of $S_0$ for each class}
        \includegraphics[width=\linewidth, page=1]{/Users/tlatrille/Documents/SelCoeff/experiments/3bins-mC-nodiv/analysis/Capra_hircus.ITCH.MutSel.polyDFE_C.pdf}
    \end{minipage}
    \subsection{Morocco}

    \begin{minipage}{0.32\linewidth}
        \flushleft {\tiny A: Expected DFE for mutations}
        \includegraphics[width=\linewidth, page=1]{/Users/tlatrille/Documents/SelCoeff/data_processed/opportunities_bins3Mask0.9/DFE.Capra_hircus.MOCH.MutSel.pdf}
    \end{minipage}
    \begin{minipage}{0.32\linewidth}
        \flushleft {\tiny B: Expected DFE for substitutions}
        \includegraphics[width=\linewidth, page=1]{/Users/tlatrille/Documents/SelCoeff/data_processed/opportunities_bins3Mask0.9/DFE.Capra_hircus.MOCH.MutSel.Flow.pdf}
    \end{minipage}
    \begin{minipage}{0.32\linewidth}
        \flushleft {\tiny C: Observed DFE for substitutions}
        \includegraphics[width=\linewidth, page=1]{/Users/tlatrille/Documents/SelCoeff/supp/hist/subs.Capra_hircus.MOCH.bins3Mask0.9.pdf}
    \end{minipage}
    \\
    \begin{minipage}{0.32\linewidth}
        \flushleft {\tiny D: Observed DFE for polymorphisms}
        \includegraphics[width=\linewidth, page=1]{/Users/tlatrille/Documents/SelCoeff/supp/hist/poly.Capra_hircus.MOCH.bins3Mask0.9.pdf}
    \end{minipage}
    \begin{minipage}{0.32\linewidth}
        \flushleft {\tiny E: Site frequency spectrum}
        \includegraphics[width=\linewidth, page=1]{/Users/tlatrille/Documents/SelCoeff/experiments/3bins-mC-nodiv/analysis/Capra_hircus.MOCH.MutSel-sfs.normalize.pdf}
    \end{minipage}
    \begin{minipage}{0.32\linewidth}
        \flushleft {\tiny F: $S$ as function of $S_0$ for each class}
        \includegraphics[width=\linewidth, page=1]{/Users/tlatrille/Documents/SelCoeff/experiments/3bins-mC-nodiv/analysis/Capra_hircus.MOCH.MutSel.polyDFE_C.pdf}
    \end{minipage}
    \section{Chlorocebus sabaeus}

    \subsection{Barbados}

    \begin{minipage}{0.32\linewidth}
        \flushleft {\tiny A: Expected DFE for mutations}
        \includegraphics[width=\linewidth, page=1]{/Users/tlatrille/Documents/SelCoeff/data_processed/opportunities_bins3Mask0.9/DFE.Chlorocebus_sabaeus.Barbados.MutSel.pdf}
    \end{minipage}
    \begin{minipage}{0.32\linewidth}
        \flushleft {\tiny B: Expected DFE for substitutions}
        \includegraphics[width=\linewidth, page=1]{/Users/tlatrille/Documents/SelCoeff/data_processed/opportunities_bins3Mask0.9/DFE.Chlorocebus_sabaeus.Barbados.MutSel.Flow.pdf}
    \end{minipage}
    \begin{minipage}{0.32\linewidth}
        \flushleft {\tiny C: Observed DFE for substitutions}
        \includegraphics[width=\linewidth, page=1]{/Users/tlatrille/Documents/SelCoeff/supp/hist/subs.Chlorocebus_sabaeus.Barbados.bins3Mask0.9.pdf}
    \end{minipage}
    \\
    \begin{minipage}{0.32\linewidth}
        \flushleft {\tiny D: Observed DFE for polymorphisms}
        \includegraphics[width=\linewidth, page=1]{/Users/tlatrille/Documents/SelCoeff/supp/hist/poly.Chlorocebus_sabaeus.Barbados.bins3Mask0.9.pdf}
    \end{minipage}
    \begin{minipage}{0.32\linewidth}
        \flushleft {\tiny E: Site frequency spectrum}
        \includegraphics[width=\linewidth, page=1]{/Users/tlatrille/Documents/SelCoeff/experiments/3bins-mC-nodiv/analysis/Chlorocebus_sabaeus.Barbados.MutSel-sfs.normalize.pdf}
    \end{minipage}
    \begin{minipage}{0.32\linewidth}
        \flushleft {\tiny F: $S$ as function of $S_0$ for each class}
        \includegraphics[width=\linewidth, page=1]{/Users/tlatrille/Documents/SelCoeff/experiments/3bins-mC-nodiv/analysis/Chlorocebus_sabaeus.Barbados.MutSel.polyDFE_C.pdf}
    \end{minipage}
    \subsection{Central Afr. Rep.}

    \begin{minipage}{0.32\linewidth}
        \flushleft {\tiny A: Expected DFE for mutations}
        \includegraphics[width=\linewidth, page=1]{/Users/tlatrille/Documents/SelCoeff/data_processed/opportunities_bins3Mask0.9/DFE.Chlorocebus_sabaeus.Central_African_Republic.MutSel.pdf}
    \end{minipage}
    \begin{minipage}{0.32\linewidth}
        \flushleft {\tiny B: Expected DFE for substitutions}
        \includegraphics[width=\linewidth, page=1]{/Users/tlatrille/Documents/SelCoeff/data_processed/opportunities_bins3Mask0.9/DFE.Chlorocebus_sabaeus.Central_African_Republic.MutSel.Flow.pdf}
    \end{minipage}
    \begin{minipage}{0.32\linewidth}
        \flushleft {\tiny C: Observed DFE for substitutions}
        \includegraphics[width=\linewidth, page=1]{/Users/tlatrille/Documents/SelCoeff/supp/hist/subs.Chlorocebus_sabaeus.Central_African_Republic.bins3Mask0.9.pdf}
    \end{minipage}
    \\
    \begin{minipage}{0.32\linewidth}
        \flushleft {\tiny D: Observed DFE for polymorphisms}
        \includegraphics[width=\linewidth, page=1]{/Users/tlatrille/Documents/SelCoeff/supp/hist/poly.Chlorocebus_sabaeus.Central_African_Republic.bins3Mask0.9.pdf}
    \end{minipage}
    \begin{minipage}{0.32\linewidth}
        \flushleft {\tiny E: Site frequency spectrum}
        \includegraphics[width=\linewidth, page=1]{/Users/tlatrille/Documents/SelCoeff/experiments/3bins-mC-nodiv/analysis/Chlorocebus_sabaeus.Central_African_Republic.MutSel-sfs.normalize.pdf}
    \end{minipage}
    \begin{minipage}{0.32\linewidth}
        \flushleft {\tiny F: $S$ as function of $S_0$ for each class}
        \includegraphics[width=\linewidth, page=1]{/Users/tlatrille/Documents/SelCoeff/experiments/3bins-mC-nodiv/analysis/Chlorocebus_sabaeus.Central_African_Republic.MutSel.polyDFE_C.pdf}
    \end{minipage}
    \subsection{Ethiopia}

    \begin{minipage}{0.32\linewidth}
        \flushleft {\tiny A: Expected DFE for mutations}
        \includegraphics[width=\linewidth, page=1]{/Users/tlatrille/Documents/SelCoeff/data_processed/opportunities_bins3Mask0.9/DFE.Chlorocebus_sabaeus.Ethiopia.MutSel.pdf}
    \end{minipage}
    \begin{minipage}{0.32\linewidth}
        \flushleft {\tiny B: Expected DFE for substitutions}
        \includegraphics[width=\linewidth, page=1]{/Users/tlatrille/Documents/SelCoeff/data_processed/opportunities_bins3Mask0.9/DFE.Chlorocebus_sabaeus.Ethiopia.MutSel.Flow.pdf}
    \end{minipage}
    \begin{minipage}{0.32\linewidth}
        \flushleft {\tiny C: Observed DFE for substitutions}
        \includegraphics[width=\linewidth, page=1]{/Users/tlatrille/Documents/SelCoeff/supp/hist/subs.Chlorocebus_sabaeus.Ethiopia.bins3Mask0.9.pdf}
    \end{minipage}
    \\
    \begin{minipage}{0.32\linewidth}
        \flushleft {\tiny D: Observed DFE for polymorphisms}
        \includegraphics[width=\linewidth, page=1]{/Users/tlatrille/Documents/SelCoeff/supp/hist/poly.Chlorocebus_sabaeus.Ethiopia.bins3Mask0.9.pdf}
    \end{minipage}
    \begin{minipage}{0.32\linewidth}
        \flushleft {\tiny E: Site frequency spectrum}
        \includegraphics[width=\linewidth, page=1]{/Users/tlatrille/Documents/SelCoeff/experiments/3bins-mC-nodiv/analysis/Chlorocebus_sabaeus.Ethiopia.MutSel-sfs.normalize.pdf}
    \end{minipage}
    \begin{minipage}{0.32\linewidth}
        \flushleft {\tiny F: $S$ as function of $S_0$ for each class}
        \includegraphics[width=\linewidth, page=1]{/Users/tlatrille/Documents/SelCoeff/experiments/3bins-mC-nodiv/analysis/Chlorocebus_sabaeus.Ethiopia.MutSel.polyDFE_C.pdf}
    \end{minipage}
    \subsection{Gambia}

    \begin{minipage}{0.32\linewidth}
        \flushleft {\tiny A: Expected DFE for mutations}
        \includegraphics[width=\linewidth, page=1]{/Users/tlatrille/Documents/SelCoeff/data_processed/opportunities_bins3Mask0.9/DFE.Chlorocebus_sabaeus.Gambia.MutSel.pdf}
    \end{minipage}
    \begin{minipage}{0.32\linewidth}
        \flushleft {\tiny B: Expected DFE for substitutions}
        \includegraphics[width=\linewidth, page=1]{/Users/tlatrille/Documents/SelCoeff/data_processed/opportunities_bins3Mask0.9/DFE.Chlorocebus_sabaeus.Gambia.MutSel.Flow.pdf}
    \end{minipage}
    \begin{minipage}{0.32\linewidth}
        \flushleft {\tiny C: Observed DFE for substitutions}
        \includegraphics[width=\linewidth, page=1]{/Users/tlatrille/Documents/SelCoeff/supp/hist/subs.Chlorocebus_sabaeus.Gambia.bins3Mask0.9.pdf}
    \end{minipage}
    \\
    \begin{minipage}{0.32\linewidth}
        \flushleft {\tiny D: Observed DFE for polymorphisms}
        \includegraphics[width=\linewidth, page=1]{/Users/tlatrille/Documents/SelCoeff/supp/hist/poly.Chlorocebus_sabaeus.Gambia.bins3Mask0.9.pdf}
    \end{minipage}
    \begin{minipage}{0.32\linewidth}
        \flushleft {\tiny E: Site frequency spectrum}
        \includegraphics[width=\linewidth, page=1]{/Users/tlatrille/Documents/SelCoeff/experiments/3bins-mC-nodiv/analysis/Chlorocebus_sabaeus.Gambia.MutSel-sfs.normalize.pdf}
    \end{minipage}
    \begin{minipage}{0.32\linewidth}
        \flushleft {\tiny F: $S$ as function of $S_0$ for each class}
        \includegraphics[width=\linewidth, page=1]{/Users/tlatrille/Documents/SelCoeff/experiments/3bins-mC-nodiv/analysis/Chlorocebus_sabaeus.Gambia.MutSel.polyDFE_C.pdf}
    \end{minipage}
    \subsection{Kenya}

    \begin{minipage}{0.32\linewidth}
        \flushleft {\tiny A: Expected DFE for mutations}
        \includegraphics[width=\linewidth, page=1]{/Users/tlatrille/Documents/SelCoeff/data_processed/opportunities_bins3Mask0.9/DFE.Chlorocebus_sabaeus.Kenya.MutSel.pdf}
    \end{minipage}
    \begin{minipage}{0.32\linewidth}
        \flushleft {\tiny B: Expected DFE for substitutions}
        \includegraphics[width=\linewidth, page=1]{/Users/tlatrille/Documents/SelCoeff/data_processed/opportunities_bins3Mask0.9/DFE.Chlorocebus_sabaeus.Kenya.MutSel.Flow.pdf}
    \end{minipage}
    \begin{minipage}{0.32\linewidth}
        \flushleft {\tiny C: Observed DFE for substitutions}
        \includegraphics[width=\linewidth, page=1]{/Users/tlatrille/Documents/SelCoeff/supp/hist/subs.Chlorocebus_sabaeus.Kenya.bins3Mask0.9.pdf}
    \end{minipage}
    \\
    \begin{minipage}{0.32\linewidth}
        \flushleft {\tiny D: Observed DFE for polymorphisms}
        \includegraphics[width=\linewidth, page=1]{/Users/tlatrille/Documents/SelCoeff/supp/hist/poly.Chlorocebus_sabaeus.Kenya.bins3Mask0.9.pdf}
    \end{minipage}
    \begin{minipage}{0.32\linewidth}
        \flushleft {\tiny E: Site frequency spectrum}
        \includegraphics[width=\linewidth, page=1]{/Users/tlatrille/Documents/SelCoeff/experiments/3bins-mC-nodiv/analysis/Chlorocebus_sabaeus.Kenya.MutSel-sfs.normalize.pdf}
    \end{minipage}
    \begin{minipage}{0.32\linewidth}
        \flushleft {\tiny F: $S$ as function of $S_0$ for each class}
        \includegraphics[width=\linewidth, page=1]{/Users/tlatrille/Documents/SelCoeff/experiments/3bins-mC-nodiv/analysis/Chlorocebus_sabaeus.Kenya.MutSel.polyDFE_C.pdf}
    \end{minipage}
    \subsection{Nevis}

    \begin{minipage}{0.32\linewidth}
        \flushleft {\tiny A: Expected DFE for mutations}
        \includegraphics[width=\linewidth, page=1]{/Users/tlatrille/Documents/SelCoeff/data_processed/opportunities_bins3Mask0.9/DFE.Chlorocebus_sabaeus.Nevis.MutSel.pdf}
    \end{minipage}
    \begin{minipage}{0.32\linewidth}
        \flushleft {\tiny B: Expected DFE for substitutions}
        \includegraphics[width=\linewidth, page=1]{/Users/tlatrille/Documents/SelCoeff/data_processed/opportunities_bins3Mask0.9/DFE.Chlorocebus_sabaeus.Nevis.MutSel.Flow.pdf}
    \end{minipage}
    \begin{minipage}{0.32\linewidth}
        \flushleft {\tiny C: Observed DFE for substitutions}
        \includegraphics[width=\linewidth, page=1]{/Users/tlatrille/Documents/SelCoeff/supp/hist/subs.Chlorocebus_sabaeus.Nevis.bins3Mask0.9.pdf}
    \end{minipage}
    \\
    \begin{minipage}{0.32\linewidth}
        \flushleft {\tiny D: Observed DFE for polymorphisms}
        \includegraphics[width=\linewidth, page=1]{/Users/tlatrille/Documents/SelCoeff/supp/hist/poly.Chlorocebus_sabaeus.Nevis.bins3Mask0.9.pdf}
    \end{minipage}
    \begin{minipage}{0.32\linewidth}
        \flushleft {\tiny E: Site frequency spectrum}
        \includegraphics[width=\linewidth, page=1]{/Users/tlatrille/Documents/SelCoeff/experiments/3bins-mC-nodiv/analysis/Chlorocebus_sabaeus.Nevis.MutSel-sfs.normalize.pdf}
    \end{minipage}
    \begin{minipage}{0.32\linewidth}
        \flushleft {\tiny F: $S$ as function of $S_0$ for each class}
        \includegraphics[width=\linewidth, page=1]{/Users/tlatrille/Documents/SelCoeff/experiments/3bins-mC-nodiv/analysis/Chlorocebus_sabaeus.Nevis.MutSel.polyDFE_C.pdf}
    \end{minipage}
    \subsection{Saint Kitts}

    \begin{minipage}{0.32\linewidth}
        \flushleft {\tiny A: Expected DFE for mutations}
        \includegraphics[width=\linewidth, page=1]{/Users/tlatrille/Documents/SelCoeff/data_processed/opportunities_bins3Mask0.9/DFE.Chlorocebus_sabaeus.Saint_Kitts.MutSel.pdf}
    \end{minipage}
    \begin{minipage}{0.32\linewidth}
        \flushleft {\tiny B: Expected DFE for substitutions}
        \includegraphics[width=\linewidth, page=1]{/Users/tlatrille/Documents/SelCoeff/data_processed/opportunities_bins3Mask0.9/DFE.Chlorocebus_sabaeus.Saint_Kitts.MutSel.Flow.pdf}
    \end{minipage}
    \begin{minipage}{0.32\linewidth}
        \flushleft {\tiny C: Observed DFE for substitutions}
        \includegraphics[width=\linewidth, page=1]{/Users/tlatrille/Documents/SelCoeff/supp/hist/subs.Chlorocebus_sabaeus.Saint_Kitts.bins3Mask0.9.pdf}
    \end{minipage}
    \\
    \begin{minipage}{0.32\linewidth}
        \flushleft {\tiny D: Observed DFE for polymorphisms}
        \includegraphics[width=\linewidth, page=1]{/Users/tlatrille/Documents/SelCoeff/supp/hist/poly.Chlorocebus_sabaeus.Saint_Kitts.bins3Mask0.9.pdf}
    \end{minipage}
    \begin{minipage}{0.32\linewidth}
        \flushleft {\tiny E: Site frequency spectrum}
        \includegraphics[width=\linewidth, page=1]{/Users/tlatrille/Documents/SelCoeff/experiments/3bins-mC-nodiv/analysis/Chlorocebus_sabaeus.Saint_Kitts.MutSel-sfs.normalize.pdf}
    \end{minipage}
    \begin{minipage}{0.32\linewidth}
        \flushleft {\tiny F: $S$ as function of $S_0$ for each class}
        \includegraphics[width=\linewidth, page=1]{/Users/tlatrille/Documents/SelCoeff/experiments/3bins-mC-nodiv/analysis/Chlorocebus_sabaeus.Saint_Kitts.MutSel.polyDFE_C.pdf}
    \end{minipage}
    \subsection{South Africa}

    \begin{minipage}{0.32\linewidth}
        \flushleft {\tiny A: Expected DFE for mutations}
        \includegraphics[width=\linewidth, page=1]{/Users/tlatrille/Documents/SelCoeff/data_processed/opportunities_bins3Mask0.9/DFE.Chlorocebus_sabaeus.South_Africa.MutSel.pdf}
    \end{minipage}
    \begin{minipage}{0.32\linewidth}
        \flushleft {\tiny B: Expected DFE for substitutions}
        \includegraphics[width=\linewidth, page=1]{/Users/tlatrille/Documents/SelCoeff/data_processed/opportunities_bins3Mask0.9/DFE.Chlorocebus_sabaeus.South_Africa.MutSel.Flow.pdf}
    \end{minipage}
    \begin{minipage}{0.32\linewidth}
        \flushleft {\tiny C: Observed DFE for substitutions}
        \includegraphics[width=\linewidth, page=1]{/Users/tlatrille/Documents/SelCoeff/supp/hist/subs.Chlorocebus_sabaeus.South_Africa.bins3Mask0.9.pdf}
    \end{minipage}
    \\
    \begin{minipage}{0.32\linewidth}
        \flushleft {\tiny D: Observed DFE for polymorphisms}
        \includegraphics[width=\linewidth, page=1]{/Users/tlatrille/Documents/SelCoeff/supp/hist/poly.Chlorocebus_sabaeus.South_Africa.bins3Mask0.9.pdf}
    \end{minipage}
    \begin{minipage}{0.32\linewidth}
        \flushleft {\tiny E: Site frequency spectrum}
        \includegraphics[width=\linewidth, page=1]{/Users/tlatrille/Documents/SelCoeff/experiments/3bins-mC-nodiv/analysis/Chlorocebus_sabaeus.South_Africa.MutSel-sfs.normalize.pdf}
    \end{minipage}
    \begin{minipage}{0.32\linewidth}
        \flushleft {\tiny F: $S$ as function of $S_0$ for each class}
        \includegraphics[width=\linewidth, page=1]{/Users/tlatrille/Documents/SelCoeff/experiments/3bins-mC-nodiv/analysis/Chlorocebus_sabaeus.South_Africa.MutSel.polyDFE_C.pdf}
    \end{minipage}
    \subsection{Zambia}

    \begin{minipage}{0.32\linewidth}
        \flushleft {\tiny A: Expected DFE for mutations}
        \includegraphics[width=\linewidth, page=1]{/Users/tlatrille/Documents/SelCoeff/data_processed/opportunities_bins3Mask0.9/DFE.Chlorocebus_sabaeus.Zambia.MutSel.pdf}
    \end{minipage}
    \begin{minipage}{0.32\linewidth}
        \flushleft {\tiny B: Expected DFE for substitutions}
        \includegraphics[width=\linewidth, page=1]{/Users/tlatrille/Documents/SelCoeff/data_processed/opportunities_bins3Mask0.9/DFE.Chlorocebus_sabaeus.Zambia.MutSel.Flow.pdf}
    \end{minipage}
    \begin{minipage}{0.32\linewidth}
        \flushleft {\tiny C: Observed DFE for substitutions}
        \includegraphics[width=\linewidth, page=1]{/Users/tlatrille/Documents/SelCoeff/supp/hist/subs.Chlorocebus_sabaeus.Zambia.bins3Mask0.9.pdf}
    \end{minipage}
    \\
    \begin{minipage}{0.32\linewidth}
        \flushleft {\tiny D: Observed DFE for polymorphisms}
        \includegraphics[width=\linewidth, page=1]{/Users/tlatrille/Documents/SelCoeff/supp/hist/poly.Chlorocebus_sabaeus.Zambia.bins3Mask0.9.pdf}
    \end{minipage}
    \begin{minipage}{0.32\linewidth}
        \flushleft {\tiny E: Site frequency spectrum}
        \includegraphics[width=\linewidth, page=1]{/Users/tlatrille/Documents/SelCoeff/experiments/3bins-mC-nodiv/analysis/Chlorocebus_sabaeus.Zambia.MutSel-sfs.normalize.pdf}
    \end{minipage}
    \begin{minipage}{0.32\linewidth}
        \flushleft {\tiny F: $S$ as function of $S_0$ for each class}
        \includegraphics[width=\linewidth, page=1]{/Users/tlatrille/Documents/SelCoeff/experiments/3bins-mC-nodiv/analysis/Chlorocebus_sabaeus.Zambia.MutSel.polyDFE_C.pdf}
    \end{minipage}
    \section{Equus caballus}

    \begin{minipage}{0.32\linewidth}
        \flushleft {\tiny A: Expected DFE for mutations}
        \includegraphics[width=\linewidth, page=1]{/Users/tlatrille/Documents/SelCoeff/data_processed/opportunities_bins3Mask0.9/DFE.Equus_caballus.up.MutSel.pdf}
    \end{minipage}
    \begin{minipage}{0.32\linewidth}
        \flushleft {\tiny B: Expected DFE for substitutions}
        \includegraphics[width=\linewidth, page=1]{/Users/tlatrille/Documents/SelCoeff/data_processed/opportunities_bins3Mask0.9/DFE.Equus_caballus.up.MutSel.Flow.pdf}
    \end{minipage}
    \begin{minipage}{0.32\linewidth}
        \flushleft {\tiny C: Observed DFE for substitutions}
        \includegraphics[width=\linewidth, page=1]{/Users/tlatrille/Documents/SelCoeff/supp/hist/subs.Equus_caballus.up.bins3Mask0.9.pdf}
    \end{minipage}
    \\
    \begin{minipage}{0.32\linewidth}
        \flushleft {\tiny D: Observed DFE for polymorphisms}
        \includegraphics[width=\linewidth, page=1]{/Users/tlatrille/Documents/SelCoeff/supp/hist/poly.Equus_caballus.up.bins3Mask0.9.pdf}
    \end{minipage}
    \begin{minipage}{0.32\linewidth}
        \flushleft {\tiny E: Site frequency spectrum}
        \includegraphics[width=\linewidth, page=1]{/Users/tlatrille/Documents/SelCoeff/experiments/3bins-mC-nodiv/analysis/Equus_caballus.up.MutSel-sfs.normalize.pdf}
    \end{minipage}
    \begin{minipage}{0.32\linewidth}
        \flushleft {\tiny F: $S$ as function of $S_0$ for each class}
        \includegraphics[width=\linewidth, page=1]{/Users/tlatrille/Documents/SelCoeff/experiments/3bins-mC-nodiv/analysis/Equus_caballus.up.MutSel.polyDFE_C.pdf}
    \end{minipage}
    \section{Homo sapiens}

    \subsection{African}

    \begin{minipage}{0.32\linewidth}
        \flushleft {\tiny A: Expected DFE for mutations}
        \includegraphics[width=\linewidth, page=1]{/Users/tlatrille/Documents/SelCoeff/data_processed/opportunities_bins3Mask0.9/DFE.Homo_sapiens.AFR.MutSel.pdf}
    \end{minipage}
    \begin{minipage}{0.32\linewidth}
        \flushleft {\tiny B: Expected DFE for substitutions}
        \includegraphics[width=\linewidth, page=1]{/Users/tlatrille/Documents/SelCoeff/data_processed/opportunities_bins3Mask0.9/DFE.Homo_sapiens.AFR.MutSel.Flow.pdf}
    \end{minipage}
    \begin{minipage}{0.32\linewidth}
        \flushleft {\tiny C: Observed DFE for substitutions}
        \includegraphics[width=\linewidth, page=1]{/Users/tlatrille/Documents/SelCoeff/supp/hist/subs.Homo_sapiens.AFR.bins3Mask0.9.pdf}
    \end{minipage}
    \\
    \begin{minipage}{0.32\linewidth}
        \flushleft {\tiny D: Observed DFE for polymorphisms}
        \includegraphics[width=\linewidth, page=1]{/Users/tlatrille/Documents/SelCoeff/supp/hist/poly.Homo_sapiens.AFR.bins3Mask0.9.pdf}
    \end{minipage}
    \begin{minipage}{0.32\linewidth}
        \flushleft {\tiny E: Site frequency spectrum}
        \includegraphics[width=\linewidth, page=1]{/Users/tlatrille/Documents/SelCoeff/experiments/3bins-mC-nodiv/analysis/Homo_sapiens.AFR.MutSel-sfs.normalize.pdf}
    \end{minipage}
    \begin{minipage}{0.32\linewidth}
        \flushleft {\tiny F: $S$ as function of $S_0$ for each class}
        \includegraphics[width=\linewidth, page=1]{/Users/tlatrille/Documents/SelCoeff/experiments/3bins-mC-nodiv/analysis/Homo_sapiens.AFR.MutSel.polyDFE_C.pdf}
    \end{minipage}
    \subsection{Admixed American}

    \begin{minipage}{0.32\linewidth}
        \flushleft {\tiny A: Expected DFE for mutations}
        \includegraphics[width=\linewidth, page=1]{/Users/tlatrille/Documents/SelCoeff/data_processed/opportunities_bins3Mask0.9/DFE.Homo_sapiens.AMR.MutSel.pdf}
    \end{minipage}
    \begin{minipage}{0.32\linewidth}
        \flushleft {\tiny B: Expected DFE for substitutions}
        \includegraphics[width=\linewidth, page=1]{/Users/tlatrille/Documents/SelCoeff/data_processed/opportunities_bins3Mask0.9/DFE.Homo_sapiens.AMR.MutSel.Flow.pdf}
    \end{minipage}
    \begin{minipage}{0.32\linewidth}
        \flushleft {\tiny C: Observed DFE for substitutions}
        \includegraphics[width=\linewidth, page=1]{/Users/tlatrille/Documents/SelCoeff/supp/hist/subs.Homo_sapiens.AMR.bins3Mask0.9.pdf}
    \end{minipage}
    \\
    \begin{minipage}{0.32\linewidth}
        \flushleft {\tiny D: Observed DFE for polymorphisms}
        \includegraphics[width=\linewidth, page=1]{/Users/tlatrille/Documents/SelCoeff/supp/hist/poly.Homo_sapiens.AMR.bins3Mask0.9.pdf}
    \end{minipage}
    \begin{minipage}{0.32\linewidth}
        \flushleft {\tiny E: Site frequency spectrum}
        \includegraphics[width=\linewidth, page=1]{/Users/tlatrille/Documents/SelCoeff/experiments/3bins-mC-nodiv/analysis/Homo_sapiens.AMR.MutSel-sfs.normalize.pdf}
    \end{minipage}
    \begin{minipage}{0.32\linewidth}
        \flushleft {\tiny F: $S$ as function of $S_0$ for each class}
        \includegraphics[width=\linewidth, page=1]{/Users/tlatrille/Documents/SelCoeff/experiments/3bins-mC-nodiv/analysis/Homo_sapiens.AMR.MutSel.polyDFE_C.pdf}
    \end{minipage}
    \subsection{East Asian}

    \begin{minipage}{0.32\linewidth}
        \flushleft {\tiny A: Expected DFE for mutations}
        \includegraphics[width=\linewidth, page=1]{/Users/tlatrille/Documents/SelCoeff/data_processed/opportunities_bins3Mask0.9/DFE.Homo_sapiens.EAS.MutSel.pdf}
    \end{minipage}
    \begin{minipage}{0.32\linewidth}
        \flushleft {\tiny B: Expected DFE for substitutions}
        \includegraphics[width=\linewidth, page=1]{/Users/tlatrille/Documents/SelCoeff/data_processed/opportunities_bins3Mask0.9/DFE.Homo_sapiens.EAS.MutSel.Flow.pdf}
    \end{minipage}
    \begin{minipage}{0.32\linewidth}
        \flushleft {\tiny C: Observed DFE for substitutions}
        \includegraphics[width=\linewidth, page=1]{/Users/tlatrille/Documents/SelCoeff/supp/hist/subs.Homo_sapiens.EAS.bins3Mask0.9.pdf}
    \end{minipage}
    \\
    \begin{minipage}{0.32\linewidth}
        \flushleft {\tiny D: Observed DFE for polymorphisms}
        \includegraphics[width=\linewidth, page=1]{/Users/tlatrille/Documents/SelCoeff/supp/hist/poly.Homo_sapiens.EAS.bins3Mask0.9.pdf}
    \end{minipage}
    \begin{minipage}{0.32\linewidth}
        \flushleft {\tiny E: Site frequency spectrum}
        \includegraphics[width=\linewidth, page=1]{/Users/tlatrille/Documents/SelCoeff/experiments/3bins-mC-nodiv/analysis/Homo_sapiens.EAS.MutSel-sfs.normalize.pdf}
    \end{minipage}
    \begin{minipage}{0.32\linewidth}
        \flushleft {\tiny F: $S$ as function of $S_0$ for each class}
        \includegraphics[width=\linewidth, page=1]{/Users/tlatrille/Documents/SelCoeff/experiments/3bins-mC-nodiv/analysis/Homo_sapiens.EAS.MutSel.polyDFE_C.pdf}
    \end{minipage}
    \subsection{European}

    \begin{minipage}{0.32\linewidth}
        \flushleft {\tiny A: Expected DFE for mutations}
        \includegraphics[width=\linewidth, page=1]{/Users/tlatrille/Documents/SelCoeff/data_processed/opportunities_bins3Mask0.9/DFE.Homo_sapiens.EUR.MutSel.pdf}
    \end{minipage}
    \begin{minipage}{0.32\linewidth}
        \flushleft {\tiny B: Expected DFE for substitutions}
        \includegraphics[width=\linewidth, page=1]{/Users/tlatrille/Documents/SelCoeff/data_processed/opportunities_bins3Mask0.9/DFE.Homo_sapiens.EUR.MutSel.Flow.pdf}
    \end{minipage}
    \begin{minipage}{0.32\linewidth}
        \flushleft {\tiny C: Observed DFE for substitutions}
        \includegraphics[width=\linewidth, page=1]{/Users/tlatrille/Documents/SelCoeff/supp/hist/subs.Homo_sapiens.EUR.bins3Mask0.9.pdf}
    \end{minipage}
    \\
    \begin{minipage}{0.32\linewidth}
        \flushleft {\tiny D: Observed DFE for polymorphisms}
        \includegraphics[width=\linewidth, page=1]{/Users/tlatrille/Documents/SelCoeff/supp/hist/poly.Homo_sapiens.EUR.bins3Mask0.9.pdf}
    \end{minipage}
    \begin{minipage}{0.32\linewidth}
        \flushleft {\tiny E: Site frequency spectrum}
        \includegraphics[width=\linewidth, page=1]{/Users/tlatrille/Documents/SelCoeff/experiments/3bins-mC-nodiv/analysis/Homo_sapiens.EUR.MutSel-sfs.normalize.pdf}
    \end{minipage}
    \begin{minipage}{0.32\linewidth}
        \flushleft {\tiny F: $S$ as function of $S_0$ for each class}
        \includegraphics[width=\linewidth, page=1]{/Users/tlatrille/Documents/SelCoeff/experiments/3bins-mC-nodiv/analysis/Homo_sapiens.EUR.MutSel.polyDFE_C.pdf}
    \end{minipage}
    \subsection{South Asian}

    \begin{minipage}{0.32\linewidth}
        \flushleft {\tiny A: Expected DFE for mutations}
        \includegraphics[width=\linewidth, page=1]{/Users/tlatrille/Documents/SelCoeff/data_processed/opportunities_bins3Mask0.9/DFE.Homo_sapiens.SAS.MutSel.pdf}
    \end{minipage}
    \begin{minipage}{0.32\linewidth}
        \flushleft {\tiny B: Expected DFE for substitutions}
        \includegraphics[width=\linewidth, page=1]{/Users/tlatrille/Documents/SelCoeff/data_processed/opportunities_bins3Mask0.9/DFE.Homo_sapiens.SAS.MutSel.Flow.pdf}
    \end{minipage}
    \begin{minipage}{0.32\linewidth}
        \flushleft {\tiny C: Observed DFE for substitutions}
        \includegraphics[width=\linewidth, page=1]{/Users/tlatrille/Documents/SelCoeff/supp/hist/subs.Homo_sapiens.SAS.bins3Mask0.9.pdf}
    \end{minipage}
    \\
    \begin{minipage}{0.32\linewidth}
        \flushleft {\tiny D: Observed DFE for polymorphisms}
        \includegraphics[width=\linewidth, page=1]{/Users/tlatrille/Documents/SelCoeff/supp/hist/poly.Homo_sapiens.SAS.bins3Mask0.9.pdf}
    \end{minipage}
    \begin{minipage}{0.32\linewidth}
        \flushleft {\tiny E: Site frequency spectrum}
        \includegraphics[width=\linewidth, page=1]{/Users/tlatrille/Documents/SelCoeff/experiments/3bins-mC-nodiv/analysis/Homo_sapiens.SAS.MutSel-sfs.normalize.pdf}
    \end{minipage}
    \begin{minipage}{0.32\linewidth}
        \flushleft {\tiny F: $S$ as function of $S_0$ for each class}
        \includegraphics[width=\linewidth, page=1]{/Users/tlatrille/Documents/SelCoeff/experiments/3bins-mC-nodiv/analysis/Homo_sapiens.SAS.MutSel.polyDFE_C.pdf}
    \end{minipage}
    \section{Ovis}

    \subsection{Iran}

    \begin{minipage}{0.32\linewidth}
        \flushleft {\tiny A: Expected DFE for mutations}
        \includegraphics[width=\linewidth, page=1]{/Users/tlatrille/Documents/SelCoeff/data_processed/opportunities_bins3Mask0.9/DFE.Ovis_aries.IROA.MutSel.pdf}
    \end{minipage}
    \begin{minipage}{0.32\linewidth}
        \flushleft {\tiny B: Expected DFE for substitutions}
        \includegraphics[width=\linewidth, page=1]{/Users/tlatrille/Documents/SelCoeff/data_processed/opportunities_bins3Mask0.9/DFE.Ovis_aries.IROA.MutSel.Flow.pdf}
    \end{minipage}
    \begin{minipage}{0.32\linewidth}
        \flushleft {\tiny C: Observed DFE for substitutions}
        \includegraphics[width=\linewidth, page=1]{/Users/tlatrille/Documents/SelCoeff/supp/hist/subs.Ovis_aries.IROA.bins3Mask0.9.pdf}
    \end{minipage}
    \\
    \begin{minipage}{0.32\linewidth}
        \flushleft {\tiny D: Observed DFE for polymorphisms}
        \includegraphics[width=\linewidth, page=1]{/Users/tlatrille/Documents/SelCoeff/supp/hist/poly.Ovis_aries.IROA.bins3Mask0.9.pdf}
    \end{minipage}
    \begin{minipage}{0.32\linewidth}
        \flushleft {\tiny E: Site frequency spectrum}
        \includegraphics[width=\linewidth, page=1]{/Users/tlatrille/Documents/SelCoeff/experiments/3bins-mC-nodiv/analysis/Ovis_aries.IROA.MutSel-sfs.normalize.pdf}
    \end{minipage}
    \begin{minipage}{0.32\linewidth}
        \flushleft {\tiny F: $S$ as function of $S_0$ for each class}
        \includegraphics[width=\linewidth, page=1]{/Users/tlatrille/Documents/SelCoeff/experiments/3bins-mC-nodiv/analysis/Ovis_aries.IROA.MutSel.polyDFE_C.pdf}
    \end{minipage}
    \subsection{Iran (O. orientalis)}

    \begin{minipage}{0.32\linewidth}
        \flushleft {\tiny A: Expected DFE for mutations}
        \includegraphics[width=\linewidth, page=1]{/Users/tlatrille/Documents/SelCoeff/data_processed/opportunities_bins3Mask0.9/DFE.Ovis_aries.IROO.MutSel.pdf}
    \end{minipage}
    \begin{minipage}{0.32\linewidth}
        \flushleft {\tiny B: Expected DFE for substitutions}
        \includegraphics[width=\linewidth, page=1]{/Users/tlatrille/Documents/SelCoeff/data_processed/opportunities_bins3Mask0.9/DFE.Ovis_aries.IROO.MutSel.Flow.pdf}
    \end{minipage}
    \begin{minipage}{0.32\linewidth}
        \flushleft {\tiny C: Observed DFE for substitutions}
        \includegraphics[width=\linewidth, page=1]{/Users/tlatrille/Documents/SelCoeff/supp/hist/subs.Ovis_aries.IROO.bins3Mask0.9.pdf}
    \end{minipage}
    \\
    \begin{minipage}{0.32\linewidth}
        \flushleft {\tiny D: Observed DFE for polymorphisms}
        \includegraphics[width=\linewidth, page=1]{/Users/tlatrille/Documents/SelCoeff/supp/hist/poly.Ovis_aries.IROO.bins3Mask0.9.pdf}
    \end{minipage}
    \begin{minipage}{0.32\linewidth}
        \flushleft {\tiny E: Site frequency spectrum}
        \includegraphics[width=\linewidth, page=1]{/Users/tlatrille/Documents/SelCoeff/experiments/3bins-mC-nodiv/analysis/Ovis_aries.IROO.MutSel-sfs.normalize.pdf}
    \end{minipage}
    \begin{minipage}{0.32\linewidth}
        \flushleft {\tiny F: $S$ as function of $S_0$ for each class}
        \includegraphics[width=\linewidth, page=1]{/Users/tlatrille/Documents/SelCoeff/experiments/3bins-mC-nodiv/analysis/Ovis_aries.IROO.MutSel.polyDFE_C.pdf}
    \end{minipage}
    \subsection{Iran (O. vignei)}

    \begin{minipage}{0.32\linewidth}
        \flushleft {\tiny A: Expected DFE for mutations}
        \includegraphics[width=\linewidth, page=1]{/Users/tlatrille/Documents/SelCoeff/data_processed/opportunities_bins3Mask0.9/DFE.Ovis_aries.IROV.MutSel.pdf}
    \end{minipage}
    \begin{minipage}{0.32\linewidth}
        \flushleft {\tiny B: Expected DFE for substitutions}
        \includegraphics[width=\linewidth, page=1]{/Users/tlatrille/Documents/SelCoeff/data_processed/opportunities_bins3Mask0.9/DFE.Ovis_aries.IROV.MutSel.Flow.pdf}
    \end{minipage}
    \begin{minipage}{0.32\linewidth}
        \flushleft {\tiny C: Observed DFE for substitutions}
        \includegraphics[width=\linewidth, page=1]{/Users/tlatrille/Documents/SelCoeff/supp/hist/subs.Ovis_aries.IROV.bins3Mask0.9.pdf}
    \end{minipage}
    \\
    \begin{minipage}{0.32\linewidth}
        \flushleft {\tiny D: Observed DFE for polymorphisms}
        \includegraphics[width=\linewidth, page=1]{/Users/tlatrille/Documents/SelCoeff/supp/hist/poly.Ovis_aries.IROV.bins3Mask0.9.pdf}
    \end{minipage}
    \begin{minipage}{0.32\linewidth}
        \flushleft {\tiny E: Site frequency spectrum}
        \includegraphics[width=\linewidth, page=1]{/Users/tlatrille/Documents/SelCoeff/experiments/3bins-mC-nodiv/analysis/Ovis_aries.IROV.MutSel-sfs.normalize.pdf}
    \end{minipage}
    \begin{minipage}{0.32\linewidth}
        \flushleft {\tiny F: $S$ as function of $S_0$ for each class}
        \includegraphics[width=\linewidth, page=1]{/Users/tlatrille/Documents/SelCoeff/experiments/3bins-mC-nodiv/analysis/Ovis_aries.IROV.MutSel.polyDFE_C.pdf}
    \end{minipage}
    \subsection{Various}

    \begin{minipage}{0.32\linewidth}
        \flushleft {\tiny A: Expected DFE for mutations}
        \includegraphics[width=\linewidth, page=1]{/Users/tlatrille/Documents/SelCoeff/data_processed/opportunities_bins3Mask0.9/DFE.Ovis_aries.ISGC.MutSel.pdf}
    \end{minipage}
    \begin{minipage}{0.32\linewidth}
        \flushleft {\tiny B: Expected DFE for substitutions}
        \includegraphics[width=\linewidth, page=1]{/Users/tlatrille/Documents/SelCoeff/data_processed/opportunities_bins3Mask0.9/DFE.Ovis_aries.ISGC.MutSel.Flow.pdf}
    \end{minipage}
    \begin{minipage}{0.32\linewidth}
        \flushleft {\tiny C: Observed DFE for substitutions}
        \includegraphics[width=\linewidth, page=1]{/Users/tlatrille/Documents/SelCoeff/supp/hist/subs.Ovis_aries.ISGC.bins3Mask0.9.pdf}
    \end{minipage}
    \\
    \begin{minipage}{0.32\linewidth}
        \flushleft {\tiny D: Observed DFE for polymorphisms}
        \includegraphics[width=\linewidth, page=1]{/Users/tlatrille/Documents/SelCoeff/supp/hist/poly.Ovis_aries.ISGC.bins3Mask0.9.pdf}
    \end{minipage}
    \begin{minipage}{0.32\linewidth}
        \flushleft {\tiny E: Site frequency spectrum}
        \includegraphics[width=\linewidth, page=1]{/Users/tlatrille/Documents/SelCoeff/experiments/3bins-mC-nodiv/analysis/Ovis_aries.ISGC.MutSel-sfs.normalize.pdf}
    \end{minipage}
    \begin{minipage}{0.32\linewidth}
        \flushleft {\tiny F: $S$ as function of $S_0$ for each class}
        \includegraphics[width=\linewidth, page=1]{/Users/tlatrille/Documents/SelCoeff/experiments/3bins-mC-nodiv/analysis/Ovis_aries.ISGC.MutSel.polyDFE_C.pdf}
    \end{minipage}
    \subsection{Morocco}

    \begin{minipage}{0.32\linewidth}
        \flushleft {\tiny A: Expected DFE for mutations}
        \includegraphics[width=\linewidth, page=1]{/Users/tlatrille/Documents/SelCoeff/data_processed/opportunities_bins3Mask0.9/DFE.Ovis_aries.MOOA.MutSel.pdf}
    \end{minipage}
    \begin{minipage}{0.32\linewidth}
        \flushleft {\tiny B: Expected DFE for substitutions}
        \includegraphics[width=\linewidth, page=1]{/Users/tlatrille/Documents/SelCoeff/data_processed/opportunities_bins3Mask0.9/DFE.Ovis_aries.MOOA.MutSel.Flow.pdf}
    \end{minipage}
    \begin{minipage}{0.32\linewidth}
        \flushleft {\tiny C: Observed DFE for substitutions}
        \includegraphics[width=\linewidth, page=1]{/Users/tlatrille/Documents/SelCoeff/supp/hist/subs.Ovis_aries.MOOA.bins3Mask0.9.pdf}
    \end{minipage}
    \\
    \begin{minipage}{0.32\linewidth}
        \flushleft {\tiny D: Observed DFE for polymorphisms}
        \includegraphics[width=\linewidth, page=1]{/Users/tlatrille/Documents/SelCoeff/supp/hist/poly.Ovis_aries.MOOA.bins3Mask0.9.pdf}
    \end{minipage}
    \begin{minipage}{0.32\linewidth}
        \flushleft {\tiny E: Site frequency spectrum}
        \includegraphics[width=\linewidth, page=1]{/Users/tlatrille/Documents/SelCoeff/experiments/3bins-mC-nodiv/analysis/Ovis_aries.MOOA.MutSel-sfs.normalize.pdf}
    \end{minipage}
    \begin{minipage}{0.32\linewidth}
        \flushleft {\tiny F: $S$ as function of $S_0$ for each class}
        \includegraphics[width=\linewidth, page=1]{/Users/tlatrille/Documents/SelCoeff/experiments/3bins-mC-nodiv/analysis/Ovis_aries.MOOA.MutSel.polyDFE_C.pdf}
    \end{minipage}

\end{document}