### 2024-12-16 manuscript/S2 File.tex

%TC:ignore
\documentclass{article}
\usepackage[hypcap=false]{caption}
\usepackage{xcolor, colortbl}
\definecolor{RED}{HTML}{EB6231}
\definecolor{BLUE}{HTML}{5D80B4}
\definecolor{LIGHTGREY}{gray}{0.9}
\definecolor{BLUELINK}{HTML}{0645AD}
\definecolor{DARKBLUELINK}{HTML}{0B0080}
\usepackage[colorlinks=false]{hyperref}
\PassOptionsToPackage{hyphens}{url}
% for linking between references, figures, TOC, etc in the pdf document
\hypersetup{colorlinks,
    linkcolor=DARKBLUELINK,
    anchorcolor=DARKBLUELINK,
    citecolor=DARKBLUELINK,
    filecolor=DARKBLUELINK,
    menucolor=DARKBLUELINK,
    urlcolor=BLUELINK
} % Color citation links in purple
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{naturalnames}{hyperref}

\usepackage[backend=biber,eprint=false,isbn=false,url=false,intitle=true,style=nature,date=year]{biblatex}
\addbibresource{codon_models.bib}

\usepackage{bbm}
\usepackage[margin=50pt]{geometry}
\usepackage{amssymb,amsfonts,amsmath,amsthm,mathtools}
\usepackage{lmodern}
\usepackage{bm,bbold}
\usepackage{verbatim}
\usepackage{float}
\usepackage{listings, enumerate, enumitem}
\usepackage[export]{adjustbox}
\usepackage{tabu}
\usepackage{longtable}
\tabulinesep=0.6mm
\newcommand\cellwidth{\TX@col@width}
\usepackage{hhline}
\setlength{\arrayrulewidth}{1.2pt}
\usepackage{multicol,multirow,array}
\usepackage{etoolbox}
\AtBeginEnvironment{tabu}{\footnotesize}
\usepackage{booktabs}
\usepackage{makecell}
\usepackage{orcidlink}
\usepackage{graphicx}
\usepackage{blkarray}
\usepackage{pgf,tikz}
\usetikzlibrary{shapes,arrows,backgrounds,fit,positioning,arrows,automata,calc}
\tikzset{res/.style={ellipse,draw,minimum height=1.0cm,minimum width=0.8cm}}
\tikzset{literal/.style={rectangle,draw,minimum height=0.5cm,minimum width=0.8cm,text width = 1.2 cm, align = center}}

\pdfinclusioncopyfonts=1

\renewcommand{\baselinestretch}{1.5}
\renewcommand{\arraystretch}{0.6}
\frenchspacing

\renewcommand{\thetable}{\Alph{table}}
\renewcommand{\thefigure}{\Alph{figure}}
\renewcommand{\theequation}{S.\arabic{equation}}

\newcommand{\UniDimArray}[1]{\bm{#1}}
\newcommand{\BiDimArray}[1]{\bm{#1}}

\newcommand{\der}{\text{d}}
\newcommand{\e}{\text{e}}
\newcommand{\Ne}{N_{\text{e}}}
\newcommand{\proba}{\mathbb{P}}
\newcommand{\pfix}{\proba_{\text{fix}}}
\newcommand{\dn}{d_N}
\newcommand{\ds}{d_S}
\newcommand{\dnds}{\dn / \ds}
\newcommand{\Sphy}{S_{0}}
\newcommand{\SphyMean}{\overline{\Sphy}}
\newcommand{\SphyDel}{\mathcal{D}_0}
\newcommand{\SphyNeu}{\mathcal{N}_0}
\newcommand{\SphyBen}{\mathcal{B}_0}
\newcommand{\Sphyclass}{x}
\newcommand{\SphyclassAlt}{y}
\newcommand{\given}{\mid}
\newcommand{\Spop}{S}
\newcommand{\SpopDel}{\mathcal{D}}
\newcommand{\SpopNeu}{\mathcal{N}}
\newcommand{\SpopBen}{\mathcal{B}}
\newcommand{\ProbaPopDel}{\proba{[} \SpopDel]}
\newcommand{\ProbaPopNeu}{\proba{[} \SpopNeu ]}
\newcommand{\ProbaPopBen}{\proba{[} \SpopBen ]}
\newcommand{\AdvMean}{\beta_b}
\newcommand{\DelMean}{\beta_d}
\newcommand{\thetaSyn}{\theta_{\text{S}}}
\newcommand{\pvalue}{p\text{-value}}

% Model
\newcommand{\submatrix}{q}
\newcommand{\Submatrix}{\BiDimArray{\submatrix}}
\newcommand{\probmatrix}{P}
\newcommand{\Probmatrix}{\BiDimArray{\probmatrix}}
\newcommand{\fit}{F}
\newcommand{\Fit}{\UniDimArray{\fit}}
\newcommand{\indice}{l}
\newcommand{\indiceexp}{^{(\indice)}}
\newcommand{\ci}{{a}}
\newcommand{\cj}{{b}}
\newcommand{\itoj}{\ci \mapsto \cj}
\newcommand{\nuc}{\mathcal{M}}
\newcommand{\fiti}{\fit_{\ci}}
\newcommand{\fitj}{\fit_{\cj}}
\newcommand{\mutmatrix}{R}
\newcommand{\Mutmatrix}{\BiDimArray{\mutmatrix}}
\newcommand{\exchan}{\rho}
\newcommand{\Exchan}{\UniDimArray{\exchan}}
\newcommand{\mutequi}{\sigma}
\newcommand{\Mutequi}{\UniDimArray{\mutequi}}
\newcommand{\Tree}{\mathcal{T}}
\newcommand{\branch}{\text{j}}
\newcommand{\branchexp}{^{(\branch)}}
\newcommand{\branchlength}{l}
% Alignment
\newcommand{\data}{D}
\newcommand{\Data}{\BiDimArray{\data}}
\newcommand{\site}{\text{i}}
\newcommand{\Nsite}{\text{N}}
\newcommand{\siteexp}{^{(\site)}}
\newcommand{\Setsite}{\site \in \{1, \hdots, \Nsite\} }
\newcommand{\branchsiteexp}{^{(\branch, \site)}}
% Categories
\newcommand{\cat}{\text{k}}
\newcommand{\Ncat}{\text{K}}
\newcommand{\catexp}{^{(\cat)}}
\newcommand{\catInterval}{\{1, \hdots, \Ncat\}}
\newcommand{\Setcat}{\cat \in \catInterval }
\newcommand{\branchcatexp}{^{(\branch, \cat)}}
\newcommand{\profile}{\phi}
\newcommand{\Profile}{\UniDimArray{\profile}}
\newcommand{\concentrationProfile}{\alpha}
\newcommand{\centerProfile}{\UniDimArray{\gamma}}
\newcommand{\catVar}{\kappa}
\newcommand{\catsite}{\catVar\left(\site\right)}
\newcommand{\catmultivar}{m}
\newcommand{\catMultiVar}{\UniDimArray{\catmultivar}}
\newcommand{\stickbreaking}{\theta}
\newcommand{\StickBreaking}{\UniDimArray{\stickbreaking}}
\newcommand{\stick}{\psi}
\newcommand{\stickbreakinghyper}{\beta}
\newcommand{\Multivariate}{\UniDimArray{Z}}
\newcommand{\subhistory}{\mathcal{H}}

\title{\textbf{Estimating the proportion of beneficial mutations that are not adaptive in mammals}}

\author{
    \large
    \textbf{T. {Latrille}$^{1\dag}$\orcidlink{0000-0002-9643-4668}, J. {Joseph}$^{2\dag}$\orcidlink{0009-0002-1312-9930}, D.~A. {Hartasánchez}$^{1}$\orcidlink{0000-0003-2596-6883}, N. {Salamin}$^{1}$\orcidlink{0000-0002-3963-4954}}\\
    \scriptsize $^{1}$Department of Computational Biology, Université de Lausanne, Lausanne, Switzerland\\
    \scriptsize $^{2}$Laboratoire de Biométrie et Biologie Evolutive, UMR5558, Université Lyon 1, Villeurbanne, France \\
    \scriptsize $^{\dag}$These authors contributed equally to this work\\
    \normalsize \texttt{\href{mailto:thibault.latrille@ens-lyon.org}{thibault.latrille@ens-lyon.org}} \\
}

\date{}

\begin{document}
    \maketitle
    \part*{S2 File}
    \vspace{-1em}
    \tableofcontents
    \newpage
    \listoffigures
    \listoftables
    \newpage
    \section{Beneficial mutations in the terminal lineages and populations}\label{sec:beneficial-mutations}

    \subsection{Example sites}\label{subsec:example-sites}

    We extracted protein-coding DNA alignments across mammals centered around the codon site (two flanking codons in white background) for which the beneficial non-adaptive mutations ($\SphyBen$) have been detected in \textit{Chlorocebus sabaeus}.
    We also show the translated amino acids of this region.
    For instance, in DNA alignment of gene SELE (\ref{fig:example-1}), the nucleotide at site 1722 has mutated (from T to C) at the basis of Simiiformes (monkeys and apes), modifying the corresponding amino acid from Serine to Proline, but has been subsequently reverted in the branch of \textit{Chlorocebus sabaeus}.
    However, other substitutions classified as $\SphyBen$ cannot be clearly interpreted as reversions \textit{sensu stricto} along the terminal branch of \textit{Chlorocebus sabaeus}.
    Indeed, we acknowledge that on a fixed fitness landscape, a deleterious mutation can be compensated by transitions to other fitter amino-acids, and not necessarily the ancestral one.

    Finally, we generated the fasta alignment for all $\SphyBen$ mutations that we detected, either in the terminal lineage or in segregating polymorphisms.
    These fasta files are available at Zenodo (\url{https://doi.org/10.5281/zenodo.7878953}), across all populations, under a zip file (\textit{alignment\_around\_non\_adaptive\_mutations.zip}) alongside a python script to obtain the figures (as in Figures~\ref{fig:example-1} and~\ref{fig:example-2}) given these alignments (\textit{plot\_variations.py}).

    \newpage
    \begin{center}
        \includegraphics[width=0.9\linewidth, page=1]{artworks/Alignment.Chlorocebus_sabaeus.Saint_Kitts-0.pdf}
        \captionof{figure}[Examples sites of $\SphyBen$ mutations in \textit{Chlorocebus sabaeus} (reversions).]{\textbf{Examples sites of $\bm{\SphyBen}$ mutations in \textit{Chlorocebus sabaeus}  (reversions).} The first line is the name of the gene, the second line the position of the mutation in the OrthoMam protein-coding DNA alignment, and the arrows in the third line correspond to the position of the mutation in the nucleotide sequence (left) and in the corresponding amino acid sequence (right)\label{fig:example-1}}
    \end{center}

    \newpage
    \begin{center}
        \includegraphics[width=0.9\linewidth, page=1]{artworks/Alignment.Chlorocebus_sabaeus.Saint_Kitts-1.pdf}
        \captionof{figure}[Examples sites of $\SphyBen$ mutations in \textit{Chlorocebus sabaeus}.]{\textbf{Examples sites of $\bm{\SphyBen}$ mutations in \textit{Chlorocebus sabaeus}.} The first line is the name of the gene, the second line the position of the mutation in the OrthoMam protein-coding DNA alignment, and the arrows in the third line correspond to the position of the mutation in the nucleotide sequence (left) and in the corresponding amino acid sequence (right).\label{fig:example-2}}
    \end{center}

    \newpage
    \subsection{Selection along the terminal branches}\label{subsec:summary-table-mutsel}

    \subsubsection{Probability of mutations and substitutions to be \texorpdfstring{$\SphyDel$}{D₀}, \texorpdfstring{$\SphyNeu$}{N₀} or \texorpdfstring{$\SphyBen$}{B₀}}
    Among all the substitutions found in each terminal branch, between 10 and 13\% were $\SphyBen$ ($\proba_{div}[ \SphyBen {]}$), while $\SphyBen$ mutations only represent between 0.9 and 1.2\% of all non-synonymous mutations ($\proba[ \SphyBen {]}$), as shown in Table~\ref{table:pdiv}.

    \begin{center}
        \scriptsize
        \begin{longtable*}{|l|l|r|r|r|r|r|r|}
            \toprule
            Population &             Species & $\proba{[} \SphyDel {]}$ & $\proba{[} \SphyNeu {]}$ & $\proba{[} \SphyBen {]}$ & $\proba_{div}[ \SphyDel {]}$ & $\proba_{div}[ \SphyNeu {]}$ & $\proba_{div}[ \SphyBen {]}$ \\
            \midrule
            \endhead
            \midrule
            \multicolumn{8}{r}{{Continued on next page}} \\
            \midrule
            \endfoot

            \bottomrule
            \endlastfoot
            \rowcolor{LIGHTGREY} Equus c. & Equus caballus & $ 0.923$ & $ 0.065$ & $ 0.012$ & $ 0.462$ & $ 0.419$ & $ 0.118$ \\
            Iran & Bos taurus & $ 0.924$ & $ 0.065$ & $ 0.011$ & $ 0.515$ & $ 0.362$ & $ 0.123$ \\
            Uganda & Bos taurus & $ 0.924$ & $ 0.065$ & $ 0.011$ & $ 0.514$ & $ 0.361$ & $ 0.125$ \\
            \rowcolor{LIGHTGREY} Australia & Capra hircus & $ 0.923$ & $ 0.066$ & $ 0.011$ & $ 0.494$ & $ 0.386$ & $ 0.121$ \\
            \rowcolor{LIGHTGREY} France & Capra hircus & $ 0.923$ & $ 0.066$ & $ 0.011$ & $ 0.494$ & $ 0.386$ & $ 0.120$ \\
            \rowcolor{LIGHTGREY} Iran (C. aegagrus) & Capra hircus & $ 0.923$ & $ 0.066$ & $ 0.011$ & $ 0.493$ & $ 0.386$ & $ 0.120$ \\
            \rowcolor{LIGHTGREY} Iran & Capra hircus & $ 0.923$ & $ 0.066$ & $ 0.011$ & $ 0.492$ & $ 0.387$ & $ 0.121$ \\
            \rowcolor{LIGHTGREY} Italy & Capra hircus & $ 0.923$ & $ 0.066$ & $ 0.011$ & $ 0.494$ & $ 0.386$ & $ 0.120$ \\
            \rowcolor{LIGHTGREY} Morocco & Capra hircus & $ 0.923$ & $ 0.066$ & $ 0.011$ & $ 0.491$ & $ 0.387$ & $ 0.122$ \\
            Iran & Ovis aries & $ 0.922$ & $ 0.067$ & $ 0.012$ & $ 0.568$ & $ 0.323$ & $ 0.109$ \\
            Iran (O. orientalis) & Ovis aries & $ 0.922$ & $ 0.067$ & $ 0.011$ & $ 0.573$ & $ 0.320$ & $ 0.108$ \\
            Iran (O. vignei) & Ovis aries & $ 0.922$ & $ 0.067$ & $ 0.012$ & $ 0.567$ & $ 0.325$ & $ 0.109$ \\
            Various & Ovis aries & $ 0.922$ & $ 0.067$ & $ 0.011$ & $ 0.572$ & $ 0.321$ & $ 0.107$ \\
            Morocco & Ovis aries & $ 0.922$ & $ 0.067$ & $ 0.012$ & $ 0.570$ & $ 0.321$ & $ 0.108$ \\
            \rowcolor{LIGHTGREY} Barbados & Chlorocebus sabaeus & $ 0.926$ & $ 0.065$ & $ 0.009$ & $ 0.485$ & $ 0.393$ & $ 0.122$ \\
            \rowcolor{LIGHTGREY} Central Afr. Rep. & Chlorocebus sabaeus & $ 0.926$ & $ 0.065$ & $ 0.009$ & $ 0.485$ & $ 0.391$ & $ 0.124$ \\
            \rowcolor{LIGHTGREY} Ethiopia & Chlorocebus sabaeus & $ 0.926$ & $ 0.065$ & $ 0.009$ & $ 0.484$ & $ 0.393$ & $ 0.124$ \\
            \rowcolor{LIGHTGREY} Gambia & Chlorocebus sabaeus & $ 0.926$ & $ 0.065$ & $ 0.009$ & $ 0.483$ & $ 0.394$ & $ 0.123$ \\
            \rowcolor{LIGHTGREY} Kenya & Chlorocebus sabaeus & $ 0.926$ & $ 0.065$ & $ 0.009$ & $ 0.485$ & $ 0.392$ & $ 0.123$ \\
            \rowcolor{LIGHTGREY} Nevis & Chlorocebus sabaeus & $ 0.926$ & $ 0.065$ & $ 0.009$ & $ 0.484$ & $ 0.393$ & $ 0.123$ \\
            \rowcolor{LIGHTGREY} South Africa & Chlorocebus sabaeus & $ 0.926$ & $ 0.065$ & $ 0.009$ & $ 0.480$ & $ 0.394$ & $ 0.125$ \\
            \rowcolor{LIGHTGREY} Saint Kitts & Chlorocebus sabaeus & $ 0.926$ & $ 0.065$ & $ 0.009$ & $ 0.483$ & $ 0.394$ & $ 0.123$ \\
            \rowcolor{LIGHTGREY} Zambia & Chlorocebus sabaeus & $ 0.926$ & $ 0.065$ & $ 0.009$ & $ 0.485$ & $ 0.393$ & $ 0.123$ \\
            African & Homo sapiens & $ 0.925$ & $ 0.065$ & $ 0.010$ & $ 0.561$ & $ 0.341$ & $ 0.099$ \\
            Admixed American & Homo sapiens & $ 0.925$ & $ 0.065$ & $ 0.010$ & $ 0.561$ & $ 0.340$ & $ 0.099$ \\
            East Asian & Homo sapiens & $ 0.925$ & $ 0.065$ & $ 0.010$ & $ 0.560$ & $ 0.341$ & $ 0.098$ \\
            European & Homo sapiens & $ 0.925$ & $ 0.065$ & $ 0.010$ & $ 0.562$ & $ 0.340$ & $ 0.098$ \\
            South Asian & Homo sapiens & $ 0.925$ & $ 0.065$ & $ 0.010$ & $ 0.561$ & $ 0.341$ & $ 0.099$ \\
        \end{longtable*}
    \captionof{table}[Probability of mutations and substitutions to be $\SphyDel$, $\SphyNeu$ or $\SphyBen$.]{\textbf{Probability of mutations and substitutions to be $\bm{\SphyDel}$, $\bm{\SphyNeu}$ or $\bm{\SphyBen}$.}
    $\proba{[} \SphyDel {]}$ (eq.~5) is the probability for a new mutation to be deleterious.
     These mutations have a selection coefficient predicted at the phylogenetic-scale lower than -1, thus toward a less fit amino-acid.
    $\proba{[} \SphyNeu {]}$ (eq.~5) is the probability for a new mutation to be nearly-neutral.
    These mutations have a selection coefficient predicted at the phylogenetic-scale between -1 and 1.
    $\proba{[} \SphyBen {]}$ (eq.~5) is the probability for a new mutation to be non-adaptive beneficial.
    These mutations have a selection coefficient predicted at the phylogenetic-scale larger than 1, thus toward a more fit amino-acid.
    $\proba_{div}[\SphyDel{]}$ is the proportion of substitutions in the terminal branch that are $\SphyDel$.
    $\proba_{div}[\SphyNeu{]}$ is the proportion of substitutions in the terminal branch that are $\SphyNeu$.
    $\proba_{div}[\SphyBen{]}$ is the proportion of substitutions in the terminal branch that are $\SphyBen$.\label{table:pdiv}}
    \end{center}
    \newpage

    \subsubsection{\texorpdfstring{$\dnds$}{dₙ/dₛ} for \texorpdfstring{$\SphyDel$}{D₀}, \texorpdfstring{$\SphyNeu$}{N₀} or \texorpdfstring{$\SphyBen$}{B₀}}
    Theoretically, $\omega = \dn / \ds$ can be related to the underlying scaled selection coefficient ($S$) with the relation $\omega = S /(1-\exp(-S))$ as in \cite[eq. 3]{nielsen_estimating_2003}.
    In our experiments, our observed $\dn(\SphyBen) / \ds$ is within the range of $1.169$ (\textit{C. hircus}) to 1.745 (\textit{H. sapiens}), as shown in Table~\ref{table:dnds}, translating into an average $S$ of $\approx0.32$ to $\approx1.24$, so indeed only slightly advantageous.\\

    Of note, observing $\dn(\SphyBen) / \ds >1$ is an important check that $\SphyBen$ mutations are indeed positively selected with an increased substitution rate.
    Since $\dn(\SphyNeu) / \ds$ is close to 1 for the predicted nearly-neutral mutations, this means that the selection coefficients predicted at the mutation-selection balance are good proxies of selection, then observing $\dn(\SphyBen) / \ds > 1$ is evidence of positive selection of predicted non-adaptive beneficial mutations.
    \begin{center}
        \scriptsize
        \begin{longtable*}{|l|r|r|r|r|}
            \toprule
            Species & $\dnds $ & $\dn ( \SphyDel ) / \ds$ & $\dn ( \SphyNeu ) / \ds$ & $\dn ( \SphyBen ) / \ds$ \\
            \midrule
            \endhead
            \midrule
            \multicolumn{5}{r}{{Continued on next page}} \\
            \midrule
            \endfoot

            \bottomrule
            \endlastfoot
            \rowcolor{LIGHTGREY} Equus caballus & $ 0.129$ & $ 0.065$ & $ 0.832$ & $ 1.267$ \\
            Bos taurus & [$ 0.114$, $ 0.116$] & [$ 0.063$, $ 0.064$] & [$ 0.629$, $ 0.638$] & [$ 1.280$, $ 1.328$] \\
            \rowcolor{LIGHTGREY} Capra hircus & [$ 0.108$, $ 0.109$] & [$ 0.058$, $ 0.058$] & [$ 0.631$, $ 0.636$] & [$ 1.169$, $ 1.183$] \\
            Ovis aries & [$ 0.127$, $ 0.129$] & [$ 0.078$, $ 0.080$] & [$ 0.619$, $ 0.621$] & [$ 1.201$, $ 1.217$] \\
            \rowcolor{LIGHTGREY} Chlorocebus sabaeus & [$ 0.118$, $ 0.119$] & [$ 0.062$, $ 0.062$] & [$ 0.713$, $ 0.720$] & [$ 1.521$, $ 1.577$] \\
            Homo sapiens & [$ 0.170$, $ 0.170$] & [$ 0.103$, $ 0.103$] & [$ 0.884$, $ 0.888$] & [$ 1.733$, $ 1.745$] \\
        \end{longtable*}
    \captionof{table}[$\dnds$ for $\SphyDel$, $\SphyNeu$ or $\SphyBen$.]{\textbf{$\bm{\dnds}$ for $\bm{\SphyDel}$, $\bm{\SphyNeu}$ or $\bm{\SphyBen}$.}
    $\dnds$ (eq.~6) is the ratio of non-synonymous over synonymous substitutions estimated for all the non-synonymous substitutions in the terminal branch.
    $\dn(\SphyDel) / \ds$ (eq.~6) is the ratio of non-synonymous over synonymous substitutions, when restricted to non-synonymous substitutions in the terminal branch that are $\SphyDel$.
    $\dn(\SphyNeu) / \ds$ (eq.~6) is the ratio of non-synonymous over synonymous substitutions, when restricted to non-synonymous substitutions in the terminal branch that are $\SphyNeu$.
    $\dn(\SphyBen) / \ds$ (eq.~6) is the ratio of non-synonymous over synonymous substitutions, when restricted to non-synonymous substitutions in the terminal branch that are $\SphyBen$.
    SNPs are considered fixed (as a substitution) in the populations if all sampled individuals are homozygous for the derived allele.
    This effect results in $\dn / \ds$ varying across populations, denoted as a range per species.
    $\dn(\SphyBen) / \ds$ values observed are above 1 and are consistent with slightly advantageous mutations.\label{table:dnds}}
    \end{center}

    \newpage
    \subsubsection{\texorpdfstring{$\dnds$}{dₙ/dₛ} over-estimation due to non-adaptive beneficial mutations}
    We estimated that between $\approx9$ and $\approx12\%$ of $\dnds$ is over-estimated, corresponding to non-adaptive beneficial mutations inflating the $\dnds$ statistic, as shown in Table~\ref{table:dnds-delta}.
    \begin{center}
        \scriptsize
        \begin{longtable*}{|l|r|r|r|}
            \toprule
            Species & $\dnds $ & $\dn(\Sphy < 1) / \ds$ & $\delta(\dnds )$ \\
            \midrule
            \endhead
            \midrule
            \multicolumn{4}{r}{{Continued on next page}} \\
            \midrule
            \endfoot

            \bottomrule
            \endlastfoot
            \rowcolor{LIGHTGREY} Equus caballus & $ 0.129$ & $ 0.115$ & $  10.7$ \\
            Bos taurus & [$ 0.114$, $ 0.116$] & [$ 0.101$, $ 0.102$] & [$  11.3$, $  11.5$] \\
            \rowcolor{LIGHTGREY} Capra hircus & [$ 0.108$, $ 0.109$] & [$ 0.096$, $ 0.097$] & [$  11.0$, $  11.2$] \\
            Chlorocebus sabaeus & [$ 0.118$, $ 0.119$] & [$ 0.104$, $ 0.105$] & [$  11.4$, $  11.7$] \\
            \rowcolor{LIGHTGREY} Homo sapiens & [$ 0.170$, $ 0.170$] & [$ 0.154$, $ 0.155$] & [$ 8.926$, $ 8.996$] \\
            Ovis aries & [$ 0.127$, $ 0.129$] & [$ 0.115$, $ 0.117$] & [$ 9.663$, $ 9.894$] \\
        \end{longtable*}
        \captionof{table}[$\dnds$ over-estimation due to non-adaptive beneficial mutations.]{\textbf{$\bm{\dnds}$ over-estimation due to non-adaptive beneficial mutations.}
        $\dnds$ (eq.~6) is the ratio of non-synonymous over synonymous substitutions estimated for all the non-synonymous substitutions in the terminal branch.
        $\dn(\Sphy < 1) / \ds$ (eq.~6) is the ratio of non-synonymous over synonymous substitutions, when restricted to non-synonymous substitutions in the terminal branch that are not $\SphyBen$.
        This is the estimated divergence when we remove non-adaptive beneficial mutations.
        $\delta(\dnds)$ (eq.~7) is the fraction of the divergence ($\dnds$) that is over-estimated: the difference between $\dnds$ and $\dn(\Sphy < 1) / \ds$.\label{table:dnds-delta}}
    \end{center}
    \newpage
    \section{Gene ontology enrichment}

    \subsection{Gene ontology enrichment in \texorpdfstring{$\SphyBen$}{B₀} SNPs}

    To assess whether $\SphyBen$ SNPs are enriched in specific gene functions, we used gene ontology (GO) terms at the gene level.
    Thus, for each $\SphyBen$ SNP detected in a focal population, we annotated this SNP with GO terms inherited from the gene in which the SNP is located.
    We restricted this analysis solely on genes annotated with at least one GO term.
    To assess the weight of each GO term in the set of $\SphyBen$ SNPs, we computed the proportion of $\SphyBen$ SNPs annotated with a given GO term.
    We then tested if the distribution of $\SphyBen$ SNPs is different between genes sharing a common GO term and the rest of the genes, as shown in Table~\ref{table:ontology}.

    \begin{center}
        \scriptsize
        \begin{longtable*}{|l|l|r|r|r|r|r|}
            \toprule
            GO id & GO name & $p$ & $r$ & Mann-Whitney U & $p_{\mathrm{v}}$ & $p_{\mathrm{v}}^{\mathrm{adj}}$ \\
            \midrule
            \endhead
            \midrule
            \multicolumn{7}{r}{Continued on next page} \\
            \midrule
            \endfoot
            \bottomrule
            \endlastfoot
            GO:0003777 & microtubule motor activity & $ 0.027$ & $ 1.802$ & $1.1\times 10^{5}$ & $1.8\times 10^{-6}$ & $\bm{0.0008{^*}}$ \\
            GO:0005578 & proteinaceous extracellular matrix & $ 0.069$ & $ 2.071$ & $4.7\times 10^{5}$ & $1.9\times 10^{-6}$ & $\bm{0.00084{^*}}$ \\
            GO:0003774 & motor activity & $ 0.023$ & $ 1.127$ & $1.1\times 10^{5}$ & $6.7\times 10^{-5}$ & $\bm{ 0.029{^*}}$ \\
            GO:0030198 & extracellular matrix organization & $ 0.039$ & $ 1.956$ & $2.7\times 10^{5}$ & $0.00055$ & $ 0.239~~$ \\
            GO:0007018 & microtubule-based movement & $ 0.023$ & $ 0.924$ & $1.3\times 10^{5}$ & $0.00073$ & $ 0.315~~$ \\
            GO:0005576 & extracellular region & $ 0.178$ & $ 2.195$ & $1.9\times 10^{6}$ & $0.00083$ & $ 0.359~~$ \\
            GO:0006898 & receptor-mediated endocytosis & $ 0.027$ & $ 1.803$ & $1.7\times 10^{5}$ & $ 0.001$ & $ 0.512~~$ \\
            GO:0006355 & regulation of transcription & $ 0.069$ & $ 0.275$ & $1.9\times 10^{6}$ & $ 0.002$ & $ 0.815~~$ \\
            GO:0005604 & basement membrane & $ 0.019$ & $ 1.638$ & $1.1\times 10^{5}$ & $ 0.002$ & $ 0.852~~$ \\
            GO:0007229 & integrin-mediated signaling pathway & $ 0.019$ & $ 1.930$ & $1.1\times 10^{5}$ & $ 0.003$ & $ 1.000~~$ \\
            GO:0016887 & ATPase activity & $ 0.031$ & $ 1.149$ & $1.8\times 10^{5}$ & $ 0.003$ & $ 1.000~~$ \\
            GO:0007155 & cell adhesion & $ 0.062$ & $ 1.438$ & $5.8\times 10^{5}$ & $ 0.005$ & $ 1.000~~$ \\
            GO:0005634 & nucleus & $ 0.266$ & $ 0.557$ & $3.9\times 10^{6}$ & $ 0.005$ & $ 1.000~~$ \\
            GO:0006351 & transcription & $ 0.073$ & $ 0.339$ & $1.9\times 10^{6}$ & $ 0.008$ & $ 1.000~~$ \\
            GO:0005654 & nucleoplasm & $ 0.124$ & $ 0.443$ & $2.6\times 10^{6}$ & $ 0.009$ & $ 1.000~~$ \\
            GO:0006366 & transcription from RNA polymerase II promoter & $ 0.008$ & $ 0.063$ & $6.2\times 10^{5}$ & $ 0.011$ & $ 1.000~~$ \\
            GO:0005829 & cytosol & $ 0.212$ & $ 0.541$ & $3.5\times 10^{6}$ & $ 0.012$ & $ 1.000~~$ \\
            GO:0016853 & isomerase activity & $ 0.015$ & $ 4.066$ & $ 1\times 10^{5}$ & $ 0.015$ & $ 1.000~~$ \\
            GO:0071356 & cellular response to tumor necrosis factor & $ 0.015$ & $ 7.069$ & $ 1\times 10^{5}$ & $ 0.016$ & $ 1.000~~$ \\
            GO:0005518 & collagen binding & $ 0.015$ & $ 2.900$ & $ 1\times 10^{5}$ & $ 0.017$ & $ 1.000~~$ \\
            GO:0043087 & regulation of GTPase activity & $ 0.015$ & $ 2.025$ & $1.1\times 10^{5}$ & $ 0.027$ & $ 1.000~~$ \\
            GO:0000139 & Golgi membrane & $ 0.012$ & $ 0.214$ & $6.3\times 10^{5}$ & $ 0.028$ & $ 1.000~~$ \\
            GO:0043235 & receptor complex & $ 0.023$ & $ 0.935$ & $1.9\times 10^{5}$ & $ 0.028$ & $ 1.000~~$ \\
            GO:0000981 & RNA polymerase II transcription factor activity & $ 0.027$ & $ 0.452$ & $9.1\times 10^{5}$ & $ 0.029$ & $ 1.000~~$ \\
            GO:0005739 & mitochondrion & $ 0.058$ & $ 0.471$ & $1.4\times 10^{6}$ & $ 0.029$ & $ 1.000~~$ \\
            GO:0005581 & collagen trimer & $ 0.015$ & $ 2.124$ & $1.1\times 10^{5}$ & $ 0.036$ & $ 1.000~~$ \\
            GO:0051015 & actin filament binding & $ 0.019$ & $ 0.912$ & $1.5\times 10^{5}$ & $ 0.036$ & $ 1.000~~$ \\
            GO:0001843 & neural tube closure & $ 0.015$ & $ 1.164$ & $1.1\times 10^{5}$ & $ 0.036$ & $ 1.000~~$ \\
            GO:0005737 & cytoplasm & $ 0.317$ & $ 0.692$ & $4.1\times 10^{6}$ & $ 0.037$ & $ 1.000~~$ \\
            GO:0043565 & sequence-specific DNA binding & $ 0.015$ & $ 0.308$ & $ 6\times 10^{5}$ & $ 0.039$ & $ 1.000~~$ \\
            GO:0004222 & metalloendopeptidase activity & $ 0.019$ & $ 1.188$ & $1.6\times 10^{5}$ & $ 0.043$ & $ 1.000~~$ \\
            GO:0007166 & cell surface receptor signaling pathway & $ 0.031$ & $ 2.871$ & $ 3\times 10^{5}$ & $ 0.046$ & $ 1.000~~$ \\
            GO:0007156 & homophilic cell adhesion via plasma membrane... & $ 0.015$ & $ 1.013$ & $1.2\times 10^{5}$ & $ 0.050$ & $ 1.000~~$ \\
            GO:0007286 & spermatid development & $ 0.015$ & $ 2.497$ & $1.2\times 10^{5}$ & $ 0.051$ & $ 1.000~~$ \\
            GO:0008202 & steroid metabolic process & $ 0.015$ & $ 2.755$ & $1.2\times 10^{5}$ & $ 0.051$ & $ 1.000~~$ \\
            GO:0061024 & membrane organization & $ 0.019$ & $ 1.516$ & $1.6\times 10^{5}$ & $ 0.052$ & $ 1.000~~$ \\
            GO:1901796 & regulation of signal transduction by p53 class... & $ 0.019$ & $ 1.333$ & $1.6\times 10^{5}$ & $ 0.058$ & $ 1.000~~$ \\
            GO:0006397 & mRNA processing & $ 0.004$ & $ 0.210$ & $3.5\times 10^{5}$ & $ 0.061$ & $ 1.000~~$ \\
            GO:0006915 & apoptotic process & $ 0.015$ & $ 0.463$ & $6.4\times 10^{5}$ & $ 0.065$ & $ 1.000~~$ \\
            GO:0043547 & positive regulation of GTPase activity & $ 0.035$ & $ 2.651$ & $3.7\times 10^{5}$ & $ 0.068$ & $ 1.000~~$ \\
            GO:0007010 & cytoskeleton organization & $ 0.019$ & $ 1.771$ & $1.7\times 10^{5}$ & $ 0.068$ & $ 1.000~~$ \\
            GO:0030308 & negative regulation of cell growth & $ 0.015$ & $ 3.055$ & $1.3\times 10^{5}$ & $ 0.070$ & $ 1.000~~$ \\
            GO:0007160 & cell-matrix adhesion & $ 0.015$ & $ 1.046$ & $1.2\times 10^{5}$ & $ 0.073$ & $ 1.000~~$ \\
            GO:0006897 & endocytosis & $ 0.023$ & $ 1.408$ & $2.2\times 10^{5}$ & $ 0.073$ & $ 1.000~~$ \\
            GO:0003677 & DNA binding & $ 0.077$ & $ 0.448$ & $1.7\times 10^{6}$ & $ 0.075$ & $ 1.000~~$ \\
            GO:0043202 & lysosomal lumen & $ 0.015$ & $ 1.522$ & $1.3\times 10^{5}$ & $ 0.075$ & $ 1.000~~$ \\
            GO:0003700 & DNA binding transcription factor activity & $ 0.027$ & $ 0.343$ & $8.7\times 10^{5}$ & $ 0.076$ & $ 1.000~~$ \\
            GO:0008134 & transcription factor binding & $ 0.004$ & $ 0.048$ & $3.2\times 10^{5}$ & $ 0.077$ & $ 1.000~~$ \\
            GO:0005488 & binding & $ 0.046$ & $ 0.641$ & $4.1\times 10^{5}$ & $ 0.078$ & $ 1.000~~$ \\
            GO:0045944 & positive regulation of transcription from RNA... & $ 0.039$ & $ 0.324$ & $1.1\times 10^{6}$ & $ 0.085$ & $ 1.000~~$ \\
            GO:0005794 & Golgi apparatus & $ 0.046$ & $ 0.430$ & $1.2\times 10^{6}$ & $ 0.087$ & $ 1.000~~$ \\
            GO:0016032 & viral process & $ 0.008$ & $ 0.321$ & $4.1\times 10^{5}$ & $ 0.089$ & $ 1.000~~$ \\
            GO:0055085 & transmembrane transport & $ 0.054$ & $ 1.533$ & $5.9\times 10^{5}$ & $ 0.092$ & $ 1.000~~$ \\
            GO:0006260 & DNA replication & $ 0.019$ & $ 1.051$ & $1.8\times 10^{5}$ & $ 0.096$ & $ 1.000~~$ \\
            GO:0030425 & dendrite & $ 0.008$ & $ 0.121$ & $3.9\times 10^{5}$ & $ 0.099$ & $ 1.000~~$ \\
            GO:0005524 & ATP binding & $ 0.127$ & $ 0.666$ & $1.5\times 10^{6}$ & $ 0.103$ & $ 1.000~~$ \\
            GO:0005615 & extracellular space & $ 0.100$ & $ 1.699$ & $1.3\times 10^{6}$ & $ 0.107$ & $ 1.000~~$ \\
            GO:0007283 & spermatogenesis & $ 0.039$ & $ 1.824$ & $3.9\times 10^{5}$ & $ 0.108$ & $ 1.000~~$ \\
            GO:0005179 & hormone activity & $ 0.012$ & $ 7.683$ & $9.9\times 10^{4}$ & $ 0.118$ & $ 1.000~~$ \\
            GO:0098793 & presynapse & $ 0.012$ & $ 2.065$ & $9.6\times 10^{4}$ & $ 0.119$ & $ 1.000~~$ \\
            GO:0005254 & chloride channel activity & $ 0.012$ & $ 1.694$ & $9.6\times 10^{4}$ & $ 0.124$ & $ 1.000~~$ \\
            GO:0036064 & ciliary basal body & $ 0.015$ & $ 1.378$ & $1.4\times 10^{5}$ & $ 0.125$ & $ 1.000~~$ \\
            GO:0070062 & extracellular exosome & $ 0.104$ & $ 0.570$ & $2.1\times 10^{6}$ & $ 0.125$ & $ 1.000~~$ \\
            GO:0090630 & activation of GTPase activity & $ 0.012$ & $ 2.519$ & $9.9\times 10^{4}$ & $ 0.127$ & $ 1.000~~$ \\
            GO:0044212 & transcription regulatory region DNA binding & $ 0.004$ & $ 0.322$ & $2.7\times 10^{5}$ & $ 0.128$ & $ 1.000~~$ \\
            GO:0055114 & oxidation-reduction process & $ 0.019$ & $ 0.482$ & $6.4\times 10^{5}$ & $ 0.129$ & $ 1.000~~$ \\
            GO:0019886 & antigen processing and presentation of exogenous... & $ 0.012$ & $ 1.393$ & $9.6\times 10^{4}$ & $ 0.129$ & $ 1.000~~$ \\
            GO:0005743 & mitochondrial inner membrane & $ 0.008$ & $ 0.463$ & $3.7\times 10^{5}$ & $ 0.133$ & $ 1.000~~$ \\
            GO:0022617 & extracellular matrix disassembly & $ 0.012$ & $ 1.009$ & $9.6\times 10^{4}$ & $ 0.135$ & $ 1.000~~$ \\
            GO:0042493 & response to drug & $ 0.004$ & $ 0.043$ & $2.6\times 10^{5}$ & $ 0.137$ & $ 1.000~~$ \\
            GO:0008380 & RNA splicing & $ 0.008$ & $ 0.098$ & $2.6\times 10^{5}$ & $ 0.139$ & $ 1.000~~$ \\
            GO:0043687 & post-translational protein modification & $ 0.008$ & $ 0.319$ & $3.6\times 10^{5}$ & $ 0.140$ & $ 1.000~~$ \\
            GO:0010951 & negative regulation of endopeptidase activity & $ 0.012$ & $ 2.037$ & $9.9\times 10^{4}$ & $ 0.140$ & $ 1.000~~$ \\
            GO:0000978 & RNA polymerase II proximal promoter... & $ 0.012$ & $ 0.250$ & $4.5\times 10^{5}$ & $ 0.140$ & $ 1.000~~$ \\
            GO:0003713 & transcription coactivator activity & $ 0.004$ & $ 0.339$ & $2.6\times 10^{5}$ & $ 0.142$ & $ 1.000~~$ \\
            GO:0001227 & transcriptional repressor activity & $ 0.012$ & $ 1.384$ & $9.9\times 10^{4}$ & $ 0.144$ & $ 1.000~~$ \\
            GO:0005623 & cell & $ 0.012$ & $ 1.927$ & $ 1\times 10^{5}$ & $ 0.150$ & $ 1.000~~$ \\
            GO:0030574 & collagen catabolic process & $ 0.012$ & $ 1.036$ & $9.9\times 10^{4}$ & $ 0.151$ & $ 1.000~~$ \\
            GO:0006457 & protein folding & $ 0.015$ & $ 2.465$ & $1.5\times 10^{5}$ & $ 0.158$ & $ 1.000~~$ \\
            GO:0016477 & cell migration & $ 0.004$ & $ 0.121$ & $2.4\times 10^{5}$ & $ 0.159$ & $ 1.000~~$ \\
            GO:0070374 & positive regulation of ERK1 and ERK2 cascade & $ 0.019$ & $ 4.238$ & $ 2\times 10^{5}$ & $ 0.160$ & $ 1.000~~$ \\
            GO:0007605 & sensory perception of sound & $ 0.015$ & $ 0.824$ & $1.5\times 10^{5}$ & $ 0.162$ & $ 1.000~~$ \\
            GO:0031225 & anchored component of membrane & $ 0.012$ & $ 0.892$ & $ 1\times 10^{5}$ & $ 0.171$ & $ 1.000~~$ \\
            GO:0016757 & transferase activity & $ 0.004$ & $ 0.435$ & $2.4\times 10^{5}$ & $ 0.177$ & $ 1.000~~$ \\
            GO:0002376 & immune system process & $ 0.012$ & $ 0.411$ & $4.3\times 10^{5}$ & $ 0.177$ & $ 1.000~~$ \\
            GO:0005759 & mitochondrial matrix & $ 0.008$ & $ 0.361$ & $3.3\times 10^{5}$ & $ 0.185$ & $ 1.000~~$ \\
            GO:0010389 & regulation of G2/M transition of mitotic cell... & $ 0.012$ & $ 0.924$ & $ 1\times 10^{5}$ & $ 0.188$ & $ 1.000~~$ \\
            GO:0005667 & transcription factor complex & $ 0.004$ & $ 0.177$ & $2.3\times 10^{5}$ & $ 0.189$ & $ 1.000~~$ \\
            GO:0030054 & cell junction & $ 0.031$ & $ 0.251$ & $8.1\times 10^{5}$ & $ 0.190$ & $ 1.000~~$ \\
            GO:0001650 & fibrillar center & $ 0.019$ & $ 1.540$ & $1.1\times 10^{5}$ & $ 0.191$ & $ 1.000~~$ \\
            GO:0014069 & postsynaptic density & $ 0.004$ & $ 0.172$ & $2.2\times 10^{5}$ & $ 0.200$ & $ 1.000~~$ \\
            GO:0046934 & phosphatidylinositol-4 & $ 0.012$ & $ 1.135$ & $1.1\times 10^{5}$ & $ 0.201$ & $ 1.000~~$ \\
            GO:0007267 & cell-cell signaling & $ 0.004$ & $ 0.354$ & $2.2\times 10^{5}$ & $ 0.201$ & $ 1.000~~$ \\
            GO:0001649 & osteoblast differentiation & $ 0.012$ & $ 1.917$ & $1.1\times 10^{5}$ & $ 0.201$ & $ 1.000~~$ \\
            GO:0045893 & positive regulation of transcription & $ 0.023$ & $ 0.596$ & $6.6\times 10^{5}$ & $ 0.205$ & $ 1.000~~$ \\
            GO:0007399 & nervous system development & $ 0.015$ & $ 0.450$ & $ 5\times 10^{5}$ & $ 0.205$ & $ 1.000~~$ \\
            GO:0001501 & skeletal system development & $ 0.015$ & $ 1.212$ & $1.6\times 10^{5}$ & $ 0.207$ & $ 1.000~~$ \\
            GO:0019903 & protein phosphatase binding & $ 0.012$ & $ 0.875$ & $1.1\times 10^{5}$ & $ 0.208$ & $ 1.000~~$ \\
            GO:1902476 & chloride transmembrane transport & $ 0.012$ & $ 1.458$ & $1.1\times 10^{5}$ & $ 0.210$ & $ 1.000~~$ \\
            GO:0007568 & aging & $ 0.015$ & $ 2.053$ & $1.6\times 10^{5}$ & $ 0.210$ & $ 1.000~~$ \\
        \end{longtable*}
        \captionof{table}[Gene ontology enrichment in $\SphyBen$ SNPs.]{\textbf{Gene ontology enrichment in $\bm{\SphyBen}$ SNPs.} GO terms that are enriched in genes with $\SphyBen$ SNPs in European humans (shown here are the first 100 rows with the lowest $p_{\mathrm{v}}$).
        For each GO term, $p$ is the proportion of $\SphyBen$ SNPs that are annotated with this focal GO term among all annotated $\SphyBen$ SNPs.
        For each gene, $p[ \SphyBen {]}$ is the proportion of sites that have a $\SphyBen$ SNP (can be 0 if no $\SphyBen$ SNP is detected for this gene).
        For each GO term, $r$ is the ratio of the mean of $p[ \SphyBen {]}$ in the set of genes sharing the focal ontology over the mean of $p[ \SphyBen {]}$ in the rest of the genes.
        Mann-Whitney U is the statistic testing if the distribution of $p[ \SphyBen {]}$ is different between the two sets of genes (p-value $p_{\mathrm{v}}$ for the two-sided test).
        $p_{\mathrm{v}}^{\mathrm{adj}}$ are corrected for multiple comparison (Holm–Bonferroni correction across the ontology terms).
        $^*$ for $p_{\mathrm{v}}^{\mathrm{adj}}$ lower than the risk $\alpha=0.05$.\label{table:ontology}
        }
    \end{center}

    \newpage
    \subsection{Enrichment in \texorpdfstring{$\SphyBen$}{B₀} SNPs across all gene ontology terms}
    Whether SNPs classified as $\SphyBen$ are associated to a particular ontology is tested with a Mann-Whitney U statistic as described in Table~\ref{table:ontology}.
    This test is performed across across all 347 gene ontology terms, giving one $p_{\mathrm{v}}$ per ontology (Table~\ref{table:ontology} for the 100 lowest $p_{\mathrm{v}}$).
    The distribution of $p_{\mathrm{v}}$ are not strongly associated to the ontology terms of their respective genes, as shown in the histogram of Figure~\ref{fig:ontology-p}.

    \begin{center}
        \includegraphics[width=0.65\linewidth]{artworks/ontology_poly.Homo_sapiens.EUR.hist.pdf}
        \captionof{figure}[Enrichment in $\SphyBen$ SNPs across all gene ontology terms.]{\textbf{Enrichment in $\bm{\SphyBen}$ SNPs across all gene ontology terms}. In order to test if the distribution of $\SphyBen$ SNPs is different between genes sharing a common GO term and the rest of the genes, we used a Mann-Whitney U test. The distribution of $p_{\mathrm{v}}$ obtained across all 347 gene ontology terms is shown as a histogram.\label{fig:ontology-p}}
    \end{center}

    \newpage

    \section{Clinically related terms for mutations}

    \subsection{Terms associated with deleterious mutations \texorpdfstring{$\SphyDel$}{D₀}}
    SNPs predicted with $\SphyDel$ are statistically associated to clinical terms such as \textit{Likely Pathogenic} and \textit{Pathogenic}, as shown in Table~\ref{table:ontology-neg}.

    \begin{center}
        \begin{tabular}{|l|r|r|r|r|r|}
            \toprule
            SNP clinical ontology & $n_{\mathrm{Observed}}$ & $n_{\mathrm{Expected}}$ & Odds ratio & $p_{\mathrm{v}}$ & $p_{\mathrm{v-adjusted}}$ \\
            \midrule
            Benign                & 2969                    & $4043.0$                & $ 0.734$   & $ 1.000$             & $ 1.000~~$                    \\
            Likely benign         & 2994                    & $3399.8$                & $ 0.881$   & $ 0.999$             & $ 1.000~~$                    \\
            Risk factor           & 102                     & $ 118.2$                & $ 0.863$   & $ 0.798$             & $ 1.000~~$                    \\
            Likely pathogenic     & 221                     & $  68.5$                  & $ 3.226$   & $1.7\times 10^{-8}$  & $\bm{6.7\times 10^{-8}{^*}}$  \\
            Pathogenic            & 560                     & $ 193.6$                & $ 2.893$   & $4.2\times 10^{-17}$ & $\bm{2.1\times 10^{-16}{^*}}$ \\
            \bottomrule
        \end{tabular}
        \captionof{table}[Terms associated with deleterious mutations $\SphyDel$.]{\textbf{Terms associated with deleterious mutations $\bm{\SphyDel}$.}
        In humans (European population), non-synonymous SNPs in the test group ($\SphyDel$) are contrasted to SNPs in the control group ($\SphyNeu$).
        For each clinical term, a 2x2 contingency table is built by counting the number of SNPs based on their selection coefficient and their clinical terms (whether they have this specific term or not).
        Fisher's exact tests are then performed for these 2x2 contingency tables.
        $^*$ for $p_{\mathrm{v}}^{\mathrm{adj}}$ corrected for multiple comparison (Holm–Bonferroni correction) lower than the risk $\alpha=0.05$.\label{table:ontology-neg}}
    \end{center}

    \subsection{Terms associated with non-adaptive beneficial mutations \texorpdfstring{$\SphyBen$}{B₀}}
    Beneficial non-adaptive mutations are associated with clinical terms such as \textit{Benign} and \textit{Likely Benign}, as shown in Table~\ref{table:ontology-pos}.
    \begin{center}
        \begin{tabular}{|l|r|r|r|r|r|}
            \toprule
            SNP clinical ontology & $n_{\mathrm{Observed}}$ & $n_{\mathrm{Expected}}$ & Odds ratio & $p_{\mathrm{v}}$ & $p_{\mathrm{v-adjusted}}$ \\
            \midrule
            Benign                & 319                     & $ 261.7$                & $ 1.219$   & $ 0.002$         & $\bm{ 0.009{^*}}$         \\
            Likely benign         & 263                     & $ 222.7$                & $ 1.181$   & $ 0.012$         & $\bm{ 0.049{^*}}$         \\
            Risk factor           & 5                       & $ 7.847$                & $ 0.637$   & $ 0.879$         & $ 0.879~~$                \\
            Likely pathogenic     & 7                       & $ 4.552$                & $ 1.538$   & $ 0.227$         & $ 0.682~~$                \\
            Pathogenic            & 16                      & $  12.9$                  & $ 1.241$   & $ 0.268$         & $ 0.682~~$                \\
            \bottomrule
        \end{tabular}
        \captionof{table}[Terms associated with non-adaptive beneficial mutations $\SphyBen$.]{\textbf{Terms associated with non-adaptive beneficial mutations $\bm{\SphyBen}$.}
        In humans (European population), non-synonymous SNPs in the test group ($\SphyBen$) are contrasted to SNPs in the control group ($\SphyNeu$).
        For each clinical term, a 2x2 contingency table is built by counting the number of SNPs based on their selection coefficient and their clinical terms (whether they have this specific term or not).
        Fisher's exact tests are then performed for these 2x2 contingency tables.
        $^*$ for $p_{\mathrm{v}}^{\mathrm{adj}}$ corrected for multiple comparison (Holm–Bonferroni correction) lower than the risk $\alpha=0.05$.\label{table:ontology-pos}}
    \end{center}

    \printbibliography
\end{document}